---
title: "Analysis of scRNA-seq data of Mk(t) mice (GSE250486)"
author: "Annika Viljamaa"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: "kable"
    toc: true
---

## Information

**Publication:** *A reversible epigenetic memory of inflammatory injury controls lineage plasticity and tumor initiation in the mouse pancreas* *Falvo et al. 2023* (<https://www.sciencedirect.com/science/article/pii/S153458072300583X>)

The original paper does not report original code.

## Import libraries

```{r}
library(Seurat)
library(dplyr)
library(tidyverse)
library(scDblFinder)


set.seed(123)
```

## Import data

```{r}

files <-  list.files(path = './GSE250486_RAW/raw_data/', recursive = F, full.names = F)

files <- files[-9]


for (x in files){
  
  name <- gsub("\\_filtered.*", "", x)
  
  cts <- Read10X_h5(paste0('./GSE250486_RAW/raw_data/', x)) 
  
  assign(name, CreateSeuratObject(counts = cts))
 
}


# merge object

seurat_object <- merge(GSM7979742_2524, y=c(GSM7979743_2544,GSM7979744_2546, GSM7979745_2547,GSM7979746_B0599,GSM7979747_B0600,GSM7979748_B0602,GSM7979749_B0603),
      add.cell.ids=c("mktC_2524_12WRec_2DRas.","mktS_2544_12WRec_2DRas.", "mtS_2546_12WRec.", "mtC_2547_12WRec.","mtC_b0599_2DRec.","mtS_b0600_2DRec.","mtS_b0602_2DRec.","mtC_b0603_2DRec."))

seurat_object
```

```{r}
head(seurat_object)
```

### Preprocessing

```{r}

# create columns for metadata
# cell id
seurat_object@meta.data$cell_id <- rownames(seurat_object@meta.data)
length(unique(seurat_object@meta.data$cell_id))

# identity
seurat_object@meta.data$orig.ident <- gsub('\\..*', '', seurat_object@meta.data$cell_id)
seurat_object@meta.data <- separate(seurat_object@meta.data, col = orig.ident,sep='_', into =c('trt', 'orig.ident', 'recovery', 'ras'), remove = F)
seurat_object@meta.data$mouse_model <- gsub('C|S','',seurat_object@meta.data$trt)
seurat_object@meta.data$trt <- gsub('mkt|mt','',seurat_object@meta.data$trt)
seurat_object@meta.data <- tidyr::unite(seurat_object@meta.data, col = recovery, c('recovery', 'ras'), sep='_', remove=T, na.rm=T)

# sanity check
unique(seurat_object@meta.data$orig.ident) 
unique(seurat_object@meta.data$mouse_model)
unique(seurat_object@meta.data$trt)
unique(seurat_object@meta.data$recovery)

head(seurat_object)


```

```{r}
# percent mt
seurat_object[["percent.mt"]]<-PercentageFeatureSet(seurat_object, pattern="^mt-")

plot1 <- FeatureScatter(seurat_object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(seurat_object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

```{r}
# based on the paper
seurat_object <- subset(seurat_object, subset = percent.mt < 10 & nFeature_RNA < 4000 & nFeature_RNA > 800)

head(seurat_object)

seurat_object

```

```{r}
plot1 <- FeatureScatter(seurat_object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(seurat_object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2


# double check filtering
min(seurat_object@meta.data$nFeature_RNA)
min(seurat_object@meta.data$nCount_RNA)
max(seurat_object@meta.data$percent.mt)
```

### Control for doublets

```{r}

for (i in c("2524",  "2544",  "2546",  "2547",  "b0599", "b0600", "b0602", "b0603")){
	sub <- subset(seurat_object, subset = orig.ident == i)
	eval(parse(text=paste("sceDblF_",i," <- scDblFinder(GetAssayData(sub, layer = 'counts'), dbr = 0.07)",sep="")))
	eval(parse(text=paste("score.",i," <- sceDblF_",i,"@colData@listData[['scDblFinder.score']]",sep="")))
	eval(parse(text=paste("names(score.",i,") <- rownames(sceDblF_",i,"@colData)",sep="")))
}

doublets.info <- rbind(sceDblF_2524@colData,sceDblF_2544@colData,sceDblF_2546@colData,sceDblF_2547@colData,sceDblF_b0599@colData,sceDblF_b0600@colData,sceDblF_b0602@colData,sceDblF_b0603@colData)
seurat_object$is.doublet <- doublets.info$scDblFinder.class

nrow(seurat_object@meta.data[seurat_object@meta.data$is.doublet=='doublet',]) # 2134 doublets were called

seurat_object <- subset(seurat_object, subset = is.doublet == 'singlet')



seurat_object

```

### Seurat standard workflow

```{r}


# default parameters
seurat_object<-NormalizeData(seurat_object)
seurat_object<-FindVariableFeatures(seurat_object, selection.method = "vst", nfeatures=2000)


all.genes<-rownames(seurat_object)
seurat_object<-ScaleData(seurat_object, features=all.genes)
seurat_object <- RunPCA(seurat_object, features = VariableFeatures(object = seurat_object))
ElbowPlot(seurat_object)
```

```{r}

seurat_object <- RunUMAP(seurat_object, reduction='pca', dims = 1:20)
seurat_object <- FindNeighbors(seurat_object, reduction = 'pca', dims = 1:20)

seurat_object <- FindClusters(seurat_object, resolution = c(1.0))


DimPlot(seurat_object, reduction='pca', label = TRUE)
DimPlot(seurat_object, label = TRUE, group.by = 'RNA_snn_res.1')


```

```{r}
Idents(seurat_object) <- 'RNA_snn_res.1'

p1 <- DimPlot(seurat_object, reduction = 'umap')
p2 <- DimPlot(seurat_object, reduction = 'umap',  group.by = c("orig.ident"))
p3 <- DimPlot(seurat_object, reduction = 'umap',  group.by = c("trt"))
p4 <- DimPlot(seurat_object, reduction = 'umap',  group.by = c("recovery"))


p1+p2+p3+p4

```

#### Integrate samples

```{r}
seurat_object <- IntegrateLayers(object = seurat_object, method = CCAIntegration, assay = "RNA", orig.reduction = "pca", new.reduction = "integrated.cca",
                                 scale.layer = "scale.data",verbose = FALSE)


seurat_object <- RunUMAP(seurat_object, reduction='integrated.cca', dims = 1:20)
seurat_object <- FindNeighbors(seurat_object, reduction = 'integrated.cca', dims = 1:20)
seurat_object <- FindClusters(seurat_object, resolution = c(0.6))
```

```{r}

DimPlot(seurat_object, label = TRUE)
Idents(seurat_object) <- 'RNA_snn_res.0.6'

```

```{r}
p1 <- DimPlot(seurat_object, reduction = 'umap', label = T)
p2 <- DimPlot(seurat_object, reduction = 'umap',  group.by = c("orig.ident"))
p3 <- DimPlot(seurat_object, reduction = 'umap',  group.by = c("trt"))
p4 <- DimPlot(seurat_object, reduction = 'umap',  group.by = c("recovery"))

p1+p2+p3+p4


```

## Identify fibroblasts

```{r}
# fibroblast + CAF markers
VlnPlot(seurat_object, features = c("Col1a1", "Col1a2", "Col3a1",
                                    "Ngfr", "Pdgfra", "Acta2", "Cd74"))

FeaturePlot(seurat_object, features = c("Col1a1", "Col1a2", "Col3a1",
                                    "Ngfr", "Pdgfra", "Acta2", "Cd74", "Plvap"))

```

```{r}

# rename identified clusters
seurat_object <- RenameIdents(object = seurat_object, `0` = "Fibroblasts_1")
seurat_object <- RenameIdents(object = seurat_object, `1` = "Fibroblasts_2")
seurat_object <- RenameIdents(object = seurat_object, `9` = "Fibroblasts_3")
seurat_object <- RenameIdents(object = seurat_object, `16` = "Fibroblasts_4")
seurat_object <- RenameIdents(object = seurat_object, `20` = "Fibroblasts_5")

seurat_object <- RenameIdents(object = seurat_object, `2` = "Ductal/ADM cells_1")
seurat_object <- RenameIdents(object = seurat_object, `10` = "Ductal/ADM cells_2")

seurat_object <- RenameIdents(object = seurat_object, `7` = "Acinar cells")

```

```{r}
# save for easier import
saveRDS(seurat_object, file = "./GSE250486_RAW/raw_data/pancreatitis_preprocessed.rds")

seurat_object <- readRDS("./GSE250486_RAW/raw_data/pancreatitis_preprocessed.rds")
```

## Subset to fibroblasts from acute pancreatitis sample

```{r}
#fibroblasts
seurat_subset <- subset(seurat_object, idents = c("Fibroblasts_1","Fibroblasts_2","Fibroblasts_3", "Fibroblasts_4", "Fibroblasts_5"))
seurat_subset@meta.data$CellType_cluster <- Idents(seurat_subset)
seurat_subset@meta.data$CellType <- gsub('\\_.*','',seurat_subset@meta.data$CellType_cluster)

unique(seurat_subset@meta.data$CellType_cluster)
unique(seurat_subset@meta.data$CellType)

seurat_subset
head(seurat_subset)
```

```{r}
# acute pancreatitis
seurat_subset@meta.data$trt_recovery <- paste(seurat_subset@meta.data$trt, seurat_subset@meta.data$recovery, sep='_')

Idents(seurat_subset) <- 'trt_recovery'

seurat_subset <- subset(seurat_subset, idents = c("C_2DRec","S_2DRec"))
unique(seurat_subset@meta.data$CellType_cluster)
unique(seurat_subset@meta.data$CellType)

seurat_subset
head(seurat_subset)

```

```{r}
p1 <- DimPlot(seurat_subset, reduction = 'umap')
p2 <- DimPlot(seurat_subset, reduction = 'umap', group.by = 'orig.ident')

p1+p2
```

## DE analysis (Ngfr+ fibroblasts vs. the all other fibroblasts)

```{r}
# join layers
seurat_subset <- JoinLayers(seurat_subset)


#Ngfr+ cells
Ngfr_fibroblasts <- WhichCells(subset(x = seurat_subset, subset = Ngfr > 0))

seurat_subset@meta.data$Ngfr_status <- ifelse(seurat_subset@meta.data$cell_id%in%Ngfr_fibroblasts, 'Ngfr+', 'Ngfr-')

# sanity check
nrow(seurat_subset@meta.data[seurat_subset@meta.data$Ngfr_status=='Ngfr+',])
nrow(seurat_subset@meta.data[seurat_subset@meta.data$Ngfr_status=='Ngfr-',])  

seurat_subset

table(seurat_subset@meta.data$orig.ident, seurat_subset@meta.data$trt, seurat_subset@meta.data$Ngfr_status)
```

```{r}
# chi squared test

table <- as.matrix(table(seurat_subset@meta.data$trt, factor(seurat_subset@meta.data$Ngfr_status, levels = c('Ngfr+', 'Ngfr-'))))
table
fisher.test(table)
```

```{r}
Idents(seurat_subset) <- 'Ngfr_status'
Ngfr.markers_all <- FindAllMarkers(seurat_subset, only.pos = FALSE)

Ngfr.markers <- FindMarkers(seurat_subset, ident.1 = "Ngfr+", ident.2 = "Ngfr-", verbose = FALSE)
head(Ngfr.markers, n = 20)

Ngfr.markers
```


```{r}
# gene signature scoring
library(UCell)

gene_ls <- rownames(Ngfr.markers[Ngfr.markers$avg_log2FC>1 & Ngfr.markers$p_val_adj<0.05,])

signatures <- list(Ngfr_fibroblasts=gene_ls)


seurat_subset <- AddModuleScore_UCell(seurat_subset, 
                                      features=signatures, name=NULL)

head(seurat_subset)

```

```{r}

library(wesanderson)
cols <- wes_palette("GrandBudapest1", n = 2)

seurat_subset@meta.data$trt_recovery <- factor(seurat_subset@meta.data$trt_recovery, levels = c("C_2DRec","S_2DRec"))


# edit colors
VlnPlot(seurat_subset, features = c("Ngfr_fibroblasts"), pt.size = 0, group.by = 'trt_recovery', cols = cols)
Idents(seurat_subset) <- 'trt_recovery'
data <- seurat_subset@meta.data%>%
  dplyr::select(orig.ident, trt_recovery, Ngfr_fibroblasts)%>%
  group_by(orig.ident, trt_recovery)%>%
  dplyr::summarise(mean=mean(Ngfr_fibroblasts))


```

## CAF marker expression pattern of Ngfr+ fibroblasts

```{r}

CAFs <- c('Ngfr', 'Pdgfra', 'Acta2', 'Cd74')

# Identify cells expressing selected genes in each cluster
fibroblast_genes <- FetchData(seurat_subset, vars = CAFs, layer = "scale.data")

count_matrix <- t(fibroblast_genes)

# Extract expressed genes as a list
expressed_genes <- apply(count_matrix, 1, function(counts) {
  genes_expressed <- colnames(count_matrix)[counts > 0]
  genes_expressed
})

#subset list to only Ngfr+ cells
Ngfr_cells <- expressed_genes$Ngfr
Ngfr_expressed <- lapply(expressed_genes, function(x) x[x %in% Ngfr_cells])

# Venn diagram
library(ggVennDiagram)
venn <- Ngfr_expressed[c(-1)]
venn_plot_data <- process_data(Venn(venn[-4]), shape_id = '301')
v <- plot_venn(venn_plot_data)+#, label = 'percent', order.intersect.by = 'name')+
  scale_fill_distiller(palette = "YlGnBu")+
  theme_void()
v

# proportional venn
library(BioVenn)
draw.venn(venn$Acta2, venn$Pdgfra, venn$Cd74, title = '', nrtype="abs",
          xtitle = 'Acta2', ytitle = 'Pdgfra', ztitle = 'Cd74')

```
