---
title: "Quant-8 cases_lobular atrophy"
author: "Sara Söderqvist"
date: "2024-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script explores how frequent acinar, earlier acinar - to ductal metaplasia (ADM) and pure ductal cells & fully transformed ADM cells are stratified on lobules in different stages of atrophy. The data stems from an immunofluorescence staining of n = 8 human PDAC resections, with the markers Keratin 19 (K19) and alpha-amylase (Amy). The marker combinations are interpreted as below:

* Amy+ K19- cells = acinar cells
* K19+Amy+ cells = earlier ADM cells
* K19+Amy- cells = "pure" ductal cells *and* fully transformed, late ADM cells which have completely lost the Amy expression.

```{r load in required libraries}
library(ggplot2)
library(ggthemes)
library(forcats)
library(rstatix)
library(ggpubr)
library(RColorBrewer)
library(wesanderson)
```


```{r read in file}
setwd("/Users/sara.soderqvist/Library/CloudStorage/OneDrive-KarolinskaInstitutet/Writing_/Adaptive PDAC paper/Code/source data files to read in/IF K19_Amy") #Adjust working directory

path <- file.path(getwd())
v.filename <- list.files(path, pattern="'*.csv", 
                         ignore.case = TRUE, 
                         full.names = TRUE)

Lobular_stage = do.call(cbind, lapply(v.filename, 
                                function(x) read.csv(x))) #725 obs

```

```{r clean up the data file}
# Clean up annotations residual from the object classifier training
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Acinar", c(colnames(Lobular_stage))] 
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "ADM", c(colnames(Lobular_stage))] 
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "AF488_dotty pattern", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Ductal_Amy_overexpr", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Immune", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Langerhans", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Non_malignant_duct", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "REG3A_neg", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "REG3A_pos", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Stroma", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Stroma_periductal", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Tumor", c(colnames(Lobular_stage))] 
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "cell_cleanup", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "AF488and647_coexpressed_background", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "Stroma_Amy_leakage", c(colnames(Lobular_stage))]
Lobular_stage <- Lobular_stage[Lobular_stage$Class != "", c(colnames(Lobular_stage))] 

# 114 obs

# remove redundant columns
Lobular_stage <- Lobular_stage[, -c(3, 5:8,19)]
Lobular_stage[is.na(Lobular_stage)] <- 0 # if there are no detections, it became NA.
```

```{r explorative plots}
classes <- split(Lobular_stage, f = Lobular_stage$Class)
lapply(classes, nrow)

scatter_cellcounts <- ggplot(Lobular_stage, aes(x = Class, y = Num.Detections, color = Class, size = 3, alpha = 0.9))
scatter_cellcounts + 
  geom_jitter(width = 0.15, height = 0.15) + 
  theme_clean() +
  ggtitle("Jitter Total cell count in all ROIs")

scatter_ROIareas <- ggplot(Lobular_stage, aes(x = Class, y = Area.µm.2, color = Class, size = 3, alpha = 0.9))
scatter_ROIareas + 
  geom_jitter(width = 0.15, height = 0.15) + 
  theme_clean() +
  ggtitle("Jitter Areas of all ROIs")
```


```{r add explorative groupings}
#Calculate ADM fractions
Lobular_stage$ADM_fraction <- (Lobular_stage$Num.ADM/Lobular_stage$Num.Detections)*100
Lobular_stage$Duct_fraction <- ((Lobular_stage$Num.Ductal_Amy_overexpr+Lobular_stage$Num.Non_malignant_duct)/Lobular_stage$Num.Detections)*100
Lobular_stage$Duct_K19spfraction <- (Lobular_stage$Num.Non_malignant_duct/Lobular_stage$Num.Detections)*100

#factorise and set the levels
Lobular_stage$Class <- as.factor(Lobular_stage$Class)
Lobular_stage$Class <- fct_relevel(Lobular_stage$Class, c('Preserved', 'e_atrophy', 'e_atrophy_tu', 'l_atrophy', 'l_atrophy_tu'))
str(Lobular_stage$Class)
```

Not normalised:

```{r summary and statistics ADM per lobualar stage of atrophy}
# The lobular stages here are not pathologist curated. However he prescence of tumor is correct (defined by the consecutive p53 stain, K19 expression and morphology)
## ADM defined by double positive for amylase and K19
summary_lobstage <- Lobular_stage %>%
  group_by(Class) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(ADM_fraction, na.rm = TRUE),
    sd = sd(ADM_fraction, na.rm = TRUE),
    median = median(ADM_fraction, na.rm = TRUE),
    IQR = IQR(ADM_fraction, na.rm = TRUE)
  )
summary_lobstage

hist_lobstage <- Lobular_stage %>%
  ggplot(aes(x = ADM_fraction, color = Class, fill = Class, alpha = 0.6)) +
  geom_histogram(position = 'identity', bins = 150) +
  labs(title = "Fractions of dp cells in lobules") +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_color_brewer(palette = "Dark2", direction = -1)+
  theme_clean()

hist_lobstage

# Kruskal wallis test and post hoc Dunns test
Kruskal_Lobstage <- Lobular_stage %>%
  kruskal_test(ADM_fraction ~ Class)
Kruskal_Lobstage

Dunn_lobstage<- Lobular_stage %>%
  dunn_test(ADM_fraction ~ Class, p.adjust.method = "BH")
Dunn_lobstage

```

```{r boxplot on how frequent ADM cells exist per lobular degradation stage}
# As defined by K19_amy dp
dp_boxplot <- ggboxplot(Lobular_stage, x = "Class", y = "ADM_fraction", color = "Class", fill = "Class", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of Amy and K19 double positive cells in lobules in various stages of atrophy") +
  theme_clean()

#dp_boxplot
Dunn_lobstage_1 <- Dunn_lobstage %>% add_xy_position(x = "Class")
dp_boxplot_with_sign <- dp_boxplot + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1, tip.length = 0) 

dp_boxplot_with_sign
```


```{r summary and statistics ductal cells per lobular stage of atrophy}
## Ductal cells
summary_lobstage_duct <- Lobular_stage %>%
  group_by(Class) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(Duct_fraction, na.rm = TRUE),
    sd = sd(Duct_fraction, na.rm = TRUE),
    median = median(Duct_fraction, na.rm = TRUE),
    IQR = IQR(Duct_fraction, na.rm = TRUE)
  )
summary_lobstage_duct

hist_lobstage_duct <- Lobular_stage %>%
  ggplot(aes(x = Duct_fraction, color = Class, fill = Class, alpha = 0.6)) +
  geom_histogram(position = 'identity', bins = 150) +
  labs(title = "Fractions of ductal cells in lobules") +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_color_brewer(palette = "Dark2", direction = -1)+
  theme_clean()

hist_lobstage_duct

# Kruskal wallis test and post hoc Dunns test
Kruskal_Lobstage_duct <- Lobular_stage %>%
  kruskal_test(Duct_fraction ~ Class)
Kruskal_Lobstage_duct

Dunn_lobstage_duct <- Lobular_stage %>%
  dunn_test(Duct_fraction ~ Class, p.adjust.method = "BH")
Dunn_lobstage_duct
```

```{r boxplot on how frequent ductal cells exist per lobular degradation stage}
dp_boxplot_duct <- ggboxplot(Lobular_stage, x = "Class", y = "Duct_fraction", color = "Class", fill = "Class", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of all ductal cells in lobules in various stages of atrophy") +
  theme_clean()

Dunn_lobstage_1_duct <- Dunn_lobstage_duct %>% add_xy_position(x = "Class")
dp_boxplot_with_sign_duct <- dp_boxplot_duct + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_duct, tip.length = 0) 

dp_boxplot_with_sign_duct
```

```{r summary and statistics ductal cells per lobular stage of atrophy 2}
## Ductal cells
summary_lobstage_ductK19sp <- Lobular_stage %>%
  group_by(Class) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(Duct_K19spfraction, na.rm = TRUE),
    sd = sd(Duct_K19spfraction, na.rm = TRUE),
    median = median(Duct_K19spfraction, na.rm = TRUE),
    IQR = IQR(Duct_K19spfraction, na.rm = TRUE)
  )
summary_lobstage_ductK19sp

hist_lobstage_ductK19sp <- Lobular_stage %>%
  ggplot(aes(x = Duct_K19spfraction, color = Class, fill = Class, alpha = 0.6)) +
  geom_histogram(position = 'identity', bins = 150) +
  labs(title = "Fractions of ductal cells, only the K19 single positives, in lobules") +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_color_brewer(palette = "Dark2", direction = -1)+
  theme_clean()

hist_lobstage_ductK19sp

# Kruskal wallis test and post hoc Dunns test
Kruskal_Lobstage_ductK19sp <- Lobular_stage %>%
  kruskal_test(Duct_K19spfraction ~ Class)
Kruskal_Lobstage_ductK19sp

Dunn_lobstage_ductK19sp <- Lobular_stage %>%
  dunn_test(Duct_K19spfraction ~ Class, p.adjust.method = "BH")
Dunn_lobstage_ductK19sp

```

```{r boxplot on how frequent ductal cells exist per lobular degradation stage 2}
dp_boxplot_ductK19sp <- ggboxplot(Lobular_stage, x = "Class", y = "Duct_K19spfraction", color = "Class", fill = "Class", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of K19 single positive ductal cells in lobules in various stages of atrophy") +
  theme_clean()

Dunn_lobstage_1_ductK19sp <- Dunn_lobstage_ductK19sp %>% add_xy_position(x = "Class")
dp_boxplot_with_sign_ductK19sp <- dp_boxplot_ductK19sp + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_ductK19sp, tip.length = 0) 

dp_boxplot_with_sign_ductK19sp

```

```{r Exploring assigning lobular atrophy based on the stromal cell content}
Lobular_stage$non.parenchyma_fraction <- ((Lobular_stage$Num.Immune+Lobular_stage$Num.Stroma+Lobular_stage$Num.Stroma_Amy_leakage+Lobular_stage$Num.Stroma_periductal)/Lobular_stage$Num.Detections)*100
Lobular_stage$non.parenchyma_fraction_area <- ((Lobular_stage$Num.Immune+Lobular_stage$Num.Stroma+Lobular_stage$Num.Stroma_Amy_leakage+Lobular_stage$Num.Stroma_periductal)/Lobular_stage$Area.µm.2)*100

dotplot_lobstage_cellcount <- ggdotplot(Lobular_stage, x = "Object.ID", y = "non.parenchyma_fraction", color = "Class", fill = "Class", palette = "Dark2", alpha = 0.5, xlab = "Pancreatic lobules", size = 0.7) +
  labs(title = "Fraction of non parenchyma cells per lobule (immune and all stroma classes)") +
  theme_clean()

dotplot_lobstage_cellcount

dotplot_lobstage_area <- ggdotplot(Lobular_stage, x = "Object.ID", y = "non.parenchyma_fraction_area", color = "Class", fill = "Class", palette = "Dark2", alpha = 0.5, xlab = "Pancreatic lobules", size = 0.7) +
  labs(title = "Ratio of non parenchyma cells per lobular AREA (immune and all stroma classes)") +
  theme_clean()

dotplot_lobstage_area
```

```{r acinar cell fraction}
Lobular_stage$acinar_fraction <- ((Lobular_stage$Num.Acinar)/Lobular_stage$Num.Detections)*100

summary_lobstage_acinar <- Lobular_stage %>%
  group_by(Class) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(acinar_fraction, na.rm = TRUE),
    sd = sd(acinar_fraction, na.rm = TRUE),
    median = median(acinar_fraction, na.rm = TRUE),
    IQR = IQR(acinar_fraction, na.rm = TRUE)
  )
summary_lobstage_acinar

Kruskal_Lobstage_acinar <- Lobular_stage %>%
  kruskal_test(acinar_fraction ~ Class)
Kruskal_Lobstage_acinar

Dunn_lobstage_acinar <- Lobular_stage %>%
  dunn_test(acinar_fraction ~ Class, p.adjust.method = "BH")
Dunn_lobstage_acinar

dp_boxplot_acinar <- ggboxplot(Lobular_stage, x = "Class", y = "acinar_fraction", color = "Class", fill = "Class", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of acinar cells in lobules in various stages of atrophy") +
  theme_clean()

Dunn_lobstage_1_acinar <- Dunn_lobstage_acinar %>% add_xy_position(x = "Class")
dp_boxplot_with_sign_acinar <- dp_boxplot_acinar + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_acinar, tip.length = 0) 
dp_boxplot_with_sign_acinar
```

Grouping based on pathologist's classification of lobular atrophy.

Each lobule was scored by collaborating pathologist CFM, blindly to the immunofluorescence stains, on HE and CD146_NGFR consecutive IHC stains to minimal, mild, moderate, severe or end-stage atrophy. The corresponding lobules are then manually assigned the atrophy staging by the individual Object ID. 

```{r classification of lobules by Carlos atrophy staging}
#This is the patholohist blinded setting of lobular class.
Lobular_stage$Atrophy_grade <- Lobular_stage$Object.ID

#remove: fffc7851-ca68-4e93-aa2b-c932cbc00c1b -> likely septum according to Carlos
Lobular_stage <- Lobular_stage[Lobular_stage$Atrophy_grade != "fffc7851-ca68-4e93-aa2b-c932cbc00c1b", c(colnames(Lobular_stage))]

# MINIMAL
Lobular_stage_car <- Lobular_stage %>% mutate(Atrophy_grade = ifelse(grepl("4f5df1de-112c-42ef-aaa9-5077347a4717|667d4dd3-1e5f-46c2-8f66-d309419749a4|9d6e64b0-baf8-450e-8ca7-51f9302c8022|4db6a66e-3c3b-40b2-8920-7baf651822cb|29ba9bdd-cacd-4911-97b4-94c029a75486|ef7c0733-ceb1-48a9-982c-126c057205a7|9ea35e2c-9d21-4761-84a8-682f10542364", .$Object.ID), 'Minimal', .$Atrophy_grade))

# MILD
Lobular_stage_car <- Lobular_stage_car %>% mutate(Atrophy_grade = ifelse(grepl("eea0d12e-bfc2-4d2a-9ba6-5eefcdaddc33|f0847fa5-c885-487d-9b0b-943d70fd68ca|40be8089-7bfd-4bb8-b39a-e285b69f711c|1dcb542e-0285-4724-9d01-7b41ab1cc914|a2133f18-21e8-4577-b73f-2a733ab80920|61770c7b-21a8-47cb-aba3-1d7a38a2b466|6e2c1f2f-97cd-4398-b4ca-327f57fc3ae7|2b430d17-25c6-43c6-882c-53ca549c1f02|8777530f-c793-49c9-9af2-5a437af2bdd8|38d75ac1-c951-4a52-9b13-2c1aef313c7a|c16a0566-0d1a-4512-81ba-0e69fb001361|a7736bb4-58be-466a-bf23-20b954f643f4|96020b39-0e30-467f-a272-b51c1654a8ff|a2f519cf-b0f3-4149-8c72-507e399f1d44|8efa07bd-8a74-4188-acc2-22210d156bb0|31ce80ac-4bb2-41f2-9993-bae4f9c3dd68|832253ff-f068-43ac-9b26-1635940eeca9", .$Object.ID), 'Mild', .$Atrophy_grade))

# MODERATE
Lobular_stage_car <- Lobular_stage_car %>% mutate(Atrophy_grade = ifelse(grepl("4a32da96-db7b-4c6b-a412-d2a831ce5992|3eec5680-b8b2-4fed-b671-84b0cd064ac4|cdd5be29-ad3c-42e7-a1df-35e31a05b1d3|338e823f-2d0b-4646-9429-030efaa4289c|3d76a326-0084-4c04-98d3-fff5507454f4|e37e14cc-b294-4f32-af3b-483e2db315fd|e8d721ca-7784-4511-98c2-eea22729d8f6|3192b9ad-fbc2-44af-a31e-aacd1cfd328e|f85cf50a-6289-4002-8d05-c2bc85accd44|5317cba0-f09a-4ec4-bc4e-d23fb1399436|d5c28a0b-0110-43bc-938d-5643f23d7295|1f3747f9-ea59-42c6-9dcc-70c0ed939287|2c9420a1-a9f8-4cfe-8cb6-00c6b3f21dee|4881ffb5-18b2-49e6-961a-77297a6450db|3006e379-4ff8-4fe3-a625-06c34aefa1f2|4be17a95-6fb2-4424-aefc-2b1bec80884c|eebbf068-3d2a-44ab-8751-2d935d62de52|fee0023c-95d0-4c42-ba3e-c0ecc5109fde|5e3782ec-457e-48e5-a017-340b8b5c0d2b|96800ff8-5020-40bd-9005-edac39faead7|cb80b292-f22d-4af3-aa8d-b5a65e9798d7|91c840cb-f08c-4a7c-91a1-c2ab32905819|3089084c-d518-4ad2-8633-b5717588d769|8d735f93-e532-40fe-aeb0-141d72e442e8|ad6231f9-176d-4d8c-8a09-3f36516f3861|508f898b-56be-4fd1-987f-331c2b55ab5a|75a528b2-0816-4f83-8c82-908efe3c021c|b54ca9ee-eea7-42dc-b121-4b4738c0e78e|076dafa5-83ed-472f-9bfe-922f968ea5b0|7c73c631-7b41-411a-a7c0-94504287ff93|d95f2901-a9f8-44cd-99f3-6f617eed8710|45c2292d-7b50-43f7-aefd-3601f833d238|bb4f5edd-ced1-4658-8737-b2ffd374fc6b|bd577789-1105-4e81-94ea-7ea465980b84|cec9d44f-8fae-4838-9060-3b3e5201ea23|d76d1e2e-b920-4691-afb5-d3bb230649a6|4851291d-9629-401d-9656-25ad390538e7|d15150e8-aae9-4ade-9419-c75f97a60cf7|a2444f31-909d-402c-90e1-a8399a033de5|e6feeb2f-2570-4347-bd9c-327f0e354ccf|e99d1cce-f881-4c7d-9d13-0b3f0c1e0ebd|10ebfb01-122d-404f-9c36-2f5b724ed5da|1a242ca2-31a0-4026-a445-5779fa3b29e7|0e491b24-bcf0-44b0-a49d-aee205e3f490|906797f0-5686-456a-918a-4c233cecd5bd|7c1af83d-3f3b-4c37-9e5f-1564ed714958|dd3f24a3-7e73-4c29-b369-f670b06623fb|e469fb76-acd0-4b9d-adeb-92f6f0b7f970", .$Object.ID), 'Moderate', .$Atrophy_grade))

# SEVERE
Lobular_stage_car <- Lobular_stage_car %>% mutate(Atrophy_grade = ifelse(grepl("a3d5e526-2907-4a6f-a310-48dbe1cb084c|29304c63-9db7-431f-b3cf-b89dc62d3838|5909ace5-2d71-4ff1-931c-89bf4c5aa4f1|89ea4a1e-4111-49d6-949e-bd68ade27f25|4e61afc6-ebf3-4c37-b83a-3f43cba7184a|c16f7887-af99-42ae-bfa7-64e4fc2a9e49|4fad870f-e695-4ffb-9373-d9b6e0389b15|7c8b6acb-82f5-4e7e-9f67-6938d68f83f0|2f07df98-d7fc-4d3c-a2bc-6b84dcf9cd32|7c148311-e3f2-4310-9aa3-64793bd87749|da12c13f-ee4c-4b09-a5e8-f7c1b106afea|f859e148-5c1d-4a8a-92d4-c19f4469c48d|00e718b3-541e-4c67-8978-59dc05bb365f|836cc42a-083c-47fd-a317-5de0af240dff|cd8efc95-708c-48c7-ae55-0b5bd3ae3e8e|5507acd7-1463-4010-9fa4-0a4f43f1c63f|69d339f3-1606-4bdf-85a7-9963196e8ee4|dae602b0-1fe1-4d16-b6f0-244dbfeba280|508dc492-3c2f-447e-9f53-8c16f831dd93|177192d9-b85c-47e4-9c7a-19d413c4d078|d08b57cc-69ee-4a26-88f6-9d404ddbdc88|6a2b6c1f-d680-4886-95ce-df737c7289d6|9dc58cfa-cc24-4fe9-9f8e-9339210b1f4b|687e1d2f-dd0c-4bc3-8f72-41d3daa3ccf0|42169441-3983-4791-ad7f-25ceab4cb4ec|a2bb24e9-a496-4006-a328-56484d0c504d|f8ed6d6a-94a2-4ace-b756-1b9721223ef1|95b34130-f239-47a3-83ee-77cee825df22|463f3550-ccfd-42cb-914e-859778a6f857", .$Object.ID), 'Severe', .$Atrophy_grade))

# END-STAGE
Lobular_stage_car <- Lobular_stage_car %>% mutate(Atrophy_grade = ifelse(grepl("f796285b-3278-4651-8948-f12ad8d75e05|5d7b3e16-4abc-4e78-b1bf-da773cb9d4bb|f3d04f0b-e2aa-4ab2-80f3-21720de5302e|e54ed928-a1d8-4330-9f73-2b3bc0846a4e|cd44e218-3ec5-4e7f-bd54-d87e4c2c19f1|6eae4da2-90db-4ef1-ac9f-aefb4c0ae8e9|09e47ca9-856b-4600-b61f-0ffa6cc12a03|e534fafb-161f-4658-ba95-ac8138d6492e|243ad731-6117-4d78-8b66-206d4ec212b1|7b573dde-575d-47da-9405-b8090d1d92ba|19c86454-cbc3-4231-bfa7-c8ca5ce07dd7|e79c35b1-b501-43c7-b666-17a01ea73518|c7172ef3-e82f-4c6c-b9d2-fe90f9e55372", .$Object.ID), 'Endstage', .$Atrophy_grade))

unique(Lobular_stage_car$Atrophy_grade)
```

```{r plots after atrophy grading }
Lobular_stage_car$Atrophy_grade <- as.factor(Lobular_stage_car$Atrophy_grade)
Lobular_stage_car$Atrophy_grade <- fct_relevel(Lobular_stage_car$Atrophy_grade, c('Minimal', 'Mild', 'Moderate', 'Severe', 'Endstage'))

str(Lobular_stage_car$Atrophy_grade)

dotplot_lobstage_atrophy_cellcount <- ggdotplot(Lobular_stage_car, x = "Object.ID", y = "non.parenchyma_fraction", color = "Atrophy_grade", fill = "Atrophy_grade", palette = "Dark2", alpha = 0.5, xlab = "Pancreatic lobules", size = 0.7) +
  labs(title = "Fraction of non parenchyma cells per lobule (immune and all stroma classes)") +
  theme_clean()

dotplot_lobstage_atrophy_cellcount

## ADM defined by double positive for amylase and K19
summary_lobstage_car <- Lobular_stage_car %>%
  group_by(Atrophy_grade) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(ADM_fraction, na.rm = TRUE),
    sd = sd(ADM_fraction, na.rm = TRUE),
    median = median(ADM_fraction, na.rm = TRUE),
    IQR = IQR(ADM_fraction, na.rm = TRUE)
  )
summary_lobstage_car

hist_lobstage_car <- Lobular_stage_car %>%
  ggplot(aes(x = ADM_fraction, color = Atrophy_grade, fill = Atrophy_grade, alpha = 0.6)) +
  geom_histogram(position = 'identity', bins = 150) +
  labs(title = "Fractions of dp cells in lobules") +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_color_brewer(palette = "Dark2", direction = -1)+
  theme_clean()

hist_lobstage_car

# Kruskal wallis test and post hoc Dunns test
Kruskal_Lobstage_car <- Lobular_stage_car %>%
  kruskal_test(ADM_fraction ~ Atrophy_grade)
Kruskal_Lobstage_car

Dunn_lobstage_car <- Lobular_stage_car %>%
  dunn_test(ADM_fraction ~ Atrophy_grade, p.adjust.method = "BH")
Dunn_lobstage_car

# As defined by K19_amy dp
dp_boxplot_car <- ggboxplot(Lobular_stage_car, x = "Atrophy_grade", y = "ADM_fraction", color = "Atrophy_grade", fill = "Atrophy_grade", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of Amy and K19 double positive cells in lobules in various stages of atrophy") +
  theme_clean()

#dp_boxplot
Dunn_lobstage_1_car <- Dunn_lobstage_car %>% add_xy_position(x = "Atrophy_grade")
dp_boxplot_with_sign_car <- dp_boxplot_car + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_car, tip.length = 0) 

dp_boxplot_with_sign_car

## Ductal cells
summary_lobstage_duct_car <- Lobular_stage_car %>%
  group_by(Atrophy_grade) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(Duct_fraction, na.rm = TRUE),
    sd = sd(Duct_fraction, na.rm = TRUE),
    median = median(Duct_fraction, na.rm = TRUE),
    IQR = IQR(Duct_fraction, na.rm = TRUE)
  )
summary_lobstage_duct_car

# Kruskal wallis test and post hoc Dunns test
Kruskal_Lobstage_duct_car <- Lobular_stage_car %>%
  kruskal_test(Duct_fraction ~ Atrophy_grade)
Kruskal_Lobstage_duct_car

Dunn_lobstage_duct_car <- Lobular_stage_car %>%
  dunn_test(Duct_fraction ~ Atrophy_grade, p.adjust.method = "BH")
Dunn_lobstage_duct_car

dp_boxplot_duct_car <- ggboxplot(Lobular_stage_car, x = "Atrophy_grade", y = "Duct_fraction", color = "Atrophy_grade", fill = "Atrophy_grade", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of all ductal cells in lobules with pathology staged atrophy") +
  theme_clean()

Dunn_lobstage_1_duct_car <- Dunn_lobstage_duct_car %>% add_xy_position(x = "Atrophy_grade")
dp_boxplot_with_sign_duct_car <- dp_boxplot_duct_car + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_duct_car, tip.length = 0) 

dp_boxplot_with_sign_duct_car

## Ductal cells, single positive for K19
summary_lobstage_ductK19sp_car <- Lobular_stage_car %>%
  group_by(Atrophy_grade) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(Duct_K19spfraction, na.rm = TRUE),
    sd = sd(Duct_K19spfraction, na.rm = TRUE),
    median = median(Duct_K19spfraction, na.rm = TRUE),
    IQR = IQR(Duct_K19spfraction, na.rm = TRUE)
  )
summary_lobstage_ductK19sp_car

# Kruskal wallis test and post hoc Dunns test
Kruskal_Lobstage_ductK19sp_car <- Lobular_stage_car %>%
  kruskal_test(Duct_K19spfraction ~ Atrophy_grade)
Kruskal_Lobstage_ductK19sp_car

Dunn_lobstage_ductK19sp_car <- Lobular_stage_car %>%
  dunn_test(Duct_K19spfraction ~ Atrophy_grade, p.adjust.method = "BH")
Dunn_lobstage_ductK19sp_car


dp_boxplot_ductK19sp_car <- ggboxplot(Lobular_stage_car, x = "Atrophy_grade", y = "Duct_K19spfraction", color = "Atrophy_grade", fill = "Atrophy_grade", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of K19 single positive ductal cells in lobules in lobules with pathology staged atrophy") +
  theme_clean()

Dunn_lobstage_1_ductK19sp_car <- Dunn_lobstage_ductK19sp_car %>% add_xy_position(x = "Atrophy_grade")
dp_boxplot_with_sign_ductK19sp_car <- dp_boxplot_ductK19sp_car + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_ductK19sp_car, tip.length = 0) 

dp_boxplot_with_sign_ductK19sp_car

## Acinar cells
summary_lobstage_acinar_car <- Lobular_stage_car %>%
  group_by(Atrophy_grade) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(acinar_fraction, na.rm = TRUE),
    sd = sd(acinar_fraction, na.rm = TRUE),
    median = median(acinar_fraction, na.rm = TRUE),
    IQR = IQR(acinar_fraction, na.rm = TRUE)
  )
summary_lobstage_acinar_car

Kruskal_Lobstage_acinar_car <- Lobular_stage_car %>%
  kruskal_test(acinar_fraction ~ Atrophy_grade)
Kruskal_Lobstage_acinar_car

Dunn_lobstage_acinar_car <- Lobular_stage_car %>%
  dunn_test(acinar_fraction ~ Atrophy_grade, p.adjust.method = "BH")
Dunn_lobstage_acinar_car

dp_boxplot_acinar_car <- ggboxplot(Lobular_stage_car, x = "Atrophy_grade", y = "acinar_fraction", color = "Atrophy_grade", fill = "Atrophy_grade", palette = "Dark2", alpha = 0.5) +
  labs(title = "Fraction of acinar cells in lobules with pathology staged atrophy") +
  theme_clean()

Dunn_lobstage_1_acinar_car <- Dunn_lobstage_acinar_car %>% add_xy_position(x = "Atrophy_grade")
dp_boxplot_with_sign_acinar_car <- dp_boxplot_acinar_car + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_acinar_car, tip.length = 0) 
dp_boxplot_with_sign_acinar_car

```

```{r Acinar, ductal, and ADM cells alltogether}
library(tidyr)
# Joint boxplot. ADM_fraction, Duct_K19spfraction, acinar_fraction in a column
# incl. removing columns that will not be used
Lobular_stage_all3 <- Lobular_stage_car[, c(2, 15, 17, 20, 21)] # 113 obs

Lobular_stage_all3 <- Lobular_stage_all3 %>%
  pivot_longer(cols = c(2:4), names_to = "Cell_type", values_to = "fraction")

Lobular_stage_all3$Group <- paste(Lobular_stage_all3$Atrophy_grade, Lobular_stage_all3$Cell_type, sep = "_")

Lobular_stage_all3$Cell_type <- as.factor(Lobular_stage_all3$Cell_type)
Lobular_stage_all3$Atrophy_grade <- fct_relevel(Lobular_stage_all3$Atrophy_grade, c('acinar_fraction', 'ADM_fraction', 'Duct_K19spfraction'))
str(Lobular_stage_all3$Cell_type)

str(Lobular_stage_all3$Atrophy_grade)


summary_lobstage_all3 <- Lobular_stage_all3 %>%
  group_by(Cell_type) %>%
  dplyr::summarise(
    count = dplyr::n(),
    mean = mean(fraction, na.rm = TRUE),
    sd = sd(fraction, na.rm = TRUE),
    median = median(fraction, na.rm = TRUE),
    IQR = IQR(fraction, na.rm = TRUE)
  )
summary_lobstage_all3


# is there a diff of % of each cell type between each stage of atrophy?

Kruskal_Lobstage_all3 <- Lobular_stage_all3 %>%
  group_by(Cell_type) %>%
  kruskal_test(fraction ~ Atrophy_grade)
Kruskal_Lobstage_all3

Dunn_lobstage_all3 <- Lobular_stage_all3 %>%
  group_by(Cell_type) %>%
  dunn_test(fraction ~ Atrophy_grade, p.adjust.method = "BH")
Dunn_lobstage_all3

Dunn_lobstage_1_all3 <- Dunn_lobstage_all3 %>% add_xy_position(x = "Atrophy_grade", group = "Cell_type")
Dunn_lobstage_1_all3

#color opt #1
dp_boxplot_all3 <- ggboxplot(Lobular_stage_all3, x = "Atrophy_grade", y = "fraction", color = "Cell_type", fill = "Cell_type", alpha = 0.8) +
  labs(title = "Fraction of three celltypes in lobules with pathology staged atrophy") +
    scale_fill_manual(values = wes_palette("GrandBudapest1", n = 4))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 4)) +
  theme_clean()

#color opt #2 matching aiforia overlays to Okabei and Ito's

aiforia_colors <- c(acinar_fraction = "#CC79A7", ADM_fraction = "#55B4E9", Duct_K19spfraction = "#D55E00")

#dp_boxplot_all3 <- ggboxplot(Lobular_stage_all3, x = #"Atrophy_grade", y = "fraction", color = "Cell_type", #fill = "Cell_type", alpha = 0.8) +
#  labs(title = "Fraction of three celltypes in lobules #with pathology staged atrophy") +
#  scale_color_manual(values = aiforia_colors) +
#  scale_fill_manual(values = (aiforia_colors))+
#  theme_clean()

dp_boxplot_with_sign_all3 <- dp_boxplot_all3 + 
  stat_pvalue_manual(label = "p.adj.signif",
                                Dunn_lobstage_1_all3, tip.length = 0, color = "Cell_type", step.group.by = "Cell_type") 
dp_boxplot_with_sign_all3 
```
```{r}
sessionInfo()
```





