---
title: "KC and KPC semiquantitative"
author: "Sara Söderqvist"
date: "2025-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This script is exploring whether the markers HMGA2, GATA6 and Galectin-4 are differentially expressed in:
 - PanINs KC vs tumor in KPC mice
 - Subanalysis for KC mice: High grade vs low-grade PanIN
 - Subanalysis for KPC mice: tumor in the overall tissue locations (adipose vs) stroma vs lobules. 

One section from each mouse was stained for Galectin-4_Vimentin, and one for HMGA2_GATA6 with immunofluorescence, counterstained with DAPI. 

Quantification was performed in a semiquantitative manner by CFM and SSÖ. 
Scores were assigned per image by opening them in Qupath setting an as strict LUT setting as possible in order to display positive vs negative cells for the respective markers. Consequently, scorings were done considering the positive marker distribution, rather than the marker intensity, and compiled manually in an excel-sheet, separate for KC and KPC.

n = 4 KC mice (named: 45, 46, 49 and 56)
n = 4 KPC mice (named: 436-3B, 436-5, 436-7 and 605-4)

A 5-grade score for the semiquantitative analysis in excel was applied as following:

```{r load required libraries}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(reshape2)
library(rstatix)
library(ggpubr)
```


```{r semiquantitive scorings legend}

semiq <- factor(c('++++', '+++', '++', '+', '-'))
numeric_semiq <- (c(4, 3, 2, 1, 0))
explanation <- c("all or almost all cells are positive",
            "more than half of the cells are positive",
            "less than half of the cells are positive",
            "few, scarce cells are positive",
            "all cells are negative")

semiq_legend <- data.frame(semiq, numeric_semiq, explanation)
semiq_legend
```

For images from KC mice, the scoring was applied considering *only* PanINs, separated for high - and -low grade.

For images from KPC mice, the scoring was applied considering *only* the tumor cells. An assignment of the compartment that most tumor cells were present in was also added. 
For the KC mice, since they were PanINs, the dominant compartment is in general implied to be lobular.

```{r read in , results='asis'}
setwd("/Users/sara.soderqvist/Library/CloudStorage/OneDrive-KarolinskaInstitutet/Writing_/Adaptive PDAC paper/Code/source data files to read in/In vivo") # adjust the working directory

KC_og <- read.csv('read in_in vivo_semiq_KC.csv',
                                  sep = ';', header = TRUE)

KPC_og <- read.csv('read in_in vivo_semiq_KPC.csv',
                                  sep = ';', header = TRUE)
# remove residuals from excel export
KC <- KC_og[, -10]
KPC <- KPC_og[-c(11:13),]

KC <- KC %>% mutate_at(c('Gal4_L', 'GATA6_L', 'HMGA2_L', 'Gal4_H', 'GATA6_H', 'HMGA2_H'), as.numeric)
str(KC)
knitr::kable(head(KPC))
knitr::kable(head(KC))
```

```{r means per KPC mouse}
KPC_s <- split(KPC, f = KPC$Mouse)
KPC_meancalc <- lapply(KPC_s, function (a) {
  a$GATA6_mean <- round(mean(a$GATA6))
  a$HMGA2_mean <- round(mean(a$HMGA2))
  a$Gal4_mean <- round(mean(a$Gal4)) 
  a
})
#KPC_meancalc

KPC_mean_b <- bind_rows(KPC_meancalc)
KPC_mean_a <- KPC_mean_b[, -c(2, 4:7)]
KPC_mean <- unique(KPC_mean_a)
KPC_mean # this contains the mean for each KPC mouse.
```

```{r means per KC mouse}
KC_s <- split(KC, f = KC$Mouse)

KC_meancalc <- lapply(KC_s, function (a) {
  a$GATA6_mean <- round((mean(a$GATA6_L, na.rm = TRUE) + mean(a$GATA6_H, na.rm = TRUE))/2) # mean of the means for both High grade and low grade PanINs
  a$HMGA2_mean <- round((mean(a$HMGA2_L, na.rm = TRUE) + mean(a$HMGA2_H, na.rm = TRUE))/2)
  a$Gal4_mean <- round((mean(a$Gal4_L, na.rm = TRUE) + mean(a$Gal4_H, na.rm = TRUE))/2)
  a
})
#KC_meancalc

KC_mean_b <- bind_rows(KC_meancalc)

KC_mean_a <- KC_mean_b[, -c(2, 4:9)]
KC_mean <- unique(KC_mean_a)
KC_mean # this contains the mean for each KC mouse.
```
Bars showing relative frequencies

```{r bars for semiq }
KC_mean$Mouse <- as.factor(KC_mean$Mouse)
KPC_mean$Mouse <- as.factor(KPC_mean$Mouse)
KC_KPCmeans <- bind_rows(KC_mean, KPC_mean)
KC_KPCmeans <- KC_KPCmeans %>% mutate_at(c('GATA6_mean', 'HMGA2_mean', 'Gal4_mean'), as.factor)

mKC_KPCmeans <- melt(KC_KPCmeans, id.vars = c("Mouse", "Mouse_type"), value.name = "Mean_value", variable.name = "Marker")

mKC_KPCmeans$semiq <- factor(nrow(mKC_KPCmeans))
mKC_KPCmeans <- mKC_KPCmeans %>% mutate(semiq = ifelse(grepl('1', .$Mean_value), '+', .$semiq))
mKC_KPCmeans <- mKC_KPCmeans %>% mutate(semiq = ifelse(grepl('2', .$Mean_value), '++', .$semiq))
mKC_KPCmeans <- mKC_KPCmeans %>% mutate(semiq = ifelse(grepl('3', .$Mean_value), '+++', .$semiq))
mKC_KPCmeans <- mKC_KPCmeans %>% mutate(semiq = ifelse(grepl('4', .$Mean_value), '++++', .$semiq))

bars_meanspermouse <- mKC_KPCmeans %>% ggplot(aes(x = Mouse_type, fill = Mean_value)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression KC and KPC mice", subtitle = "Means of all images per mouse, KC: all PanINs, KPC: invasive tumor", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds")  
bars_meanspermouse

bars_meanspermouse_semiq <- mKC_KPCmeans %>% ggplot(aes(x = Mouse_type, fill = semiq)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression KC and KPC mice", subtitle = "Means of all images per mouse. KC: all PanINs, KPC: invasive tumor", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds")
bars_meanspermouse_semiq
```

Focusing on Gal4 and HMGA2

```{r exclude GATA6}
mKC_KPCmeans_GH <- mKC_KPCmeans[mKC_KPCmeans$Marker != "GATA6_mean", ]
mKC_KPCmeans_GH$Marker <- factor(mKC_KPCmeans_GH$Marker, levels = c("HMGA2_mean", "Gal4_mean"))
levels(mKC_KPCmeans_GH$Marker)
unique((mKC_KPCmeans_GH$Marker))

```

```{r wilcox test with stacked barchart HMGA2 and Gal4 in green}

#Ensure numeric for statistical test
mKC_KPCmeans_GH$Mean_value <- as.numeric(mKC_KPCmeans_GH$Mean_value)

# Wilcox test (non-paired) for stacked barchart. Comparison: is there more of the respective markers in either of the two mouse models?
res.wil_KCKPC_GH_bar <-  mKC_KPCmeans_GH %>%
  group_by(Marker) %>%
  wilcox_test(Mean_value ~ Mouse_type)%>%
  adjust_pvalue(method = 'BH')%>%
  add_significance()
res.wil_KCKPC_GH_bar <- res.wil_KCKPC_GH_bar %>% add_xy_position(x = 'Mouse_type')
res.wil_KCKPC_GH_bar [11] <- 1.2 #Making the plotting more compact
res.wil_KCKPC_GH_bar

#here change the Mean_value to factor to enable making a stacked barplot
mKC_KPCmeans_GH$Mean_value <- as.factor(mKC_KPCmeans_GH$Mean_value)

bars_meanspermouse_sign <- mKC_KPCmeans_GH %>% ggplot() +
  geom_bar(aes(x = Mouse_type, fill = Mean_value), position = "fill") +
labs(title = "Degree of expression KC and KPC mice", subtitle = "Means of all images per mouse, KC: all PanINs, KPC: invasive tumor", x = "Marker") +
 facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Greens")  +
  stat_pvalue_manual(res.wil_KCKPC_GH_bar, label = 'p.adj.signif')
bars_meanspermouse_sign
```

# Focusing on the PanINs of KC mice: separating high and low grade

```{r means per KC mouse PanIN}
KC_panincalc <- lapply(KC_s, function (a) {
  a$mean_GATA6_Low <- round((mean(a$GATA6_L, na.rm = TRUE)))
  a$mean_GATA6_High <- round((mean(a$GATA6_H, na.rm = TRUE)))
  a$mean_HMGA2_Low <- round((mean(a$HMGA2_L, na.rm = TRUE)))
  a$mean_HMGA2_High <- round((mean(a$HMGA2_H, na.rm = TRUE)))
  a$mean_Gal4_Low <- round((mean(a$Gal4_L, na.rm = TRUE)))
  a$mean_Gal4_High <- round((mean(a$Gal4_H, na.rm = TRUE)))
  a
})
#KC_panincalc

KC_meanpanin_b <- bind_rows(KC_panincalc)

KC_meanpanin_a <- KC_meanpanin_b[, -c(2, 4:9)]
KC_meanpanin <- unique(KC_meanpanin_a)
KC_meanpanin # this contains the mean for each KC mouse separated for High - and low grade PanIN.

```

Bars of relative frequences for PanINs, separated on high - and lowgrade PanIN.

```{r PanIN barplots}
KC_meanpanin$Mouse <- as.factor(KC_meanpanin$Mouse)
KC_meanpanin$Mouse_type <- as.factor(KC_meanpanin$Mouse_type)
KC_meanpanin <- KC_meanpanin %>% mutate_at(c('mean_GATA6_Low', 'mean_GATA6_High', 'mean_HMGA2_Low', 'mean_HMGA2_High', 'mean_Gal4_Low', 'mean_Gal4_High'), as.factor)

mKC_meanpanin <- melt(KC_meanpanin, id.vars = c("Mouse", "Mouse_type"), value.name = "Mean_value", variable.name = "Marker")

#factor: PanIN differentiation
mKC_meanpanin$PanIN_grade <- factor(nrow(mKC_meanpanin))
mKC_meanpanin <- mKC_meanpanin %>% mutate(PanIN_grade = ifelse(grepl('Low', .$Marker), 'Low_grade', .$PanIN_grade))
mKC_meanpanin <- mKC_meanpanin %>% mutate(PanIN_grade = ifelse(grepl('High', .$Marker), 'High_grade', .$PanIN_grade))

#Now simplify the marker names
mKC_meanpanin$Marker <- as.character(mKC_meanpanin$Marker) # will otherwise return factor levels if 'no' in ifelse statement.
mKC_meanpanin <- mKC_meanpanin %>% mutate(Marker = ifelse(grepl('Gal4', .$Marker), 'Gal4', .$Marker))
mKC_meanpanin <- mKC_meanpanin %>% mutate(Marker = ifelse(grepl('GATA6', .$Marker), 'GATA6', .$Marker))
mKC_meanpanin <- mKC_meanpanin %>% mutate(Marker = ifelse(grepl('HMGA2', .$Marker), 'HMGA2', .$Marker))
mKC_meanpanin$Marker <- as.factor(mKC_meanpanin$Marker) #change back to factor

mKC_meanpanin$semiq <- factor(nrow(mKC_meanpanin))
mKC_meanpanin <- mKC_meanpanin %>% mutate(semiq = ifelse(grepl('0', .$Mean_value), '-', .$semiq))
mKC_meanpanin <- mKC_meanpanin %>% mutate(semiq = ifelse(grepl('1', .$Mean_value), '+', .$semiq))
mKC_meanpanin <- mKC_meanpanin %>% mutate(semiq = ifelse(grepl('2', .$Mean_value), '++', .$semiq))
mKC_meanpanin <- mKC_meanpanin %>% mutate(semiq = ifelse(grepl('3', .$Mean_value), '+++', .$semiq))
mKC_meanpanin <- mKC_meanpanin %>% mutate(semiq = ifelse(grepl('4', .$Mean_value), '++++', .$semiq))

bars_meansperPanIN <- mKC_meanpanin %>% ggplot(aes(x = PanIN_grade, fill = Mean_value)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression in PanINs from KC mice", subtitle = "Means of all images per mouse", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds") 
bars_meansperPanIN

bars_meansperPanIN_semiq <- mKC_meanpanin %>% ggplot(aes(x = PanIN_grade, fill = semiq)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression in PanINs from KC mice", subtitle = "Means of all images per mouse", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds")
bars_meansperPanIN_semiq
```

# Focusing on the KPC mice overall image compartments: Lobules and stroma

```{r KPC mice all rois}

KPC_mean_c <- KPC_mean_b[, -c(8:10)]
KPC_mean_c # this contains the score for each image from the KPC mice
colnames(KPC_mean_c) [7] <- "Compartment"

KPC_mean_c <- KPC_mean_c %>% mutate_at(c('GATA6', 'HMGA2', 'Gal4'), as.factor)

mKPC_mean_c <- melt(KPC_mean_c, id.vars = c("Mouse", "Image", "Mouse_type", "Compartment"), value.name = "Semiq_score", variable.name = "Marker")

mKPC_mean_c$semiq <- factor(nrow(mKPC_mean_c))
mKPC_mean_c <- mKPC_mean_c %>% mutate(semiq = ifelse(grepl('1', .$Semiq_score), '+', .$semiq))
mKPC_mean_c <- mKPC_mean_c %>% mutate(semiq = ifelse(grepl('2', .$Semiq_score), '++', .$semiq))
mKPC_mean_c <- mKPC_mean_c %>% mutate(semiq = ifelse(grepl('3', .$Semiq_score), '+++', .$semiq))
mKPC_mean_c <- mKPC_mean_c %>% mutate(semiq = ifelse(grepl('4', .$Semiq_score), '++++', .$semiq))

bars_KPCcompartment <- mKPC_mean_c %>% ggplot(aes(x = Compartment, fill = Semiq_score)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression of KPC mice", subtitle = "All ROIs, no averaging", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds") 
bars_KPCcompartment

bars_KPCcompartment_semiq <- mKPC_mean_c %>% ggplot(aes(x = Compartment, fill = semiq)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression KPC mice", subtitle = "All ROIs, no averaging", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds")
bars_KPCcompartment_semiq

```

Only 1 ROI was stromal. Pool this with the adipose:

```{r KPC mice all ROIs poolign stromal and adipose.}

mKPC_mean_cs <- mKPC_mean_c
mKPC_mean_cs$Compartments_v2 <- factor(nrow(mKPC_mean_cs))

mKPC_mean_cs <- mKPC_mean_cs %>% mutate(Compartments_v2 = ifelse(grepl('Adipose', .$Compartment), 'Stroma',  .$Compartments_v2))
mKPC_mean_cs <- mKPC_mean_cs %>% mutate(Compartments_v2 = ifelse(grepl('Stroma', .$Compartment), 'Stroma',  .$Compartments_v2))
mKPC_mean_cs <- mKPC_mean_cs %>% mutate(Compartments_v2 = ifelse(grepl('Lobular', .$Compartment), 'Lobule',  .$Compartments_v2))
  
bars_KPCcompartment2 <- mKPC_mean_cs %>% ggplot(aes(x = Compartments_v2, fill = Semiq_score)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression of KPC mice, pooling stromal and adipose", subtitle = "All ROIs, no averaging", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds") 
bars_KPCcompartment2

bars_KPCcompartment2_semiq <- mKPC_mean_cs %>% ggplot(aes(x = Compartments_v2, fill = semiq)) +
  geom_bar(position = "fill") +
  labs(title = "Degree of expression KPC mice, pooling stromal and adipose", subtitle = "All ROIs, no averaging", x = "Marker") +
  facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Reds")
bars_KPCcompartment2_semiq
```

```{r barplot with statistical testing numeric KC vs KPC}
mKC_KPCmeans$Mean_value <- as.numeric(mKC_KPCmeans$Mean_value)

# Wilcox test (non-paired)
res.wil_KCKPC <-  mKC_KPCmeans %>%
  group_by(Marker) %>%
  wilcox_test(Mean_value ~ Mouse_type)%>%
  adjust_pvalue(method = 'BH')%>%
  add_significance()
res.wil_KCKPC <- res.wil_KCKPC %>% add_xy_position(x = 'Mouse_type')
res.wil_KCKPC

boxKCKPC <- mKC_KPCmeans %>%
  ggplot(aes(x = Mouse_type, y = Mean_value))+
  geom_boxplot(outlier.shape = NA, aes(fill = Mouse_type, alpha = 0.8))+
  geom_jitter(aes(color = Mouse_type), alpha = 1, position=position_jitter(0.1), size = 3)+
  scale_color_brewer(palette = 'Accent')+
  scale_fill_brewer(palette = 'Accent')+
  labs(x = '', y = 'Semiquantitative score (0-4)')+
  theme_clean()+
  facet_grid(Marker ~ .) +
  stat_pvalue_manual(res.wil_KCKPC, label = 'p.adj.signif')
boxKCKPC
```

```{r barplot with statistical testing numeric PanINs}
mKC_meanpanin$Mean_value <- as.numeric(mKC_meanpanin$Mean_value)

res.wil_panin <-  mKC_meanpanin %>%
  group_by(Marker) %>%
  wilcox_test(Mean_value ~ PanIN_grade)%>%
  adjust_pvalue(method = 'BH')%>%
  add_significance()
res.wil_panin <- res.wil_panin %>% add_xy_position(x = 'PanIN_grade')
res.wil_panin

box_panin <- mKC_meanpanin %>%
  ggplot(aes(x = PanIN_grade, y = Mean_value))+
  geom_boxplot(outlier.shape = NA, aes(fill = PanIN_grade, alpha = 0.8))+
  geom_jitter(aes(color = PanIN_grade), alpha = 1, position=position_jitter(0.1), size = 3)+
  scale_color_brewer(palette = 'Accent')+
  scale_fill_brewer(palette = 'Accent')+
          labs(x = '', y = 'Semiquantitative score (0-4)')+
  theme_clean()+
  facet_grid(Marker ~ .) +
    stat_pvalue_manual(res.wil_panin, label = 'p.adj.signif')
box_panin
```

```{r box, geom stat and stacked barchart with statistical testing numeric KC vs KPC only Gal4 and HMGA2}

mKC_KPCmeans_GH <- mKC_KPCmeans[mKC_KPCmeans$Marker != "GATA6_mean", ]
mKC_KPCmeans_GH$Marker <- factor(mKC_KPCmeans_GH$Marker, levels = c("HMGA2_mean", "Gal4_mean"))
levels(mKC_KPCmeans_GH$Marker)
unique((mKC_KPCmeans_GH$Marker))

# Wilcox test (non-paired)
res.wil_KCKPC_GH <-  mKC_KPCmeans_GH %>%
  group_by(Marker) %>%
  wilcox_test(Mean_value ~ Mouse_type)%>%
  adjust_pvalue(method = 'BH')%>%
  add_significance()
res.wil_KCKPC_GH <- res.wil_KCKPC_GH %>% add_xy_position(x = 'Mouse_type')
res.wil_KCKPC_GH

boxKCKPC_GH <- mKC_KPCmeans_GH %>%
  ggplot(aes(x = Mouse_type, y = Mean_value))+
  geom_boxplot(outlier.shape = NA, aes(fill = Mouse_type, alpha = 0.8))+
  geom_jitter(aes(color = Mouse_type), alpha = 1, position = position_jitter(0.1), size = 3)+
  scale_color_brewer(palette = 'Accent')+
  scale_fill_brewer(palette = 'Accent')+
  labs(x = '', y = 'Semiquantitative score (0-4)')+
  theme_clean()+
  facet_grid(Marker ~ .) +
  stat_pvalue_manual(res.wil_KCKPC_GH, label = 'p.adj.signif')
boxKCKPC_GH

pointKCKPC_GH <- mKC_KPCmeans_GH %>%
  ggplot(aes(x = Mouse_type, y = Mean_value))+
  geom_count(aes(fill = Mouse_type, color = Mouse_type))+
  scale_color_brewer(palette = 'Accent')+
  scale_fill_brewer(palette = 'Accent')+
  labs(x = '', y = 'Semiquantitative score (0-4)')+
  theme_clean()+
  facet_grid(Marker ~ .) +
  stat_pvalue_manual(res.wil_KCKPC_GH, label = 'p.adj.signif') +
  scale_size(range = c(2, 10))
pointKCKPC_GH
```

```{r barplot with statistical testing numeric PanINs only HMGA2 and Gal4}
mKC_meanpanin_GH <- mKC_meanpanin[mKC_meanpanin$Marker != "GATA6", ]
mKC_meanpanin_GH$Marker <- factor(mKC_meanpanin_GH$Marker, levels = c("HMGA2", "Gal4"))
levels(mKC_meanpanin_GH$Marker)
unique(mKC_meanpanin_GH$Marker)

#wilcox test, unpaired
res.wil_panin_GH <-  mKC_meanpanin_GH %>%
  group_by(Marker) %>%
  wilcox_test(Mean_value ~ PanIN_grade)%>%
  adjust_pvalue(method = 'BH')%>%
  add_significance()
res.wil_panin_GH <- res.wil_panin_GH %>% add_xy_position(x = 'PanIN_grade')
res.wil_panin_GH

box_panin_GH <- mKC_meanpanin_GH %>%
  ggplot(aes(x = PanIN_grade, y = Mean_value))+
  geom_boxplot(outlier.shape = NA, aes(fill = PanIN_grade, alpha = 0.8))+
  geom_jitter(aes(color = PanIN_grade), alpha = 1, position=position_jitter(0.1), size = 3)+
  scale_color_brewer(palette = 'Accent')+
  scale_fill_brewer(palette = 'Accent')+
  labs(x = '', y = 'Semiquantitative score (0-4)')+
  theme_clean()+
  facet_grid(Marker ~ .) +
  stat_pvalue_manual(res.wil_panin_GH, label = 'p.adj.signif')
box_panin_GH

point_panin_GH <- mKC_meanpanin_GH %>%
  ggplot(aes(x = PanIN_grade, y = Mean_value))+
  geom_count(aes(fill = PanIN_grade, color = PanIN_grade))+
  scale_color_brewer(palette = 'Accent')+
  scale_fill_brewer(palette = 'Accent')+
  labs(x = '', y = 'Semiquantitative score (0-4)')+
  theme_clean()+
  facet_grid(Marker ~ .) +
  stat_pvalue_manual(res.wil_panin_GH, label = 'p.adj.signif') +
  scale_size(range = c(2, 10))
point_panin_GH
```

KC PanINS - Gal4 and HMGA2 only, high grade vs low grade with significance indicators

```{r KC PanINs focusing on only Gal4 and GATA6 with sign ind}
#Ensure numeric for statistical test
mKC_meanpanin_GH$Mean_value <- as.numeric(mKC_meanpanin_GH$Mean_value)

# Wilcox test (non-paired) for stacked barchart. Comparison: is there more of the respective markers in either of the two PanIN differentiation grades?
res.wil_KCPanINs_GH_bar <-  mKC_meanpanin_GH %>%
  group_by(Marker) %>%
  wilcox_test(Mean_value ~ PanIN_grade)%>%
  adjust_pvalue(method = 'BH')%>%
  add_significance()
res.wil_KCPanINs_GH_bar <- res.wil_KCPanINs_GH_bar %>% add_xy_position(x = 'PanIN_grade')
res.wil_KCPanINs_GH_bar [11] <- 1.2 #Making the plotting more compact
res.wil_KCPanINs_GH_bar

#here change the Mean_value to factor to enable making a stacked barplot
mKC_meanpanin_GH$Mean_value <- as.factor(mKC_meanpanin_GH$Mean_value)

bars_meanspermouse_sign_KC <- mKC_meanpanin_GH %>% ggplot() +
  geom_bar(aes(x = PanIN_grade, fill = Mean_value), position = "fill") +
labs(title = "Degree of expression of PanINs in KC mice", subtitle = "all PanINs, Means of all images per mouse", x = "Marker") +
 facet_grid(. ~ Marker) +
  theme_igray() +
  scale_fill_brewer(palette = "Greens")  +
  stat_pvalue_manual(res.wil_KCPanINs_GH_bar, label = 'p.adj.signif')
bars_meanspermouse_sign_KC
```

```{r}
sessionInfo()
```


