---
title: "NGFRstroma"
author: "Sara Söderqvist"
date: "2024-02-07"
output: html_document
---
## NGFR analysis

This script explores the expression of stromal cells existing in the vincinity of tumor cells, in pancreatic lobules only. The data stems from a multiplexed immunofluorescence (m-IF) staining of 8 human PDAC resections, of the markers ASMA, NGFR, p53, PDGFRalpha, CD74 and Vimentin. During th eimage analysis, lobular areas were annotated in Qupath, in which cells were detected. The cellular content was enriched for stromal cells by removing cells with round nucleus, and below, cells with p53-reactivity are removed from the analysis. The cellular expression profiles were assigned with classifiers trained in qupath.

The goal is to explore what other stromal markers that NGFR+ cells, co-express.

```{r read in required libraries}
library(dplyr)
library(ggVennDiagram)
library(survminer)
library(dichromat)
library(BioVenn)
library(reshape2)
library(tidyr)
library(ggthemes)
library(wesanderson)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r Data read in_cell_NGFR, echo = FALSE, message = FALSE, results = "hide", warning = FALSE}
setwd("/Users/sara.soderqvist/Library/CloudStorage/OneDrive-KarolinskaInstitutet/Writing_/Adaptive PDAC paper/Code/source data files to read in/m-IF stroma/NGFR") #adjust the working directory to read in the csv file
path <- file.path(getwd())
v.filename <- list.files(path, pattern="'*.csv", 
                         ignore.case = TRUE, 
                         full.names = TRUE)

StromaNGFR = do.call(cbind, lapply(v.filename, 
                                function(x) read.csv(x)))
```


```{r NGFR data clean up}
StromaNGFR$Tumor <- factor(nrow(StromaNGFR))

StromaNGFR <- StromaNGFR %>% mutate(Tumor = ifelse(grepl(' Tumor: ', StromaNGFR$Class), 'Yes', .$Tumor))
StromaNGFR <- StromaNGFR %>% mutate(Tumor = ifelse(grepl(' Not tumor: ', StromaNGFR$Class), 'No', .$Tumor))

#Filter our p53+ contaminations within the annotations
s_tu_StromaNGFR <- split(StromaNGFR, f = StromaNGFR$Tumor)
nottumor_StromaNGFR <- s_tu_StromaNGFR$No #30 538 cells

# Adjust classes to enable grouping
nottumor_StromaNGFR$NGFR_pos <- factor(nrow(nottumor_StromaNGFR))
nottumor_StromaNGFR$ASMA_pos <- factor(nrow(nottumor_StromaNGFR))
nottumor_StromaNGFR$CD74_pos <- factor(nrow(nottumor_StromaNGFR))
nottumor_StromaNGFR$Vim_pos <- factor(nrow(nottumor_StromaNGFR))
nottumor_StromaNGFR$PDGFR_pos <- factor(nrow(nottumor_StromaNGFR))

nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(NGFR_pos = ifelse(grepl(' NGFR_pos: ', nottumor_StromaNGFR$Class), 'Yes', .$NGFR_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(NGFR_pos = ifelse(grepl(' NGFR_neg: ', nottumor_StromaNGFR$Class), 'No', .$NGFR_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(ASMA_pos = ifelse(grepl(': ASMA_pos', nottumor_StromaNGFR$Class), 'true', .$ASMA_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(ASMA_pos = ifelse(grepl(': ASMA_neg', nottumor_StromaNGFR$Class), 'false', .$ASMA_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(CD74_pos = ifelse(grepl('CD74_pos: ', nottumor_StromaNGFR$Class), 'true', .$CD74_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(CD74_pos = ifelse(grepl('CD74_neg: ', nottumor_StromaNGFR$Class), 'false', .$CD74_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(Vim_pos = ifelse(grepl(' Vim_pos: ', nottumor_StromaNGFR$Class), 'true', .$Vim_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(Vim_pos = ifelse(grepl(' Vim_neg: ', nottumor_StromaNGFR$Class), 'false', .$Vim_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(PDGFR_pos = ifelse(grepl(' PDGFR_pos: ', nottumor_StromaNGFR$Class), 'true', .$PDGFR_pos))
nottumor_StromaNGFR <- nottumor_StromaNGFR %>% mutate(PDGFR_pos = ifelse(grepl(' PDGFR_neg: ', nottumor_StromaNGFR$Class), 'false', .$PDGFR_pos))

# Count how many of the NGFR+ cells, that also are positive for other stains. 
# First filter our NGFR- cells
s_NGFR_StromaNGFR <- split(nottumor_StromaNGFR, f = nottumor_StromaNGFR$NGFR_pos)
NGFR_pos <- s_NGFR_StromaNGFR$Yes #6 341 cells

```


```{r venn diagrams}
NGFR_Venn <- NGFR_pos[, c(2, 9:12)]
NGFR_Venn <- NGFR_Venn  %>% mutate(across(ends_with("pos"), as.logical))

#Venn_NGFR <- ggVennDiagram(lapply(NGFR_Venn[, c(2:5)], function (x)
#  which(x == 1)),
#  label = "percent")
#Venn_all_NGFR <- Venn_NGFR +
#  scale_fill_fermenter(type = "seq", palette = "BuPu")+
#  scale_color_brewer(type = "seq", palette = "BuPu")+
#  theme_cleantable() +
#  scale_x_continuous(expand = expansion(mult = .2)) +
#  labs(title = "All regions and all stains")
  

# non-Vimentin
Vennfunc <- lapply(NGFR_Venn[, c(2, 3, 5)], function (x)
  which(x == 1))

Venn_NGFR_nonVim <- ggVennDiagram(Vennfunc)

Venn_all_NGFR_nonVim <- Venn_NGFR_nonVim +
  scale_fill_gradient(low = "#0072B2", high = "#D55E00") 
  scale_colour_gradient(low = "#0072B2", high = "#D55E00") 
  theme_cleantable() 
  labs(title = "All regions, excluding Vim")

#Venn_all_NGFR
Venn_all_NGFR_nonVim
```


```{r venn diagrams with proportional circle size}
# Venn, version where the size of Venn circles correspond to the numbers of cells in the respective group
# colors here correspond to the group identity, instead of nr of cells in each group as above.

propVenn_all_NGFR_nonVim <- draw.venn(list_x = Vennfunc$ASMA_pos, list_y = Vennfunc$CD74_pos, list_z = Vennfunc$PDGFR_pos, title = "Venn diagram showing coexpreesion of all NGFR+ cells", xtitle = "ASMA", ytitle = "CD74", ztitle = "PDGFRa", x_c = "#F1BB7B", z_c = "#D67236", y_c = "#FD6467",  nrtype = "pct")
propVenn_all_NGFR_nonVim
```

```{r count the NGFR co expression}
# first by calculating the positive fractions ROI-wise.
NGFR_cell <- NGFR_pos[, -c(3, 4, 6:8)]

# Melt it and make prettier
molten_stromaNGFR <- melt(NGFR_cell, id = c("Image", "Object.ID", "Parent"))
colnames(molten_stromaNGFR) [4] <- "Stain" #rename, because colnames such as "variable" and "value" will interfere in other commands
colnames(molten_stromaNGFR) [5] <- "Cell_pos"

molten_stromaNGFR <- separate(data = molten_stromaNGFR, col = "Image", into="Image", sep=c("_"), remove=TRUE)
molten_stromaNGFR$Group <- character(nrow(molten_stromaNGFR))
molten_stromaNGFR$Group <- paste(molten_stromaNGFR$Image, molten_stromaNGFR$Parent, molten_stromaNGFR$Stain, sep = "_")

s_molten_stromaNGFR <- split(molten_stromaNGFR, f = molten_stromaNGFR$Group)

# Here considering only single positive profiles, so not considering weather individual cells are double, triple, quadraple positive for other stains.
tot_det_NGFR <- lapply(X = s_molten_stromaNGFR, FUN = function (x) {
  x$sum_detections_ROI <- nrow(x)
  x
})

fun_NGFRfrac <- function (z) {
  nr_pos<- z %>% filter(Cell_pos == "true") # get nr positive cells for each stain
  pos_fr <- nrow(nr_pos)/nrow(z) #calculate the fraction out of all the NGFR+ cell in that specific ROI
  z$pos_fraction <- pos_fr # put it in a column
  z
} 

NGFR_calc_fr <- lapply (X = tot_det_NGFR, FUN = fun_NGFRfrac)
NGFR_calc_frsub <- bind_rows(NGFR_calc_fr)

NGFR_calc_frsub <- NGFR_calc_frsub[, -c(2, 5)] #25 364 rows (1 row = 1 cell)
NGFR_calc_frsub <- unique(NGFR_calc_frsub) # 212 rows shows no data loss (1 row = 1 ROI + stain combinaiton)
NGFR_calc_frsub$Lobule_stage <- NGFR_calc_frsub$Parent

NGFR_calc_frsub$Lobule_stage <- gsub("Lobule_","", as.character(NGFR_calc_frsub$Lobule_stage))
NGFR_calc_frsub$Lobule_stage <- gsub("_A","", as.character(NGFR_calc_frsub$Lobule_stage))
NGFR_calc_frsub$Lobule_stage <- gsub("_B","", as.character(NGFR_calc_frsub$Lobule_stage))
NGFR_calc_frsub$Lobule_stage <- gsub("_C","", as.character(NGFR_calc_frsub$Lobule_stage))
NGFR_calc_frsub$Lobule_stage <- gsub("_D","", as.character(NGFR_calc_frsub$Lobule_stage))
NGFR_calc_frsub$Lobule_stage <- gsub("_E","", as.character(NGFR_calc_frsub$Lobule_stage))
NGFR_calc_frsub$Lobule_stage <- as.factor(NGFR_calc_frsub$Lobule_stage)

levels(NGFR_calc_frsub$Lobule_stage) <- c("wellp", "partiallyd", "endstage") # not applied in the end but keeping this information

```


```{r plot ngfr single positives}
#Exploratory distribution plots
density_NGFR<- ggdensity(NGFR_calc_frsub, x = "pos_fraction", y = "density", 
                        color = "Stain", fill = "Stain", alpha=0.6) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 4))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 4)) +
  labs(title = "NGFR+ cells expression of other stains") +
  xlab(NULL) + ylab("Density of average positive cells per ROI") +
  theme_clean()
  
density_NGFR

densitycount_NGFR<- ggdensity(NGFR_calc_frsub, x = "pos_fraction", y = "count", 
                        color = "Stain", fill = "Stain", alpha=0.6) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 4))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 4)) +
  labs(title = "NGFR+ cells expression of other stains") +
  xlab(NULL) + ylab("Density (count) of average positive cells per ROI") +
  theme_clean()
  
densitycount_NGFR

violin_NGFR<- ggviolin(NGFR_calc_frsub, x = "Stain", y = "pos_fraction", 
                        color = "Stain", fill = "Stain", alpha=0.6, draw_quantiles = 0.5) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 4))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 4)) +
  labs(title = "NGFR+ cells expression of other stains", subtitle = "line shows quantile 0.5") +
  xlab(NULL) + ylab("Kernel probablility density of average positive cells per ROI") +
  theme_clean()
  
violin_NGFR

# Out of interest, separation on 3 different stages of lobular atrophy
boxes_NGFR_lobulestage<- ggboxplot(NGFR_calc_frsub, x = "Lobule_stage", y = "pos_fraction", color = "Lobule_stage",  alpha=0.6,
                         fill = "Lobule_stage", facet.by = "Stain") +
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 3))+
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 3))+
  labs(title = "NGFR+ cells expression of other stains, separated by 3 stages of lobular intactness") +
  xlab(NULL) + ylab("Average positive cells per ROI") +
  theme_clean()
  
boxes_NGFR_lobulestage

boxes_NGFR<- ggboxplot(NGFR_calc_frsub, x = "Stain", y = "pos_fraction", 
                        color = "Stain", fill = "Stain", alpha=0.6) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 4))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 4)) +
  labs(title = "NGFR+ cells expression of other stains") +
  xlab(NULL) + ylab("Average positive cells per ROI") +
  theme_clean()
  
boxes_NGFR
```


```{r summary of the ngfr+ co expression data file}
#number of ROIs data originated from 
NGFR_calc_frsub$Group_ImageParent <- paste(NGFR_calc_frsub$Image, NGFR_calc_frsub$Parent, sep = "_" )
allunique_ROIs <- unique(NGFR_calc_frsub$Group_ImageParent)
summary(allunique_ROIs)

NGFR_cellsdf <- NGFR_calc_frsub[, -c(3, 4, 6)]# 212 rows
NGFR_cellsdf <- unique(NGFR_cellsdf) # 53 rows (good, so corresponds to all ROIs)

sum(NGFR_cellsdf$sum_detections_ROI) # how many cells

```

```{r}
sessionInfo()
```

