---
title: "PDGFRa analysis"
author: "Sara Söderqvist"
date: "2024-02-28"
output: html_document
---

## PDGFRa analysis
Analysis of the expression profile of stroma cells in stroma-transformed/desmoplastic compartment 
We have shown that in lobules, NGFR+ stroma cells frequently co-express PDGFRa.

To show that these are separated from the PDGFRa+ cells in stroma-transformed/desmoplastic compartment ( ~iCAF proxy), we show, in particular, that these stroma cells don't express NGFR frequently.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r Data read in_cell_PDGFRa, echo = FALSE, message = FALSE, results = "hide", warning = FALSE}
setwd("/Users/sara.soderqvist/Library/CloudStorage/OneDrive-KarolinskaInstitutet/Writing_/Adaptive PDAC paper/Code/source data files to read in/m-IF stroma/PDGFRa") #Adjust the working directory to read in the file from
path <- file.path(getwd())
v.filename <- list.files(path, pattern="'*.csv", 
                         ignore.case = TRUE, 
                         full.names = TRUE)

StromaPDGFRa = do.call(cbind, lapply(v.filename, 
                                function(x) read.csv(x)))
```

```{r PDGFRa data clean up}
StromaPDGFRa$Tumor <- factor(nrow(StromaPDGFRa))

library(dplyr)
StromaPDGFRa <- StromaPDGFRa %>% mutate(Tumor = ifelse(grepl(' Tumor: ', StromaPDGFRa$Class), 'Yes', .$Tumor))
StromaPDGFRa <- StromaPDGFRa %>% mutate(Tumor = ifelse(grepl(' Not tumor: ', StromaPDGFRa$Class), 'No', .$Tumor))

#Filter our p53+ contaminations within the annotations
s_tu_StromaPDGFRa <- split(StromaPDGFRa, f = StromaPDGFRa$Tumor)
nottumor_StromaPDGFRa <- s_tu_StromaPDGFRa$No #54 256 cells

# have format to enable grouping
nottumor_StromaPDGFRa$NGFR_pos <- factor(nrow(nottumor_StromaPDGFRa))
nottumor_StromaPDGFRa$ASMA_pos <- factor(nrow(nottumor_StromaPDGFRa))
nottumor_StromaPDGFRa$CD74_pos <- factor(nrow(nottumor_StromaPDGFRa))
nottumor_StromaPDGFRa$Vim_pos <- factor(nrow(nottumor_StromaPDGFRa))
nottumor_StromaPDGFRa$PDGFR_pos <- factor(nrow(nottumor_StromaPDGFRa))

nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(NGFR_pos = ifelse(grepl(' NGFR_pos: ', nottumor_StromaPDGFRa$Class), '1', .$NGFR_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(NGFR_pos = ifelse(grepl(' NGFR_neg: ', nottumor_StromaPDGFRa$Class), '0', .$NGFR_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(ASMA_pos = ifelse(grepl(': ASMA_pos', nottumor_StromaPDGFRa$Class), '1', .$ASMA_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(ASMA_pos = ifelse(grepl(': ASMA_neg', nottumor_StromaPDGFRa$Class), '0', .$ASMA_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(CD74_pos = ifelse(grepl('CD74_pos: ', nottumor_StromaPDGFRa$Class), '1', .$CD74_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(CD74_pos = ifelse(grepl('CD74_neg: ', nottumor_StromaPDGFRa$Class), '0', .$CD74_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(Vim_pos = ifelse(grepl(' Vim_pos: ', nottumor_StromaPDGFRa$Class), '1', .$Vim_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(Vim_pos = ifelse(grepl(' Vim_neg: ', nottumor_StromaPDGFRa$Class), '0', .$Vim_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(PDGFR_pos = ifelse(grepl(' PDGFR_pos: ', nottumor_StromaPDGFRa$Class), 'Yes', .$PDGFR_pos))
nottumor_StromaPDGFRa <- nottumor_StromaPDGFRa %>% mutate(PDGFR_pos = ifelse(grepl(' PDGFR_neg: ', nottumor_StromaPDGFRa$Class), 'No', .$PDGFR_pos))

# Establish a count of how many of the PDGFRa+ cells, that also are positive for other stains. 
# First, filter our PDGFRa- cells
s_PDGFRa_StromaPDGFRa <- split(nottumor_StromaPDGFRa, f = nottumor_StromaPDGFRa$PDGFR_pos)
PDGFRa_pos <- s_PDGFRa_StromaPDGFRa$Yes #30 945 cells
```


```{r count the single positives (dataset now includes only p53-PDGFRa+ cells) }
# the goal is to show what pdgfra+ stroma cells, in desmoplasia, co-express, and the working hypothesis is that they are low for ngfr. 

# first calculating the positive fractions ROI-wise.
PDGFRa_cell <- PDGFRa_pos[, -c(3, 4, 6, 7, 12)] # removing name and class columns, and p53 and pdgra columns.

library(reshape2)
library(tidyr)
# Melt the df and tidy up
molten_stromaPDGFRa <- melt(PDGFRa_cell, id = c("Image", "Object.ID", "Parent"))
colnames(molten_stromaPDGFRa) [4] <- "Stain" #colnames such as "variable" and "value" will interfere with other terms
colnames(molten_stromaPDGFRa) [5] <- "Cell_pos"

molten_stromaPDGFRa <- separate(data = molten_stromaPDGFRa, col = "Image", into="Image", sep=c("_"), remove=TRUE)
molten_stromaPDGFRa$Group <- character(nrow(molten_stromaPDGFRa))
molten_stromaPDGFRa$Group <- paste(molten_stromaPDGFRa$Image, molten_stromaPDGFRa$Parent, molten_stromaPDGFRa$Stain, sep = "_")

s_molten_stromaPDGFRa <- split(molten_stromaPDGFRa, f = molten_stromaPDGFRa$Group)

#For now, any cell positive will be counted as positive, so now not taking to account wheather individual cells are double, triple, quadraple positive for combinations of stains.
tot_det_PDGFRa <- lapply(X = s_molten_stromaPDGFRa, FUN = function (x) {
  x$sum_detections_ROI <- nrow(x)
  x
})

fun_PDGFRafrac <- function (z) {
  nr_pos<- z %>% filter(Cell_pos == "1") # get nr positive cells for each stain
  pos_fr <- nrow(nr_pos)/nrow(z) #calculate the fraction out of all the PDGFRa+ cell in that specific ROI
  z$pos_fraction <- pos_fr # put this number in in a column
  z
} 

PDGFRa_calc_fr <- lapply (X = tot_det_PDGFRa, FUN = fun_PDGFRafrac)

PDGFRa_calc_frsub <- bind_rows(PDGFRa_calc_fr)

PDGFRa_calc_frsub <- PDGFRa_calc_frsub[, -c(2, 5)] #123 780 rows (1 row = 1 cell)
PDGFRa_calc_frsub <- unique(PDGFRa_calc_frsub) # 160 rows, means no data loss. 8 cases, 5 rois, 4 stains. (1 row = per case stain ROI)

```


```{r plot PDGFR+ co-expression }
library(ggthemes)
library(wesanderson)

boxes_PDGFRa<- ggboxplot(PDGFRa_calc_frsub, x = "Stain", y = "pos_fraction", 
                        color = "Stain", fill = "Stain", alpha=0.6) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 4))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 4)) +
  labs(title = "PDGFRa+ cells expression of other stains") +
  xlab(NULL) + ylab("Average positive cells per ROI") +
  theme_clean()
  
boxes_PDGFRa

Suppl_15c <- boxes_PDGFRa
```

```{r control nr of rois}
#number of ROIs data originated from 
PDGFRa_calc_frsub$Group_ImageParent <- paste(PDGFRa_calc_frsub$Image, PDGFRa_calc_frsub$Parent, sep = "_" )
allunique_ROIs <- unique(PDGFRa_calc_frsub$Group_ImageParent)
summary(allunique_ROIs) # 40 unique ROIs

```







