---
title: "m-IF_Stroma layering_cleaned up"
author: "Sara Söderqvist"
date: "2024-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CAF analysis  based on expression of m-IF markers.
Most plots are separated on the stromalayer. Layer here refers to the how close the stroma cell (nucleus) was to the tumor cluster tumor. 
Inner: within 15 µm to tumor
Middle: 15-30 µm from tumor
Outer: 30-45 µm from tumor.

``` {r Data cleaunup_intensity cell, echo = FALSE, message = FALSE, results = "hide", warning = FALSE}
setwd("/Users/sara.soderqvist/Library/CloudStorage/OneDrive-KarolinskaInstitutet/Writing_/Adaptive PDAC paper/Code/source data files to read in/m-IF stroma/Layers")

path <- file.path(getwd())
v.filename <- list.files(path, pattern="'*.csv", 
                         ignore.case = TRUE, 
                         full.names = TRUE)

ls_composite_intense = do.call(cbind, lapply(v.filename, 
                                function(x) read.csv(x)))
library(reshape2)
library(dplyr)

ls_composite_intense$Compartment <- factor(nrow(ls_composite_intense))
ls_composite_intense$Compartment <- "Stroma"
ls_composite_intense <- ls_composite_intense %>% mutate(Compartment = ifelse(grepl('lobe', .$Parent), 'Lobule', .$Compartment))
ls_composite_intense$Compartment <- as.factor(ls_composite_intense$Compartment)

ls_composite_intense$Layer <- factor(nrow(ls_composite_intense))
ls_composite_intense <- ls_composite_intense %>% mutate(Layer = ifelse(grepl('Inner', .$Parent), 'Inner', .$Layer))
ls_composite_intense <- ls_composite_intense %>% mutate(Layer = ifelse(grepl('Outer', .$Parent), 'Outer', .$Layer))
ls_composite_intense <- ls_composite_intense %>% mutate(Layer = ifelse(grepl('Middle', .$Parent), 'Middle', .$Layer))
ls_composite_intense$Layer <- as.factor(ls_composite_intense$Layer)

ls_composite_intense$G_ROI <- factor(nrow(ls_composite_intense))

ls_composite_intense <- ls_composite_intense[ls_composite_intense$Parent != "Image", c(colnames(ls_composite_intense))]

ls_composite_intense$G_ROI <- ls_composite_intense$Parent

ls_composite_intense$G_ROI <- gsub("Inner_stroma_", "", x = ls_composite_intense$G_ROI)
ls_composite_intense$G_ROI <- gsub("Middle_stroma_", "", x = ls_composite_intense$G_ROI)
ls_composite_intense$G_ROI <- gsub("Outer_stroma_", "", x = ls_composite_intense$G_ROI)
ls_composite_intense$G_ROI <- gsub("_lobe_", "", x = ls_composite_intense$G_ROI)
ls_composite_intense$G_ROI <- gsub("lobe_", "", x = ls_composite_intense$G_ROI)

ls_composite_intense$Group <- paste(ls_composite_intense$Image, ls_composite_intense$Compartment, ls_composite_intense$Layer, ls_composite_intense$G_ROI, sep="_")
ls_composite_intense <- ls_composite_intense[, -c(3, 6:8, 14)]


m_raw_intense <- melt(ls_composite_intense, id = c("Image", "Object.ID", "Class", "Parent", "Compartment", "Layer", "Group", "G_ROI"))

colnames(m_raw_intense) [9] <- "Stain"
colnames(m_raw_intense) [10] <- "mean_intensity"

# Only the 3 stains stains of interest:
ls_composite_intense_3 <- ls_composite_intense[, -c(6, 9)]
m_raw_intense_3 <- melt(ls_composite_intense_3, id = c("Image", "Object.ID", "Class", "Parent", "Compartment", "Layer","Group", "G_ROI"))

colnames(m_raw_intense_3) [9] <- "Stain"
colnames(m_raw_intense_3) [10] <- "mean_intensity"

```


```{r composite intensity cell initial histograms}
library(ggplot2)
library(ggthemes)
library(RColorBrewer)

# Exploratory histograms for all five stains, to show distributions. (with NGFR having so many negative or low intensity cells, difficult to see the ASMA, PDGFRa, CD74. Below there will therefore also be summaries and plots where NGFR and and Vimentin are excluded.
hist_intense_raw <- m_raw_intense %>%
  ggplot(aes(x=mean_intensity, fill=Stain, color = Stain)) +
    geom_histogram(position = 'identity', bins = 150) +
    scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_color_brewer(palette = "Dark2", direction = -1)+
  facet_grid(~ Stain) +
    theme_clean()

hist_intense_raw

hist_intense_2 <- m_raw_intense %>%
  ggplot(aes(x=mean_intensity, fill=Stain)) +
  geom_histogram(color="#e9ecef", stat = "count", bins = 150) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  facet_grid(Compartment ~ Stain) +
    theme_clean()
hist_intense_2

hist_intense_3 <- m_raw_intense %>%
  ggplot(aes(x=mean_intensity, fill=Stain)) +
  geom_histogram(color="#e9ecef", stat = "count", bins = 150) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  facet_grid(Layer ~ Stain) +
    theme_clean()
hist_intense_3

# Only for three stains
hist_intense_raw_3 <- m_raw_intense_3 %>%
  ggplot(aes(x=mean_intensity, fill=Stain)) +
    geom_histogram(color="#e9ecef", position = 'identity', bins = 150) +
    scale_fill_brewer(palette = "Dark2", direction = -1) +
  facet_grid( ~ Stain) +
    theme_clean()
hist_intense_raw_3

hist_intense_raw_3_2 <- m_raw_intense_3 %>%
  ggplot(aes(x=mean_intensity, fill=Stain)) +
    geom_histogram(color="#e9ecef", position = 'identity', bins = 150) +
    scale_fill_brewer(palette = "Dark2", direction = -1) +
  facet_grid(Compartment ~ Stain) +
    theme_clean()
hist_intense_raw_3_2

hist_intense_raw_3_3 <- m_raw_intense_3 %>%
  ggplot(aes(x=mean_intensity, fill=Stain)) +
    geom_histogram(color="#e9ecef", position = 'identity', bins = 150) +
    scale_fill_brewer(palette = "Dark2", direction = -1) +
  facet_grid(Layer ~ Stain) +
    theme_clean()
hist_intense_raw_3_3

# Adjusting the classes
m_raw_intense$Stain <- as.factor(m_raw_intense$Stain)
m_raw_intense$Class <- as.factor(m_raw_intense$Class)
m_raw_intense$mean_intensity <- as.numeric(m_raw_intense$mean_intensity)
m_raw_intense$Parent <- as.factor(m_raw_intense$Parent)
m_raw_intense$Object.ID <- as.factor(m_raw_intense$Object.ID)
m_raw_intense$Compartment <- as.factor(m_raw_intense$Compartment)
m_raw_intense$G_ROI <- as.factor(m_raw_intense$G_ROI)
m_raw_intense$Group <- as.factor(m_raw_intense$Group)
levels(m_raw_intense$Stain)
levels(m_raw_intense$Parent)
levels(m_raw_intense$Compartment)
levels(m_raw_intense$G_ROI)

m_raw_intense_3$Stain <- as.factor(m_raw_intense_3$Stain)
m_raw_intense_3$Class <- as.factor(m_raw_intense_3$Class)
m_raw_intense_3$mean_intensity <- as.numeric(m_raw_intense_3$mean_intensity)
m_raw_intense_3$Parent <- as.factor(m_raw_intense_3$Parent)
m_raw_intense_3$Object.ID <- as.factor(m_raw_intense_3$Object.ID)
levels(m_raw_intense_3$Stain)
levels(m_raw_intense_3$Parent)
```

```{r summary statistics}
library(BiocGenerics)
library(dplyr)

summary_raw_intense <- m_raw_intense %>%
  group_by(Stain) %>%
   dplyr::summarise(
     count = n(),
     mean = mean(mean_intensity, na.rm = TRUE),
     sd = sd(mean_intensity, na.rm = TRUE),
     median = median(mean_intensity, na.rm = TRUE),
     IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
   )
summary_raw_intense

# Summary by the stroma layers
summary_raw_intense_double <- m_raw_intense %>%
  group_by(Stain, Layer) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mean_intensity, na.rm = TRUE),
    sd = sd(mean_intensity, na.rm = TRUE),
    median = median(mean_intensity, na.rm = TRUE),
    IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
  )
summary_raw_intense_double

# Summary by compartment
summary_raw_intense_comp <- m_raw_intense %>%
  group_by(Stain, Compartment) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mean_intensity, na.rm = TRUE),
    sd = sd(mean_intensity, na.rm = TRUE),
    median = median(mean_intensity, na.rm = TRUE),
    IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
  )
summary_raw_intense_comp

# Summary by both compartment and layer (all vs all)

summary_raw_intense_CompLayer <- m_raw_intense %>%
  group_by(Stain, Compartment, Layer) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mean_intensity, na.rm = TRUE),
    sd = sd(mean_intensity, na.rm = TRUE),
    median = median(mean_intensity, na.rm = TRUE),
    IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
  )
summary_raw_intense_CompLayer

## With only three stains
summary_raw_intense_3 <- m_raw_intense_3 %>%
  group_by(Stain) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mean_intensity, na.rm = TRUE),
    sd = sd(mean_intensity, na.rm = TRUE),
    median = median(mean_intensity, na.rm = TRUE),
    IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
  )
summary_raw_intense_3

# Summary by the stroma layers, three stains
summary_raw_intense_layers_3 <- m_raw_intense_3 %>%
  group_by(Stain, Layer) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mean_intensity, na.rm = TRUE),
    sd = sd(mean_intensity, na.rm = TRUE),
    median = median(mean_intensity, na.rm = TRUE),
    IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
  )
summary_raw_intense_layers_3

# Summary by compartment, three stains
summary_raw_intense_comp_3 <- m_raw_intense_3 %>%
  group_by(Stain, Compartment) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mean_intensity, na.rm = TRUE),
    sd = sd(mean_intensity, na.rm = TRUE),
    median = median(mean_intensity, na.rm = TRUE),
    IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
  )
summary_raw_intense_comp_3

# Summary by both compartment and layer (all vs all), three stains
summary_raw_intense_CompLayer_3 <- m_raw_intense_3 %>%
  group_by(Stain, Compartment, Layer) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mean_intensity, na.rm = TRUE),
    sd = sd(mean_intensity, na.rm = TRUE),
    median = median(mean_intensity, na.rm = TRUE),
    IQR = IQR(mean_intensity, na.rm = TRUE, type = 7)
  )
summary_raw_intense_CompLayer_3
```

## Plots applied in the manuscript: single positivity per stain, applied from the QuPath classifier.

```{r sp dunn test with classifier, sp for each stain}
sp_fraction <- ls_composite_intense[, -c(5:9, 14:34)]

sp_fraction$NGFR_pos <- factor(nrow(sp_fraction))
sp_fraction$ASMA_pos <- factor(nrow(sp_fraction))
sp_fraction$CD74_pos <- factor(nrow(sp_fraction))
sp_fraction$Vim_pos <- factor(nrow(sp_fraction))
sp_fraction$PDGFR_pos <- factor(nrow(sp_fraction))

sp_fraction <- sp_fraction %>% mutate(NGFR_pos = ifelse(grepl(' NGFR_pos: ', sp_fraction$Class), '1', .$NGFR_pos))
sp_fraction <- sp_fraction %>% mutate(NGFR_pos = ifelse(grepl(' NGFR_neg: ', sp_fraction$Class), '0', .$NGFR_pos))
sp_fraction <- sp_fraction %>% mutate(ASMA_pos = ifelse(grepl(': ASMA_pos', sp_fraction$Class), '1', .$ASMA_pos))
sp_fraction <- sp_fraction %>% mutate(ASMA_pos = ifelse(grepl(': ASMA_neg', sp_fraction$Class), '0', .$ASMA_pos))
sp_fraction <- sp_fraction %>% mutate(CD74_pos = ifelse(grepl('CD74_pos: ', sp_fraction$Class), '1', .$CD74_pos))
sp_fraction <- sp_fraction %>% mutate(CD74_pos = ifelse(grepl('CD74_neg: ', sp_fraction$Class), '0', .$CD74_pos))
sp_fraction <- sp_fraction %>% mutate(Vim_pos = ifelse(grepl(' Vim_pos: ', sp_fraction$Class), '1', .$Vim_pos))
sp_fraction <- sp_fraction %>% mutate(Vim_pos = ifelse(grepl(' Vim_neg: ', sp_fraction$Class), '0', .$Vim_pos))
sp_fraction <- sp_fraction %>% mutate(PDGFR_pos = ifelse(grepl(' PDGFR_pos: ', sp_fraction$Class), '1', .$PDGFR_pos))
sp_fraction <- sp_fraction %>% mutate(PDGFR_pos = ifelse(grepl(' PDGFR_neg: ', sp_fraction$Class), '0', .$PDGFR_pos))

# Calculate fractions of positive cells for each stain per ROI 
melt_sp_fraction <- melt(sp_fraction, id = c("Image", "Object.ID", "Class", "Parent", "Compartment", "Layer", "G_ROI", "Group"))

colnames(melt_sp_fraction) [9] <- "Stain" 
colnames(melt_sp_fraction) [10] <- "Cell_pos"

library(tidyr)
melt_sp_fraction <- separate(data = melt_sp_fraction, col = "Image", into="Image", sep=c("_"), remove=TRUE)
melt_sp_fraction$Group <- paste(melt_sp_fraction$Group, melt_sp_fraction$Stain, sep = "")
s_melt_sp_fraction <- split(melt_sp_fraction, f = melt_sp_fraction$Group)

tot_det_sp <- lapply(X = s_melt_sp_fraction, FUN = function (x) {
  x$sum_detections_ROI <- nrow(x)
  x
})

fun_sp <- function (z) {
  nr_pos<- z %>% filter(Cell_pos == "1") # get nr positive cells for each stain
  pos_fr <- nrow(nr_pos)/nrow(z) #calculate the fraction out of all the NGFR+ cell in that specific ROI
  z$pos_fraction <- pos_fr # put it in a column
  z
} 

sp_calc_fr <- lapply (X = tot_det_sp, FUN = fun_sp) 
spcell_calc_frsub <- bind_rows(sp_calc_fr) #55 830 rows

spcell_calc_frsub <- spcell_calc_frsub[, -c(2, 3, 10, 11)] 
spcell_calc_frsub <- unique(spcell_calc_frsub) # 1815 rows

```


```{r sp stroma cells statistics and plots}
summary_comp_sp <- spcell_calc_frsub %>%
  group_by(Stain, Layer, Compartment) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(pos_fraction, na.rm = TRUE),
    sd = sd(pos_fraction, na.rm = TRUE),
    median = median(pos_fraction, na.rm = TRUE),
    IQR = IQR(pos_fraction, na.rm = TRUE)
  )
summary_comp_sp

library(rstatix)
Kruskal_PKRstroma_slcomp_sp <- spcell_calc_frsub %>%
  group_by(Stain, Compartment) %>%
  kruskal_test(pos_fraction ~ Layer) 
Kruskal_PKRstroma_slcomp_sp

Dunn_PKRstroma_comp_sp <- spcell_calc_frsub %>%
  group_by(Stain, Compartment) %>%
  dunn_test(pos_fraction ~ Layer, p.adjust.method = "BH") 
Dunn_PKRstroma_comp_sp

# comparing the stroma layer per stain and compartment.
Dunn_PKRstroma_comp_1_sp <- Dunn_PKRstroma_comp_sp %>% add_xy_position(x = "Layer")  # The positions were a bit scewed, fixed in below two lines:
Dunn_PKRstroma_comp_1_sp$xmax <- Dunn_PKRstroma_comp_1_sp$xmax-1
Dunn_PKRstroma_comp_1_sp$xmin <- Dunn_PKRstroma_comp_1_sp$xmin-1
#Dunn_PKRstroma_comp_1_sp

library(wesanderson)
library(ggpubr)
Dunn_boxes_comp_sp <- ggboxplot(spcell_calc_frsub, x = "Layer", y = "pos_fraction", 
                        color = "Layer", fill = "Layer", alpha = 0.6) +
  stat_pvalue_manual(Dunn_PKRstroma_comp_1_sp, label = "p.adj.signif")+
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 3))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 3)) +
  facet_grid(Compartment ~ Stain) +
  labs(title = "Comparison: The stroma layers per stain and compartment, cell classifier", subtitle = "Q: does the prescence of sp stroma cells in stroma and in lobes (separately) differ between inner, middle and outer layers?")
  
Dunn_boxes_comp_sp

## Not separate by compratment:
summary_sp <- spcell_calc_frsub %>%
  group_by(Stain, Layer) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(pos_fraction, na.rm = TRUE),
    sd = sd(pos_fraction, na.rm = TRUE),
    median = median(pos_fraction, na.rm = TRUE),
    IQR = IQR(pos_fraction, na.rm = TRUE)
  )
summary_sp

Kruskal_PKRstroma_sl_sp <- spcell_calc_frsub %>%
  group_by(Stain) %>%
  kruskal_test(pos_fraction ~ Layer)
Kruskal_PKRstroma_sl_sp

Dunn_PKRstroma_sp <- spcell_calc_frsub %>%
  group_by(Stain) %>%
  dunn_test(pos_fraction ~ Layer, p.adjust.method = "BH")
Dunn_PKRstroma_sp

# comparing the stroma layer per stain and compartment
Dunn_PKRstroma_1_sp <- Dunn_PKRstroma_sp %>% add_xy_position(x = "Layer")
Dunn_PKRstroma_1_sp$xmax <- Dunn_PKRstroma_1_sp$xmax-1
Dunn_PKRstroma_1_sp$xmin <- Dunn_PKRstroma_1_sp$xmin-1
#Dunn_PKRstroma_comp_1_sp

Dunn_boxes_sp <- ggboxplot(spcell_calc_frsub, x = "Layer", y = "pos_fraction", 
                        color = "Layer", fill = "Layer", alpha = 0.6) +
  stat_pvalue_manual(Dunn_PKRstroma_1_sp, label = "p.adj.signif")+
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 3))+
  scale_color_manual(values = wes_palette("GrandBudapest1", n = 3)) +
  facet_grid(. ~ Stain) +
  labs(title = "Comparison: The stroma layers per stain, cell classifier", subtitle = "Q: does the prescence of sp stroma cells in both stroma and in lobes differ between inner, middle and outer layers?")
  
Dunn_boxes_sp

Suppl_Fig_14b <- Dunn_boxes_comp_sp

Fig_4f <- Dunn_boxes_sp

```
added after biorxiv submission. 2024 03 15

```{r lobule vs stroma expression comparison}
# Question: is there a diff between lobular and stromal for each respective Stain and stroma layer? 

spcell_calc_frsub$Group_StainLayer <- paste(spcell_calc_frsub$Stain, spcell_calc_frsub$Layer, sep = "_") 
spcell_calc_frsub$Group_LayerCompartment <- paste(spcell_calc_frsub$Layer, spcell_calc_frsub$Compartment, sep = "_") 

summary_ls_msp <- spcell_calc_frsub %>%
  group_by(Stain, Layer, Compartment) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(pos_fraction, na.rm = TRUE),
    sd = sd(pos_fraction, na.rm = TRUE),
    median = median(pos_fraction, na.rm = TRUE),
    IQR = IQR(pos_fraction, na.rm = TRUE)
  )
summary_ls_msp

wilcox_PKRstroma_sllayer <- spcell_calc_frsub %>%
  group_by(Stain, Layer) %>%
  wilcox_test(pos_fraction ~ Group_LayerCompartment, paired = FALSE) %>% 
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
wilcox_PKRstroma_sllayer

# comparing the stroma layer per stain and compartment
wilcox_PKRstroma_sllayer_1 <- wilcox_PKRstroma_sllayer %>% add_xy_position(x = "Layer")

wilcox_PKRstroma_sllayer_1$xNEW <- wilcox_PKRstroma_sllayer_1$x
wilcox_PKRstroma_sllayer_1$xNEW <- wilcox_PKRstroma_sllayer_1$xNEW-1
wilcox_PKRstroma_sllayer_1$x <- wilcox_PKRstroma_sllayer_1$xNEW
wilcox_PKRstroma_sllayer_1 <- wilcox_PKRstroma_sllayer_1[, -17]
wilcox_PKRstroma_sllayer_1$xmax <- wilcox_PKRstroma_sllayer_1$xmax-1
wilcox_PKRstroma_sllayer_1$xmin <- wilcox_PKRstroma_sllayer_1$xmin-1

wilcox_PKRstroma_sllayer_1

wilcox_boxes_slcomp<- ggboxplot(spcell_calc_frsub, x = "Layer", y = "pos_fraction", 
                        color = "Compartment", fill = "Compartment", alpha = 0.6) +
  stat_pvalue_manual(wilcox_PKRstroma_sllayer_1, label = "p.adj.signif")+
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  facet_grid(. ~ Stain) +
  labs(title = "Comparison: The stroma layer per stain and compartment", subtitle = "Q: does the expression of sp [stain] in stroma and in lobes (separately) differ between inner, middle and outer stroma?")
  
wilcox_boxes_slcomp

### Make a new test, not separated between all the layers. 
summary_ls_msp_notlayer <- spcell_calc_frsub %>%
  group_by(Stain, Compartment) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(pos_fraction, na.rm = TRUE),
    sd = sd(pos_fraction, na.rm = TRUE),
    median = median(pos_fraction, na.rm = TRUE),
    IQR = IQR(pos_fraction, na.rm = TRUE)
  )
summary_ls_msp_notlayer

wilcox_PKRstroma_sl <- spcell_calc_frsub %>%
  group_by(Stain) %>%
  wilcox_test(pos_fraction ~ Compartment, paired = FALSE) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
wilcox_PKRstroma_sl

# comparing the stroma layer per stain
wilcox_PKRstroma_sl_1 <- wilcox_PKRstroma_sl %>% add_xy_position(x = "Stain")
wilcox_PKRstroma_sl_1

wilcox_boxes_slcomp_notlayer<- ggboxplot(spcell_calc_frsub, x = "Stain", y = "pos_fraction", 
                        color = "Compartment", fill = "Compartment", alpha = 0.6) +
  stat_pvalue_manual(wilcox_PKRstroma_sl_1, label = "p.adj.signif")+
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Comparison: Differential expression in the 2 compartments per stain", subtitle = "Q: does the expression of sp [stain] differ between stroma and lobes?")
  
wilcox_boxes_slcomp_notlayer

Suppl_Fig_14c <- wilcox_boxes_slcomp
Fig_4g <- wilcox_boxes_slcomp_notlayer


```


