---
title: "Adaptive PDAC"
author: "Sara Söderqvist"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Adaptive PDAC - Background

This is a study of compartment-dependent heterogeneity in pancreatic cancer. In conclusion, we can reveal that intratumor heterogeneity in many cases is related to the invasion front properties; Tumor invading stroma or tumor invading lobes - the acinar invasion.

31 cases had 6 immunohistochemistry markers quantified in QuPath, 4 related to the basal like subtype (CK/KRT17, CK/KRT5, CA125 and HMGA2) and 2 related to the classical subtype (Muc5 and CDX2, which also relates to intestinal expression). Not all stains were available for all cases, which is stated in the supplementary material.

For each case, up to 9 regions of interest (ROIs) for "tumor in septa" (=stroma) and 9 ROIs representing "tumor in acinar (=lobular) invasion". The ROIs are overlapping across the different stains. All quantifications extracted are only including tumor cells. Data cleanup include removal of ev. annotations in the images that helped correct ROI setting.

```{r read in required libraries}
library(tidyr)
library(dplyr)
library(rstatix)
library(reactable)
library(ggthemes)
library(ggpubr)
library(data.table)
library(reshape2)
library(corrplot)
library(patchwork)
library(ggplot2)
library(ComplexHeatmap)
library(viridis) 
library(RColorBrewer)
library(ggVennDiagram)
library(dichromat)
library(survminer)
library(BioVenn)
library(Hmisc)
library(stringr)
library(forcats)
library(ggridges)
library(hrbrthemes)
```

``` {r Data cleaunup, echo = FALSE, message = FALSE, results = "hide", warning = FALSE}
setwd("/Users/sara.soderqvist/Library/CloudStorage/OneDrive-KarolinskaInstitutet/Writing_/Adaptive PDAC paper/Code/source data files to read in/IHC-spatially dependent heterogeneity") # adjust the working directory

path <- file.path(getwd())
v.filename <- list.files(path, pattern="'*.csv", 
                         ignore.case = TRUE, 
                         full.names = TRUE)

all_PKR = do.call(rbind, lapply(v.filename, 
                                function(x) read.csv(x)))
all_comb <- separate(data = all_PKR, col = "Image", into=c("PKR_nr", "Stain"), sep=c("PKR-"), remove=T)
all_comb <- separate(data = all_comb, col = "Stain", into=c("PKR", "Stain"), sep=c("-"), remove=T)
all_comb <- all_comb[,-1]

all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('ck17', .$Stain), 'CK17', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('CK17', .$Stain), 'CK17', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('muc5', .$Stain), 'Muc5', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('Muc5', .$Stain), 'Muc5', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('ca125', .$Stain), 'CA125', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('CA125', .$Stain), 'CA125', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('ck5', .$Stain), 'CK5', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('CK5.', .$Stain), 'CK5', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('hmga2', .$Stain), 'HMGA2', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('HMGA2', .$Stain), 'HMGA2', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('cdx2', .$Stain), 'CDX2', .$Stain))
all_comb <- all_comb %>% mutate(Stain = ifelse(grepl('CDX2', .$Stain), 'CDX2', .$Stain))

# clean up from QuPath annotations
all_comb <- all_comb[all_comb$Class != "Lob (C)", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "tumor septa", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "tumor lobule", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Presumed lobes", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Lob normal", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Stroma", c("PKR", "Stain", "Class", "Num.Detections","Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Tumor", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Septum tumorinv", c("PKR", "Stain", "Class", "Num.Detections","Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Lobule tumorinv", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Full slide", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Debris", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "intravssc endothel lymph spread", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "NO", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Stain != "SMAD4_Nuclear", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Stain != "SMAD4_Cytoplasm", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "not tumor", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "duct probably!", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "lob", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "tumor in stroma", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "new lob", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "Not lob", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "duxt¨", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_comb <- all_comb[all_comb$Class != "septa", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive")]
all_combined <- all_comb

## Some cases are visibly negative. These cases, or individual ROIs, are imputated here with 0% positive cells, containing 300 cells per ROI.
# Fill with 0
ac <- all_combined
ac_sub49 <- rbind(ac, 
                  ac %>%
                    filter(Stain == "CK17",
                           PKR == "49",
                           Class != "Septa 9") %>% 
                    mutate(Stain = "CA125",
                           Tumor..Positive.. = 0,
                           Num.Detections = 300,
                           Num.Tumor..Negative = 300,
                           Num.Tumor..Positive = 0))

ac_sub24 <- rbind(ac_sub49,
                  ac_sub49 %>% 
                    filter(Stain == "CK17",
                           PKR == "24",
                           Class %in% c("Septa 1", "Septa 2", "Septa 3", "Septa 5", "Septa 7", "Septa 8", "Septa 9")) %>%
                    mutate(Stain = "Muc5",
                           Tumor..Positive.. = 0,
                           Num.Detections = 300,
                           Num.Tumor..Negative = 300,
                           Num.Tumor..Positive = 0))
ac_sub11 <- rbind(ac_sub24,
                  ac_sub24 %>% 
                    filter(Stain == "CK17",
                           PKR == "11",
                           !Class %in% c("Septa 6", "Lob 2", "Lob 7")) %>%
                    mutate(Stain = "CK5",
                           Tumor..Positive.. = 0,
                           Num.Detections = 300,
                           Num.Tumor..Negative = 300,
                           Num.Tumor..Positive = 0))
all_combined <- ac_sub11
all_comb_exl <- ac_sub11
all_comb_exl$Group <- paste("PKR-", all_comb_exl$PKR, " ", all_comb_exl$Stain, " ", all_comb_exl$Class, sep="")
all_comb_exl$Group_Case_Stain <- paste("PKR-", all_comb_exl$PKR, " ", all_comb_exl$Stain, sep="")
all_comb_exl$Group_Stain_Class <- paste(all_comb_exl$Stain, all_comb_exl$Class, sep=" ")
all_comb_exl$Group_Case_Class <- paste("PKR-", all_comb_exl$PKR, all_comb_exl$Class, sep = " ")
all_combined_exl <- all_comb_exl
str(all_combined_exl) 
all_combined_exl$Tissue_compartment <- character(2105)
all_combined_exl <- all_combined_exl %>% mutate(Tissue_compartment = ifelse(grepl('Lob ', all_combined_exl$Class), 'Lobule', .$Tissue_compartment))
all_combined_exl <- all_combined_exl %>% mutate(Tissue_compartment = ifelse(grepl('Septa ', all_combined_exl$Class), 'Stroma', .$Tissue_compartment))

## Make neccessary columns numeric
all_combined_exl$Num.Detections <- as.numeric(all_combined_exl$Num.Detections)
all_combined_exl$Num.Tumor..Positive <- as.numeric(all_combined_exl$Num.Tumor..Positive)
all_combined_exl$Num.Tumor..Negative<- as.numeric(all_combined_exl$Num.Tumor..Negative)

## Control: make sure there are not doublets
length(unique(all_combined_exl$Group))
str(all_combined_exl)
n_occur <- data.frame(table(all_combined_exl$Group))
n_occur[n_occur$Freq > 1,] # should become 0

ac_s <- rbind(ac_sub11, 
              ac_sub11 %>%
                filter(Stain == "CK17",
                       PKR =="1") %>% 
                mutate(Stain = "HMGA2",
                       Tumor..Positive.. = 0,
                       Num.Detections = 300,
                       Num.Tumor..Negative = 300,
                       Num.Tumor..Positive = 0))
ac_s1 <- rbind(ac_s, 
               ac_s %>%
                 filter(Stain == "CK17",
                        PKR =="13") %>% 
                 mutate(Stain = "CA125",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s2 <- rbind(ac_s1, 
               ac_s1 %>%
                 filter(Stain == "CK17",
                        PKR =="32") %>% 
                 mutate(Stain = "CA125",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s3 <- rbind(ac_s2, 
               ac_s2 %>%
                 filter(Stain == "CK17",
                        PKR =="32") %>% 
                 mutate(Stain = "HMGA2",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s4 <- rbind(ac_s3, 
               ac_s3 %>%
                 filter(Stain == "CK17",
                        PKR =="4") %>% 
                 mutate(Stain = "CA125",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s5 <- rbind(ac_s4, 
               ac_s4 %>%
                 filter(Stain == "CK17",
                        PKR =="12") %>% 
                 mutate(Stain = "CA125",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s6 <- rbind(ac_s5, 
               ac_s5 %>%
                 filter(Stain == "CK17",
                        PKR =="15") %>% 
                 mutate(Stain = "HMGA2",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s7 <- rbind(ac_s6, 
               ac_s6 %>%
                 filter(Stain == "CK17",
                        PKR =="15") %>% 
                 mutate(Stain = "CK5",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s8 <- rbind(ac_s7, 
               ac_s7 %>%
                 filter(Stain == "CK17",
                        PKR =="29") %>% 
                 mutate(Stain = "HMGA2",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s9 <- rbind(ac_s8, 
               ac_s8 %>%
                 filter(Stain == "CK17",
                        PKR =="17") %>% 
                 mutate(Stain = "HMGA2",
                        Tumor..Positive.. = 0,
                        Num.Detections = 300,
                        Num.Tumor..Negative = 300,
                        Num.Tumor..Positive = 0))
ac_s10 <- rbind(ac_s9, 
                ac_s9 %>%
                  filter(Stain == "CK17",
                         PKR =="17") %>% 
                  mutate(Stain = "CA125",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s11 <- rbind(ac_s10, 
                ac_s10 %>%
                  filter(Stain == "CK17",
                         PKR =="18") %>% 
                  mutate(Stain = "CA125",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s12 <- rbind(ac_s11, 
                ac_s11 %>%
                  filter(Stain == "CK17",
                         PKR =="49") %>% 
                  mutate(Stain = "HMGA2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s13 <- rbind(ac_s12, 
                ac_s12 %>%
                  filter(Stain == "CK17",
                         PKR =="50") %>% 
                  mutate(Stain = "CDX2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s14 <- rbind(ac_s13, 
                ac_s13 %>%
                  filter(Stain == "CK17",
                         PKR =="52a") %>% 
                  mutate(Stain = "CDX2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s15 <- rbind(ac_s14, 
                ac_s14 %>%
                  filter(Stain == "CK17",
                         PKR =="55") %>% 
                  mutate(Stain = "HMGA2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s16 <- rbind(ac_s15, 
                ac_s15 %>%
                  filter(Stain == "CK17",
                         PKR =="55") %>% 
                  mutate(Stain = "CDX2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s17 <- rbind(ac_s16, 
                ac_s16 %>%
                  filter(Stain == "CK17",
                         PKR =="47") %>% 
                  mutate(Stain = "HMGA2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s18 <- rbind(ac_s17, 
                ac_s17 %>%
                  filter(Stain == "CK17",
                         PKR =="57") %>% 
                  mutate(Stain = "HMGA2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s19 <- rbind(ac_s18, 
                ac_s18 %>%
                  filter(Stain == "CK17",
                         PKR =="57") %>% 
                  mutate(Stain = "CA125",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s20 <- rbind(ac_s19, 
                ac_s19 %>%
                  filter(Stain == "CK17",
                         PKR =="58") %>% 
                  mutate(Stain = "CDX2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s21 <- rbind(ac_s20, 
                ac_s20 %>%
                  filter(Stain == "CK17",
                         PKR =="62") %>% 
                  mutate(Stain = "CDX2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s22 <- rbind(ac_s21, 
                ac_s21 %>%
                  filter(Stain == "CK17",
                         PKR =="62") %>% 
                  mutate(Stain = "CK5",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s23 <- rbind(ac_s22, 
                ac_s22 %>%
                  filter(Stain == "CK17",
                         PKR =="63") %>% 
                  mutate(Stain = "CA125",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s24 <- rbind(ac_s23, 
                ac_s23 %>%
                  filter(Stain == "CK17",
                         PKR =="66") %>% 
                  mutate(Stain = "HMGA2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s25 <- rbind(ac_s24, 
                ac_s24 %>%
                  filter(Stain == "CK17",
                         PKR =="66") %>% 
                  mutate(Stain = "CK5",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))
ac_s26 <- rbind(ac_s25, 
                ac_s25 %>%
                  filter(Stain == "CK17",
                         PKR =="60") %>% 
                  mutate(Stain = "CDX2",
                         Tumor..Positive.. = 0,
                         Num.Detections = 300,
                         Num.Tumor..Negative = 300,
                         Num.Tumor..Positive = 0))

all_comb_imputated <- ac_s26
str(all_comb_imputated)

# Make grouping columns of combinations of case ID, marker/stain, and compartment 
all_comb_imputated$Group <- paste("PKR-", all_comb_imputated$PKR, " ", all_comb_imputated$Stain, " ", all_comb_imputated$Class, sep="")
all_comb_imputated$Group_Case_Stain <- paste("PKR-", all_comb_imputated$PKR, " ", all_comb_imputated$Stain, sep="")
all_comb_imputated$Group_Stain_Class <- paste(all_comb_imputated$Stain, all_comb_imputated$Class, sep=" ")
all_comb_imputated$Group_Case_Class <- paste("PKR-", all_comb_imputated$PKR, all_comb_imputated$Class, sep = " ")
all_combined <- all_comb_imputated
all_combined$Tissue_compartment <- character(nrow(all_combined))
all_combined <- all_combined %>% mutate(Tissue_compartment = ifelse(grepl('Lob ', all_combined$Class), 'Lobule', .$Tissue_compartment)) #terminology more accurate
all_combined <- all_combined %>% mutate(Tissue_compartment = ifelse(grepl('Septa ', all_combined$Class), 'Stroma', .$Tissue_compartment)) # terminology more accurate
all_combined$Num.Detections <- as.numeric(all_combined$Num.Detections)
all_combined$Num.Tumor..Positive <- as.numeric(all_combined$Num.Tumor..Positive)
all_combined$Num.Tumor..Negative<- as.numeric(all_combined$Num.Tumor..Negative)
length(unique(all_combined$Group))
str(all_combined)

#Control that there are no doublets
n_occur <- data.frame(table(all_combined$Group))
n_occur[n_occur$Freq > 1,] 
```
  
### The data and groupings:

```{r Wilcoxon test per stain, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
all_CK5 <- all_combined[all_combined$Stain == "CK5", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
all_CK17 <- all_combined[all_combined$Stain == "CK17", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
all_CA125 <- all_combined[all_combined$Stain == "CA125", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
all_Muc5 <- all_combined[all_combined$Stain == "Muc5", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
all_HMGA2 <- all_combined[all_combined$Stain == "HMGA2", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
all_CDX2 <- all_combined[all_combined$Stain == "CDX2", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]

## Statistical testing, with wilcox.  ##
# Adding multiple testing correction, with benjamini-hochberg.
# It is not possible to test the cases and stains that have 0:s in *all* ROIs for one or two Tissue_compartment. Therefore, the statistical testing will be on the 
# dataset before the "filling of 0% positive". 
# Called "all_combined_exl". 

exl_CK5 <- all_combined_exl[all_combined_exl$Stain == "CK5", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
exl_CK17 <- all_combined_exl[all_combined_exl$Stain == "CK17", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
exl_CA125 <- all_combined_exl[all_combined_exl$Stain == "CA125", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
exl_Muc5 <- all_combined_exl[all_combined_exl$Stain == "Muc5", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
exl_HMGA2 <- all_combined_exl[all_combined_exl$Stain == "HMGA2", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]
exl_CDX2 <- all_combined_exl[all_combined_exl$Stain == "CDX2", c("PKR", "Stain", "Class", "Num.Detections", "Tumor..Positive..", "Num.Tumor..Negative", "Num.Tumor..Positive", "Group", "Group_Case_Stain", "Group_Stain_Class", "Group_Case_Class", "Tissue_compartment")]

stat.test_CK17 <- exl_CK17 %>%
  group_by(PKR) %>%
  wilcox_test(
    Tumor..Positive.. ~ Tissue_compartment, paired = F) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
stat.test_CK17

stat.test_Muc5 <- exl_Muc5 %>%
  group_by(PKR) %>%
  wilcox_test(
    Tumor..Positive.. ~ Tissue_compartment, paired = F) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
stat.test_Muc5

stat.test_HMGA2 <- exl_HMGA2 %>%
  group_by(PKR) %>%
  wilcox_test(
    Tumor..Positive.. ~ Tissue_compartment, paired = F) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
stat.test_HMGA2

stat.test_CA125 <- exl_CA125 %>%
  group_by(PKR) %>%
  wilcox_test(
    Tumor..Positive.. ~ Tissue_compartment, paired = F) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
stat.test_CA125

stat.test_CK5 <- exl_CK5 %>%
  group_by(PKR) %>%
  wilcox_test(
    Tumor..Positive.. ~ Tissue_compartment, paired = F) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
stat.test_CK5

stat.test_CDX2 <- exl_CDX2 %>%
  group_by(PKR) %>%
  wilcox_test(
    Tumor..Positive.. ~ Tissue_compartment, paired = F) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj") 
stat.test_CDX2

## Add median in each table of statistical testing ##
# CK17
s17 <- split(x = all_CK17, f = all_CK17$PKR)
med_CK17 <-  lapply(s17, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  med_Stroma <- median(y)
  med_Lobule <- median(v)
  tibble(Median_CK17_Stroma = med_Stroma, Median_CK17_Lobule = med_Lobule)
})

tb_CK17 <- bind_rows(med_CK17, .id = "PKR")
tb_CK17 <- tibble (tb_CK17)
tail(tb_CK17)

stat.test.CK17_m <- full_join(stat.test_CK17, tb_CK17, by = "PKR")

# HMGA2
sHMGA2 <- split(x = all_HMGA2, f = all_HMGA2$PKR)
med_HMGA2 <-  lapply(sHMGA2, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  med_Stroma <- median(y)
  med_Lobule <- median(v)
  tibble(Median_HMGA2_Stroma = med_Stroma, Median_HMGA2_Lobule = med_Lobule)
})

tb_HMGA2 <- bind_rows(med_HMGA2, .id = "PKR")
tb_HMGA2 <- tibble (tb_HMGA2)
tail(tb_HMGA2)

stat.test.HMGA2_m <- full_join(stat.test_HMGA2, tb_HMGA2, by = "PKR")

# CA125
sCA125 <- split(x = all_CA125, f = all_CA125$PKR)
med_CA125 <-  lapply(sCA125, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  med_Stroma <- median(y)
  med_Lobule <- median(v)
  tibble(Median_CA125_Stroma = med_Stroma, Median_CA125_Lobule = med_Lobule)
})

tb_CA125 <- bind_rows(med_CA125, .id = "PKR")
tb_CA125 <- tibble (tb_CA125)
tail(tb_CA125)

stat.test.CA125_m <- full_join(stat.test_CA125, tb_CA125, by = "PKR")

# CK5
sCK5 <- split(x = all_CK5, f = all_CK5$PKR)
med_CK5 <-  lapply(sCK5, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  med_Stroma <- median(y)
  med_Lobule <- median(v)
  tibble(Median_CK5_Stroma = med_Stroma, Median_CK5_Lobule = med_Lobule)
})

tb_CK5 <- bind_rows(med_CK5, .id = "PKR")
tb_CK5 <- tibble (tb_CK5)
tail(tb_CK5)

stat.test.CK5_m <- full_join(stat.test_CK5, tb_CK5, by = "PKR")

# Muc5
sMuc5 <- split(x = all_Muc5, f = all_Muc5$PKR)
med_Muc5 <-  lapply(sMuc5, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  med_Stroma <- median(y)
  med_Lobule <- median(v)
  tibble(Median_Muc5_Stroma = med_Stroma, Median_Muc5_Lobule = med_Lobule)
})

tb_Muc5 <- bind_rows(med_Muc5, .id = "PKR")
tb_Muc5 <- tibble (tb_Muc5)
tail(tb_Muc5)

stat.test.Muc5_m <- full_join(stat.test_Muc5, tb_Muc5, by = "PKR")

# CDX2
sCDX2 <- split(x = all_CDX2, f = all_CDX2$PKR)
med_CDX2 <-  lapply(sCDX2, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  med_Stroma <- median(y)
  med_Lobule <- median(v)
  tibble(Median_CDX2_Stroma = med_Stroma, Median_CDX2_Lobule = med_Lobule)
})

tb_CDX2 <- bind_rows(med_CDX2, .id = "PKR")
tb_CDX2 <- tibble (tb_CDX2)
tail(tb_CDX2)

stat.test.CDX2_m <- full_join(stat.test_CDX2, tb_CDX2, by = "PKR")


## Add means in each table of statistical testing ##
# CK17
s17 <- split(x = all_CK17, f = all_CK17$PKR)
mean_CK17 <-  lapply(s17, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  mean_Stroma <- mean(y)
  mean_Lobule <- mean(v)
  tibble(Mean_CK17_Stroma = mean_Stroma, Mean_CK17_Lobule = mean_Lobule)
})

tb_CK17 <- bind_rows(mean_CK17, .id = "PKR")
tb_CK17 <- tibble (tb_CK17)
tail(tb_CK17)

stat.test.CK17_m <- full_join(stat.test.CK17_m, tb_CK17, by = "PKR") #make sure to also include the medians, with the "_m" dataset.

# HMGA2
sHMGA2 <- split(x = all_HMGA2, f = all_HMGA2$PKR)
mean_HMGA2 <-  lapply(sHMGA2, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  mean_Stroma <- mean(y)
  mean_Lobule <- mean(v)
  tibble(Mean_HMGA2_Stroma = mean_Stroma, Mean_HMGA2_Lobule = mean_Lobule)
})

tb_HMGA2 <- bind_rows(mean_HMGA2, .id = "PKR")
tb_HMGA2 <- tibble (tb_HMGA2)
tail(tb_HMGA2)

stat.test.HMGA2_m <- full_join(stat.test.HMGA2_m, tb_HMGA2, by = "PKR")

# CA125
sCA125 <- split(x = all_CA125, f = all_CA125$PKR)
mean_CA125 <-  lapply(sCA125, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  mean_Stroma <- mean(y)
  mean_Lobule <- mean(v)
  tibble(Mean_CA125_Stroma = mean_Stroma, Mean_CA125_Lobule = mean_Lobule) 
})

tb_CA125 <- bind_rows(mean_CA125, .id = "PKR")
tb_CA125 <- tibble (tb_CA125)
tail(tb_CA125)

stat.test.CA125_m <- full_join(stat.test.CA125_m, tb_CA125, by = "PKR")

# CK5
sCK5 <- split(x = all_CK5, f = all_CK5$PKR)
mean_CK5 <-  lapply(sCK5, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  mean_Stroma <- mean(y)
  mean_Lobule <- mean(v)
  tibble(Mean_CK5_Stroma = mean_Stroma, Mean_CK5_Lobule = mean_Lobule)
})

tb_CK5 <- bind_rows(mean_CK5, .id = "PKR")
tb_CK5 <- tibble (tb_CK5)
tail(tb_CK5)

stat.test.CK5_m <- full_join(stat.test.CK5_m, tb_CK5, by = "PKR")

# Muc5
sMuc5 <- split(x = all_Muc5, f = all_Muc5$PKR)
mean_Muc5 <-  lapply(sMuc5, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  mean_Stroma <- mean(y)
  mean_Lobule <- mean(v)
  tibble(Mean_Muc5_Stroma = mean_Stroma, Mean_Muc5_Lobule = mean_Lobule)
})

tb_Muc5 <- bind_rows(mean_Muc5, .id = "PKR")
tb_Muc5 <- tibble (tb_Muc5)
tail(tb_Muc5)

stat.test.Muc5_m <- full_join(stat.test.Muc5_m, tb_Muc5, by = "PKR")

# CDX2
sCDX2 <- split(x = all_CDX2, f = all_CDX2$PKR)
mean_CDX2 <-  lapply(sCDX2, function (x) {
  z <- subset(x, Tissue_compartment == "Stroma")
  u <- subset(x, Tissue_compartment == "Lobule")
  y <- z$Tumor..Positive..
  v <- u$Tumor..Positive..
  mean_Stroma <- mean(y)
  mean_Lobule <- mean(v)
  tibble(Mean_CDX2_Stroma = mean_Stroma, Mean_CDX2_Lobule = mean_Lobule)
})

tb_CDX2 <- bind_rows(mean_CDX2, .id = "PKR")
tb_CDX2 <- tibble (tb_CDX2)
tail(tb_CDX2)

stat.test.CDX2_m <- full_join(stat.test.CDX2_m, tb_CDX2, by = "PKR")

```

## Tables of Wilcoxon-tests with median and mean, separated on each stain

```{r Write wilcoxon tables}
reactable(stat.test.CK17_m, striped = T, bordered = T, defaultPageSize = 16, showPageInfo = FALSE, showPageSizeOptions = TRUE, filterable = T)
reactable(stat.test.Muc5_m, striped = T, bordered = T, defaultPageSize = 16, showPageInfo = FALSE, showPageSizeOptions = TRUE, filterable = T)
reactable(stat.test.CA125_m, striped = T, bordered = T, defaultPageSize = 16, showPageInfo = FALSE, showPageSizeOptions = TRUE, filterable = T)
reactable(stat.test.CK5_m, striped = T, bordered = T, defaultPageSize = 16, showPageInfo = FALSE, showPageSizeOptions = TRUE, filterable = T)
reactable(stat.test.HMGA2_m, striped = T, bordered = T, defaultPageSize = 16, showPageInfo = FALSE, showPageSizeOptions = TRUE, filterable = T)
```


```{r Write wilcoxon tables 2}
reactable(stat.test.CDX2_m, striped = T, bordered = T, defaultPageSize = 16, showPageInfo = FALSE, showPageSizeOptions = TRUE, filterable = T)
reactable(all_combined, striped = T, bordered = T, defaultPageSize = 16, showPageInfo = FALSE, showPageSizeOptions = TRUE, filterable = T)

```

## Boxplots
Separate for each stain, faceted on each case.  
Each dot = 1 ROI.

``` {r Boxplot CK17}
boxes_ck17 <- ggboxplot(all_CK17, x = "Tissue_compartment", y = "Tumor..Positive..",
                        color = "Tissue_compartment", palette = "Set2",
                        add = "jitter",
                        facet.by = "PKR", ncol = 4, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CK17+ cells in Lobule and Stroma", subtitle = "All cases") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 12))+
  xlab(NULL) + ylab("% CK17+ tumor cells")  +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm"))

stat.test_CK17_1 <- stat.test_CK17 %>% add_xy_position(x = "Tissue_compartment")
Suppl_Fig_4b <- boxes_ck17 + 
  stat_pvalue_manual(label = "p.adj.signif",
                                stat.test_CK17_1, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

Suppl_Fig_4b
```

``` {r Boxplot CK5}
boxes_ck5 <- ggboxplot(all_CK5, x = "Tissue_compartment", y = "Tumor..Positive..",
                        color = "Tissue_compartment", palette = "Set2",
                        add = "jitter",
                        facet.by = "PKR", ncol = 4, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CK5+ cells in Lobule and Stroma", subtitle = "All cases") +
  xlab(NULL) + ylab("% CK5+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 12)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm"))

stat.test_CK5_1 <- stat.test_CK5 %>% add_xy_position(x = "Tissue_compartment")
Suppl_5b <- boxes_ck5 +
  stat_pvalue_manual(label = "p.adj.signif",
                                stat.test_CK5_1, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

Suppl_5b
```

``` {r Boxplot CA125}
boxes_ca125 <- ggboxplot(all_CA125, x = "Tissue_compartment", y = "Tumor..Positive..",
                        color = "Tissue_compartment",
                        add = "jitter",
                        facet.by = "PKR", ncol = 4, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CA125+ cells in Lobule and Stroma", subtitle = "All cases") +
  xlab(NULL) + ylab("% CA125+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 12)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm")) +
  scale_color_manual(values = c("#FC8D62", "#66C2A5"))

stat.test_CA125_1 <- stat.test_CA125 %>% add_xy_position(x = "Tissue_compartment")

Suppl_Fig_7b <- boxes_ca125 + stat_pvalue_manual(label = "p.adj.signif",
                                stat.test_CA125_1, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

Suppl_Fig_7b

```

``` {r Boxplot HMGA2}
boxes_HMGA2 <- ggboxplot(all_HMGA2, x = "Tissue_compartment", y = "Tumor..Positive..",
                         color = "Tissue_compartment",
                         add = "jitter",
                         facet.by = "PKR", ncol = 3, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC HMGA2+ cells in Lobule and Stroma", subtitle = "All cases") +
  xlab(NULL) + ylab("% HMGA2+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 12)) +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=12)) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm"))+
  scale_color_manual(values = c("#FC8D62", "#66C2A5"))

stat.test_HMGA2_1 <- stat.test_HMGA2 %>% add_xy_position(x = "Tissue_compartment")

Suppl_Fig_6b <- boxes_HMGA2 + stat_pvalue_manual(label = "p.adj.signif",
                                 stat.test_HMGA2_1, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

Suppl_Fig_6b

```

``` {r Boxplot Muc5}
boxes_Muc5 <- ggboxplot(all_Muc5, x = "Tissue_compartment", y = "Tumor..Positive..",
                       color = "Tissue_compartment",
                       add = "jitter",
                       facet.by = "PKR", ncol = 4, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC Muc5+ cells in Lobule and Stroma", subtitle = "All cases") +
  xlab(NULL) + ylab("% Muc5+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 12)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm"))+
  scale_color_manual(values = c("#FC8D62", "#66C2A5"))

stat.test_Muc5_1 <- stat.test_Muc5 %>% add_xy_position(x = "Tissue_compartment")

Suppl_Fig_8b <- boxes_Muc5 + stat_pvalue_manual(label = "p.adj.signif",
                               stat.test_Muc5_1, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
Suppl_Fig_8b

```

``` {r Boxplot CDX2}
boxes_CDX2 <- ggboxplot(all_CDX2, x = "Tissue_compartment", y = "Tumor..Positive..",
                        color = "Tissue_compartment", palette = "Set2",
                        add = "jitter",
                        facet.by = "PKR", ncol = 4, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CDX2+ cells in Lobule and Stroma", subtitle = "All cases") +
  xlab(NULL) + ylab("% CDX2+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 12)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm"))

stat.test_CDX2_1 <- stat.test_CDX2 %>% add_xy_position(x = "Tissue_compartment")

Suppl_Fig_9b <- boxes_CDX2 + stat_pvalue_manual(label = "p.adj.signif",
                                stat.test_CDX2_1, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
Suppl_Fig_9b

```

``` {r sign dataset for boxplots, echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
ck17_s <- stat.test_CK17 %>% filter(p.adj < "0.05")
ck5_s <- stat.test_CK5 %>% filter(p.adj < "0.05")
ca125_s <- stat.test_CA125 %>% filter(p.adj < "0.05")
hmga2_s <- stat.test_HMGA2 %>% filter(p.adj < "0.05")
muc5_s <- stat.test_Muc5 %>% filter(p.adj < "0.05")
cdx2_s <- stat.test_CDX2 %>% filter(p.adj < "0.05")

ck17_sign <- merge(ck17_s, all_CK17, all.x = T)
str(ck17_sign)
ck5_sign <- merge(ck5_s, all_CK5, all.x = T)
ca125_sign <- merge(ca125_s, all_CA125, all.x = T)
hmga2_sign <- merge(hmga2_s, all_HMGA2, all.x = T)
muc5_sign <- merge(muc5_s, all_Muc5, all.x = T)
cdx2_sign <- merge(cdx2_s, all_CDX2, all.x = T)
```

## Boxplots
Only cases significant difference between lobular - and stromal invasion.  
Each dot = 1 ROI.

```{r Boxplot sign CK17, echo = FALSE, message = FALSE, warning = TRUE, fig.height=12, fig.width = 10}
stat.test_CK17_2 <- ck17_s %>% add_xy_position(x = "Tissue_compartment")
boxes_CK17_sign <- ggboxplot(ck17_sign, x = "Tissue_compartment", y = "Tumor..Positive..",
                              color = "Tissue_compartment", palette = "jco",
                              add = "jitter",
                              facet.by = "PKR", ncol = 5, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CK17+ cells in Lobule and Stroma", subtitle = "Only cases with significant difference betweeen acinar and stromal ROIs") +
  xlab(NULL) + ylab("% CK17+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 16)) +
  theme(axis.text=element_text(size=12, face = "bold"),
        axis.title=element_text(size=12,face="bold")) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm"))

boxes_CK17_sign + stat_pvalue_manual(label = "p.adj.signif",
                                      stat.test_CK17_2, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r Boxplot sign CK5, echo = FALSE, message = FALSE, warning = TRUE, fig.height=6, fig.width = 7}
stat.test_CK5_2 <- ck5_s %>% add_xy_position(x = "Tissue_compartment")
boxes_CK5_sign <- ggboxplot(ck5_sign, x = "Tissue_compartment", y = "Tumor..Positive..",
                             color = "Tissue_compartment", palette = "jco",
                             add = "jitter",
                             facet.by = "PKR", ncol = 5, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CK5+ cells in Lobule and Stroma", subtitle = "Only cases with significant difference betweeen acinar and stromal ROIs") +
  xlab(NULL) + ylab("% CK5+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 16)) +
  theme(axis.text=element_text(size=12, face = "bold"),
        axis.title=element_text(size=12,face="bold")) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm"))

boxes_CK5_sign + stat_pvalue_manual(label = "p.adj.signif",
                                     stat.test_CK5_2, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r Boxplot sign CA125, echo = FALSE, message = FALSE, warning = TRUE, fig.height=8, fig.width = 8}
stat.test_CA125_2 <- ca125_s %>% add_xy_position(x = "Tissue_compartment")
boxes_CA125_sign <- ggboxplot(ca125_sign, x = "Tissue_compartment", y = "Tumor..Positive..",
                            color = "Tissue_compartment", palette = "jco",
                            add = "jitter",
                            facet.by = "PKR", ncol = 3, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CA125+ cells in Lobule and Stroma", subtitle = "Only cases with significant difference betweeen acinar and stromal ROIs") +
  xlab(NULL) + ylab("% CA125+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 16)) +
  theme(axis.text=element_text(size=12, face = "bold"),
        axis.title=element_text(size=12,face="bold")) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm"))

boxes_CA125_sign + stat_pvalue_manual(label = "p.adj.signif",
                                    stat.test_CA125_2, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r Boxplot sign HMGA2, echo = FALSE, message = FALSE, warning = TRUE, fig.height=5, fig.width = 8}
stat.test_HMGA2_2 <- hmga2_s %>% add_xy_position(x = "Tissue_compartment")
boxes_HMGA2_sign <- ggboxplot(hmga2_sign, x = "Tissue_compartment", y = "Tumor..Positive..",
                         color = "Tissue_compartment", palette = "jco",
                         add = "jitter",
                         facet.by = "PKR", ncol = 5, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC HMGA2+ cells in Lobule and Stroma", subtitle = "Only cases with significant difference betweeen acinar and stromal ROIs") +
  xlab(NULL) + ylab("% HMGA2+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 16)) +
  theme(axis.text=element_text(size=12, face = "bold"),
        axis.title=element_text(size=12,face="bold")) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm"))

boxes_HMGA2_sign + stat_pvalue_manual(label = "p.adj.signif",
                                 stat.test_HMGA2_2, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r Boxplot sign Muc5, echo = FALSE, message = FALSE, warning = TRUE, fig.height=13, fig.width = 10}
stat.test_Muc5_2 <- muc5_s %>% add_xy_position(x = "Tissue_compartment")
boxes_Muc5_sign <- ggboxplot(muc5_sign, x = "Tissue_compartment", y = "Tumor..Positive..",
                              color = "Tissue_compartment", palette = "jco",
                              add = "jitter",
                              facet.by = "PKR", ncol = 5, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC Muc5+ cells in Lobule and Stroma", subtitle = "Only cases with significant difference betweeen acinar and stromal ROIs") +
  xlab(NULL) + ylab("% Muc5+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 16)) +
  theme(axis.text=element_text(size=12, face = "bold"),
        axis.title=element_text(size=12,face="bold")) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm"))

boxes_Muc5_sign + stat_pvalue_manual(label = "p.adj.signif",
                                      stat.test_Muc5_2, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r Boxplot sign CDX2, echo = FALSE, message = FALSE, warning = TRUE, fig.height=6, fig.width = 10}
stat.test_CDX2_2 <- cdx2_s %>% add_xy_position(x = "Tissue_compartment")
boxes_CDX2_sign <- ggboxplot(cdx2_sign, x = "Tissue_compartment", y = "Tumor..Positive..",
                             color = "Tissue_compartment", palette = "jco",
                             add = "jitter",
                             facet.by = "PKR", ncol = 5, short.panel.labs = TRUE) +
  labs(title = "Relative PDAC CDX2+ cells in Lobule and Stroma", subtitle = "Only cases with significant difference betweeen acinar and stromal ROIs") +
  xlab(NULL) + ylab("% CDX2+ tumor cells") +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"), plot.subtitle =element_text(hjust = 0.5, size = 16)) +
  theme(axis.text=element_text(size=12, face = "bold"),
        axis.title=element_text(size=12,face="bold")) +
  ylim(-1, 120) +
  theme(plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm"))

boxes_CDX2_sign + stat_pvalue_manual(label = "p.adj.signif",
                                     stat.test_CDX2_2, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Correlation plot of the markers in IHC cohort, No imputation, enabled by using the data matrix all_combined_exl, and only using complete obervations.

```{r Cor datamatrix_updated, echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
matrix_A <- dcast(data = all_combined_exl, Group_Case_Class ~ Stain,
                  value.var = "Tumor..Positive..")
#matrix_A[is.na(matrix_A)] <- 0
rownames(matrix_A) = matrix_A$Group_Case_Class
matrix_A <- matrix_A[, -1]

cor_A <- cor(matrix_A, use="complete.obs")
testRes = cor.mtest(matrix_A, conf.level = 0.95)
```

```{r Draw cor summaries_updated}
head(cor_A)
head(testRes)
```


## Draw correlation matrix with complex heatmap package
```{r correlation complex heatmap, echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
cor_p<- testRes$p
cor_p[is.nan(cor_p)]<- 1
cor_p

# the diagonal are NA, make them 1 
cor_p[is.na(cor_p)]<- 1

col_fun<- circlize::colorRamp2(c(-1, 0, 1), c("#008837", "#f7f7f7", "#7b3294"))

cell_fun = function(j, i, x, y, w, h, fill){
  if(as.numeric(x) <= 1 - as.numeric(y) + 1e-6) {
    grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
  }
  
  if (cor_p[i, j]  < 0.01 & as.numeric(x) <= 1 - as.numeric(y) + 1e-6){
    grid.text(paste0(sprintf("%.2f", cor_A[i, j]),"**"), x, y, gp = gpar(fontsize = 10))
  } else if (cor_p[i, j]  <= 0.05 & as.numeric(x) <= 1 - as.numeric(y) + 1e-6){
    grid.text(paste0(sprintf("%.2f", cor_A[i, j]),"*"), x, y, gp = gpar(fontsize = 10))
  }
}

hp<- ComplexHeatmap::Heatmap(cor_A,
                             rect_gp = gpar(type = "none"),
                             column_dend_side = "bottom",
                             column_title = "Correlation of stains",
                             column_title_gp = gpar(fontsize = 12, fontface = "bold"),
                             name = "Spearman's rank correlation coefficient", col = col_fun,
                             cell_fun = cell_fun,
                             cluster_rows = T, cluster_columns = T,
                             row_names_side = "left")

lgd_list = list(
  Legend( labels = c("<0.01", "<0.05"), title = "pvalue",
          graphics = list(
            function(x, y, w, h) grid.text("**", x = x, y = y,
                                           gp = gpar(fill = "black")),
            function(x, y, w, h) grid.text("*", x = x, y = y,
                                           gp = gpar(fill = "black")))
  ))
```


```{r draw Complex heatmap correlation_updated}
Suppl_Fig_3b <- draw(hp, annotation_legend_list = lgd_list, ht_gap = unit(1, "cm") )
Suppl_Fig_3b

#Sanity-check:
dim(matrix_A) # 488*6 = 2928

#get nr of NA:
sum(is.na(matrix_A)) #823 
2928-823
nrow(all_combined_exl)


#(dim counts) - NA = the quantified ROIs, as in the rownumbers for all_combined.

i_dim <- 2928-823
i_rows <- as.numeric(nrow(all_combined_exl))

identical(i_dim, i_rows) #2105 

```


# Heatmaps
## Heatmap type A: Case id + ROIs across Stains. 

```{r Heatmap A and matrix B, echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
m_matrix_A <- data.matrix(matrix_A) 
t_m_matrix_A <- t(m_matrix_A)


### Heatmap type B: ROIs across Case id + Stains. ###
matrix_B <- dcast(data = all_combined, Class ~ Group_Case_Stain,
                  value.var = "Tumor..Positive..")
rownames(matrix_B) = matrix_B$Class
matrix_B <- matrix_B[, -1]
m_matrix_B <- data.matrix(matrix_B) 
t_m_matrix_B <- t(m_matrix_B)

## Making the heatmaps A and B, and adding some metadata.
# First, metadata #1 for Lobule or septa ROI identity.

m_A <- data.frame(Tissue_compartment = character(488))
rownames(m_A) <- rownames(matrix_A)
m_A <- m_A %>% mutate(Tissue_compartment = ifelse(grepl(' Lob ', rownames(m_A)), 'Lobule', .$Tissue_compartment))
m_A <- m_A %>% mutate(Tissue_compartment = ifelse(grepl(' Septa ', rownames(m_A)), 'Stroma', .$Tissue_compartment))
m_A <- m_A

# Metadata #2: if they exhibit any sign of plasticity or not.
# Defined from the wilcox testing with BH correction that also are exported, and how many markers that show a significant difference
# Extensive plasticity: 3 or more markers (one basal-like connected and one classical-connected): PKR-2 16 51 30
# Moderate plasticity: 2 markers (one basal-like connected and one classical-connected): PKR-24 13 29 46 49 60 67
# Slight plasticity: One marker with significant difference: PKR-1 12 70 11 18 4 50 65 55 62 
# Basal plasticity: PKR-14 57
# Classical plasticity: PKR-63


m_A$Plasticity <- character(488)
m_A$Plasticity <- "None"
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 2 ', rownames(m_A)), 'Extensive', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 16', rownames(m_A)), 'Extensive', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 30', rownames(m_A)), 'Extensive', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 51', rownames(m_A)), 'Extensive', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 24', rownames(m_A)), 'Moderate', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 13', rownames(m_A)), 'Moderate', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 29', rownames(m_A)), 'Moderate', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 46', rownames(m_A)), 'Moderate', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 49', rownames(m_A)), 'Moderate', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 60', rownames(m_A)), 'Moderate', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 67', rownames(m_A)), 'Moderate', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 1 ', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 70', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 11', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 18', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 4', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 50', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 65', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 12', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 55', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 62', rownames(m_A)), 'Slight', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 63', rownames(m_A)), 'Classical', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 14', rownames(m_A)), 'Basal', .$Plasticity))
m_A <- m_A %>% mutate(Plasticity = ifelse(grepl('PKR- 57', rownames(m_A)), 'Basal', .$Plasticity))

annot_colors_A <- list(Plasticity=c(Extensive="#190033", Moderate="#B266FF", Slight="#E5CCFF", None="#9C9C9C", Classical = "#97C8FF", Basal = "#FF6242"), 
                     Tissue_compartment=c(Lobule = "#0066CC", Stroma = "#FF9999"))

```

```{r Pheatmap A with metadata}
pheatmap(t_m_matrix_A, fontsize_col = 6, annotation_col = m_A, annotation_colors = annot_colors_A, show_colnames = F, labels_col = "Case ID and Stain",  main = "Heatmap A: Case ID and ROIs over Stains. % Positive tumor cells", cluster_cols = T, cluster_rows = T)
```

```{r Heatmap B data frame, echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
## B:
m_B <- data.frame(Marker_type = character(159), Marker = character(159))
rownames(m_B) <- rownames(t_m_matrix_B)
#Change to stroma and Lobule name for ROIs.
colnames(t_m_matrix_B) <- gsub('Lob', 'Lobule', colnames(t_m_matrix_B))
colnames(t_m_matrix_B) <- gsub('Septa', 'Stroma', colnames(t_m_matrix_B))

m_B <- m_B %>% mutate(Marker_type = ifelse(grepl('CA125', rownames(m_B)), 'Basal_like', .$Marker_type))
m_B <- m_B %>% mutate(Marker = ifelse(grepl('CA125', rownames(m_B)), 'CA125', .$Marker))

m_B <- m_B %>% mutate(Marker_type = ifelse(grepl('CK17', rownames(m_B)), 'Basal_like', .$Marker_type))
m_B <- m_B %>% mutate(Marker = ifelse(grepl('CK17', rownames(m_B)), 'KRT17', .$Marker))

m_B <- m_B %>% mutate(Marker_type = ifelse(grepl('CK5', rownames(m_B)), 'Basal_like', .$Marker_type))
m_B <- m_B %>% mutate(Marker = ifelse(grepl('CK5', rownames(m_B)), 'KRT5', .$Marker))

m_B <- m_B %>% mutate(Marker_type = ifelse(grepl('HMGA2', rownames(m_B)), 'Basal_like', .$Marker_type))
m_B <- m_B %>% mutate(Marker = ifelse(grepl('HMGA2', rownames(m_B)), 'HMGA2', .$Marker))

m_B <- m_B %>% mutate(Marker_type = ifelse(grepl('Muc5', rownames(m_B)), 'Classical', .$Marker_type))
m_B <- m_B %>% mutate(Marker = ifelse(grepl('Muc5', rownames(m_B)), 'MUC5AC', .$Marker))

m_B <- m_B %>% mutate(Marker_type = ifelse(grepl('CDX2', rownames(m_B)), 'Classical', .$Marker_type))
m_B <- m_B %>% mutate(Marker = ifelse(grepl('CDX2', rownames(m_B)), 'CDX2', .$Marker))

col.pal <- RColorBrewer::brewer.pal(11, "RdYlBu")
col.pal2 <- plasma(9)
annot_colors_B <- list(Marker_type=c(Classical="#0066CC", Basal_like = "#FF9999"),
                     Marker = c(CDX2 = "lightsteelblue4" , MUC5AC = "#005A9C", CA125 = "brown2", KRT17 = "red4", KRT5 = "grey3", HMGA2 = "sienna4"))

```

## Heatmap type B: 
```{r Heatmap B with metadata} 
Fig_2e <- pheatmap(t_m_matrix_B, cluster_rows = T, cluster_cols = T, annotation_row = m_B, color=col.pal2, annotation_colors = annot_colors_B, show_rownames = F, cutree_cols = 2, main = "Heatmap B: Case ID and stains over ROI:s. Values: % Positive tumor cells") 

Fig_2e

#Sanity-check:
dim(t_m_matrix_B) # 159*18

#get nr of NA:
sum(is.na(t_m_matrix_B)) #354

#(dim counts) - NA = the quantified ROIs, as in the rownumbers for all_combined.

i_dim <- (159*18)-354
i_rows <- as.numeric(nrow(all_combined))

identical(i_dim, i_rows)
```

```{r Heatmap C data matrix 1,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
## C:
#One row per case.
matrix_C <- dcast(data = all_combined, PKR~ Group_Stain_Class,
                  value.var = "Tumor..Positive..")
rownames(matrix_C) = matrix_C$PKR
matrix_C <- matrix_C[, -1]
m_matrix_C <- data.matrix(matrix_C) 
t_m_matrix_C <- t(m_matrix_C)
# Change lob/septa ROI in rownames to Lobule/stroma.
rownames(t_m_matrix_C) <- gsub('Lob', 'Lobule', rownames(t_m_matrix_C))
rownames(t_m_matrix_C) <- gsub('Septa', 'Stroma', rownames(t_m_matrix_C))

## Make metadata for..
# Columns: Stain (basal or classical) & ROI (Lobule or stroma)
# Rows: plasticity 

m_CC <- data.frame(Marker_type = character(108), Tissue_compartment = character(108))
rownames(m_CC) <- colnames(m_matrix_C)

m_CC <- m_CC %>% mutate(Tissue_compartment = ifelse(grepl(' Lob ', rownames(m_CC)), 'Lobule', .$Tissue_compartment))
m_CC <- m_CC %>% mutate(Tissue_compartment = ifelse(grepl(' Septa ', rownames(m_CC)), 'Stroma', .$Tissue_compartment))

m_CC <- m_CC %>% mutate(Marker_type = ifelse(grepl('CK5', rownames(m_CC)), 'Basal_like', .$Marker_type))
m_CC <- m_CC %>% mutate(Marker_type = ifelse(grepl('CK17', rownames(m_CC)), 'Basal_like', .$Marker_type))
m_CC <- m_CC %>% mutate(Marker_type = ifelse(grepl('CA125', rownames(m_CC)), 'Basal_like', .$Marker_type))
m_CC <- m_CC %>% mutate(Marker_type = ifelse(grepl('HMGA2', rownames(m_CC)), 'Basal_like', .$Marker_type))
m_CC <- m_CC %>% mutate(Marker_type = ifelse(grepl('Muc5', rownames(m_CC)), 'Classical', .$Marker_type))
m_CC <- m_CC %>% mutate(Marker_type = ifelse(grepl('CDX2', rownames(m_CC)), 'Classical', .$Marker_type))

#I needed to add "PKR" and a blank space in the rownames, otherwise would be able to call exactly/only single PKR id:s.
str(m_matrix_C)
m_Crows <- data.frame(Plasticity = character(31))
rownames(m_Crows) <- rownames(m_matrix_C)
rownames(m_Crows) <- paste("PKR-", rownames(m_Crows), " ", sep = "")
m_Crows$Plasticity <- "None"

m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-2 ', rownames(m_Crows)), 'Extensive', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-16', rownames(m_Crows)), 'Extensive', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-30', rownames(m_Crows)), 'Extensive', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-51', rownames(m_Crows)), 'Extensive', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-24', rownames(m_Crows)), 'Moderate', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-13', rownames(m_Crows)), 'Moderate', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-29', rownames(m_Crows)), 'Moderate', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-46', rownames(m_Crows)), 'Moderate', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-49', rownames(m_Crows)), 'Moderate', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-60', rownames(m_Crows)), 'Moderate', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-67', rownames(m_Crows)), 'Moderate', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-1 ', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-70', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-11', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-18', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-4 ', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-50', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-65', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-12', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-55', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-62', rownames(m_Crows)), 'Slight', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-63', rownames(m_Crows)), 'Classical', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-14', rownames(m_Crows)), 'Basal', .$Plasticity))
m_Crows <- m_Crows %>% mutate(Plasticity = ifelse(grepl('PKR-57', rownames(m_Crows)), 'Basal', .$Plasticity))

rownames(m_Crows)<-gsub("PKR-","", as.character(rownames(m_Crows)))

annot_colors_Call <- list(Plasticity=c(Extensive="#190033", Moderate="#B266FF", Slight="#E5CCFF", None="#9C9C9C", Classical = "#97C8FF", Basal = "#FF6242"),
                        Marker_type=c(Classical="#0066CC", Basal_like = "#FF9999"),
                        Tissue_compartment=c(Lobule = "#CCE5FF", Stroma = "#FFCCCC"))
```

## Heatmap type C: 
```{r Draw heatmap type C}
pheatmap(m_matrix_C, annotation_colors = annot_colors_Call, show_colnames = F, color=col.pal2, annotation_col = m_CC, annotation_row = m_Crows, main = "Heatmap C: Case ID  over ROI:s and stains. Values: % Positive tumor cells", cluster_rows = T, cluster_cols = T)
```

```{r Heatmap C data matrix,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
## Heatmap C, but with averaged ROI:s; summarize across compartments lobular and stromal compartment.
mCA_CA125 <- matrix_C[, c(1:9)]
mCA_CA125$Avg_CA125_Lobule = rowMeans(mCA_CA125[,c(1:9)], na.rm = TRUE)
mCS_CA125 <- matrix_C[, c(10:18)]
mCS_CA125$Avg_CA125_Septa = rowMeans(mCS_CA125[,c(1:9)], na.rm = TRUE)

mCA_CDX2 <- matrix_C[, c(19:27)]
mCA_CDX2$Avg_CDX2_Lobule = rowMeans(mCA_CDX2[,c(1:9)], na.rm = TRUE)
mCS_CDX2 <- matrix_C[, c(28:36)]
mCS_CDX2$Avg_CDX2_Septa = rowMeans(mCS_CDX2[,c(1:9)], na.rm = TRUE)

mCA_CK17 <- matrix_C[, c(37:45)]
mCA_CK17$Avg_CK17_Lobule = rowMeans(mCA_CK17[,c(1:9)], na.rm = TRUE)
mCS_CK17 <- matrix_C[, c(46:54)]
mCS_CK17$Avg_CK17_Septa = rowMeans(mCS_CK17[,c(1:9)], na.rm = TRUE)

mCA_CK5 <- matrix_C[, c(55:63)]
mCA_CK5$Avg_CK5_Lobule = rowMeans(mCA_CK5[,c(1:9)], na.rm = TRUE)
mCS_CK5 <- matrix_C[, c(64:72)]
mCS_CK5$Avg_CK5_Septa = rowMeans(mCS_CK5[,c(1:9)], na.rm = TRUE)

mCA_HMGA2 <- matrix_C[, c(73:81)]
mCA_HMGA2$Avg_HMGA2_Lobule = rowMeans(mCA_HMGA2[,c(1:9)], na.rm = TRUE)
mCS_HMGA2 <- matrix_C[, c(82:90)]
mCS_HMGA2$Avg_HMGA2_Septa = rowMeans(mCS_HMGA2[,c(1:9)], na.rm = TRUE)

mCA_Muc5 <- matrix_C[, c(91:99)]
mCA_Muc5$Avg_Muc5_Lobule = rowMeans(mCA_Muc5[,c(1:9)], na.rm = TRUE)
mCS_Muc5 <- matrix_C[, c(100:108)]
mCS_Muc5$Avg_Muc5_Septa = rowMeans(mCS_Muc5[,c(1:9)], na.rm = TRUE)

# Now merge all average columns, they are matched by rownames (= PKR ID:s)
C_AVG <- cbind(mCA_CA125, mCS_CA125, mCA_CDX2, mCS_CDX2, mCA_CK17, mCS_CK17, mCA_CK5, mCS_CK5, mCA_HMGA2, mCS_HMGA2, mCA_Muc5, mCS_Muc5)
C_AVG_2 <- C_AVG[, -c(1:9, 11:19, 21:29, 31:39, 41:49, 51:59, 61:69, 71:79, 81:89, 91:99, 101:109, 111:119)] 
```

## Heatmap C, averaged compartments

```{r Heatmap C averaged }
pheatmap(C_AVG_2, cutree_rows = 2, annotation_colors = annot_colors_Call, color=col.pal2, annotation_row = m_Crows, cutree_cols = 2, main = "Heatmap C: Case ID  over ROI:s and stains. Values: Average % Positive tumor cells", cluster_rows = T, cluster_cols = T)
```

```{r Heatmap B data matrix, separate on stains,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
## Separate heatmap B on individual stains.
tmB_CK17 <- as.data.frame(t_m_matrix_B[grep("CK17", rownames(t_m_matrix_B)), ])
tmB_Muc5 <- as.data.frame(t_m_matrix_B[grep("Muc5", rownames(t_m_matrix_B)), ])
tmB_CK5 <- as.data.frame(t_m_matrix_B[grep("CK5", rownames(t_m_matrix_B)), ])
tmB_HMGA2 <- as.data.frame(t_m_matrix_B[grep("HMGA2", rownames(t_m_matrix_B)), ])
tmB_CA125 <- as.data.frame(t_m_matrix_B[grep("CA125", rownames(t_m_matrix_B)), ])
tmB_CDX2 <- as.data.frame(t_m_matrix_B[grep("CDX2", rownames(t_m_matrix_B)), ])
```

```{r Heatmap B per stain}

pheatmap(tmB_CK17, cutree_cols = 2, color=col.pal2, main = "Heatmap B for CK17: Case ID over ROI:s. Values: % Positive tumor cells")

pheatmap(tmB_Muc5, cutree_cols = 2,color=col.pal2, main = "Heatmap B for Muc5: Case ID over ROI:s. Values: % Positive tumor cells")

pheatmap(tmB_CK5, cutree_cols = 2, color=col.pal2,main = "Heatmap B for CK5: Case ID over ROI:s. Values: % Positive tumor cells")

pheatmap(tmB_HMGA2, cutree_cols = 2, color=col.pal2,main = "Heatmap B for HMGA2: Case ID over ROI:s. Values: % Positive tumor cells")

pheatmap(tmB_CA125, cutree_cols = 2, color=col.pal2,main = "Heatmap B for CA125: Case ID over ROI:s. Values: % Positive tumor cells")

pheatmap(tmB_CDX2, cutree_cols = 2,color=col.pal2, main = "Heatmap B for CDX2: Case ID over ROI:s. Values: % Positive tumor cells")
# -> only Muc5 have a complete separation
```

```{r Oa data matrix,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
##### Overall different markers' expression between acinar and stromal ROIs. #####

# Take the average of each marker in ROIs from lobules and stroma per case
# Make a paired wilcoxon, (the pair should be Lobule and stroma for the respective case) including all the stains.
# Thereby obtain which stains that differ, overall, between lobular and stromal ROIs.
# Adapting from the already created matrix C_AVG_2

oa_markdiff <- C_AVG_2
oa_markdiff$PKR <- rownames(oa_markdiff)
colnames(oa_markdiff)

a <- oa_markdiff %>% pivot_longer(cols=c('Avg_CA125_Lobule', 'Avg_CA125_Septa', 'Avg_CDX2_Lobule',  'Avg_CDX2_Septa', 'Avg_CK17_Lobule', 'Avg_CK17_Septa', 
                                         'Avg_CK5_Septa', 'Avg_CK5_Lobule', 'Avg_HMGA2_Lobule', 'Avg_HMGA2_Septa', 'Avg_Muc5_Lobule', 'Avg_Muc5_Septa'),
                                  values_to='Avg_value')
a$Tissue_compartment <- character(372)
a$Stain <- character(372)

a <- a %>% mutate(Tissue_compartment=ifelse(grepl('Lobule', a$name), 'Lobule', .$Tissue_compartment))
a <- a %>% mutate(Tissue_compartment=ifelse(grepl('Septa', a$name), 'Stroma', .$Tissue_compartment))

a <- a %>% mutate(Stain=ifelse(grepl('CK17', a$name), 'CK17', .$Stain))
a <- a %>% mutate(Stain=ifelse(grepl('CK5', a$name), 'CK5', .$Stain))
a <- a %>% mutate(Stain=ifelse(grepl('HMGA2', a$name), 'HMGA2', .$Stain))
a <- a %>% mutate(Stain=ifelse(grepl('CA125', a$name), 'CA125', .$Stain))
a <- a %>% mutate(Stain=ifelse(grepl('Muc5', a$name), 'Muc5', .$Stain))
a <- a %>% mutate(Stain=ifelse(grepl('CDX2', a$name), 'CDX2', .$Stain))
colnames(a) [5] <- 'Average_Stain'

stat.test_overall <- a %>%
  group_by(Average_Stain) %>%
  wilcox_test(
    Avg_value ~ Tissue_compartment, paired = T) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")
stat.test_overall

boxes_overall <- ggboxplot(a, x = "Tissue_compartment", y = "Avg_value",
                           color = "Tissue_compartment",
                           add = "jitter") +
  labs(title = "Average expression of PDAC cells in Lobule and Stroma", subtitle = "All cases") +
  xlab(NULL) + ylab("Average + tumor cells") +
  theme_clean() +
  ylim(-1, 120) +
  scale_color_brewer(palette= "Set2", type = "qual") +
  theme(plot.margin = margin(1.1, 1.1, 1.1, 1.1, "cm")) +
  facet_wrap(~ Average_Stain)


boxes_overall 

stat.test_overallavg <- stat.test_overall %>% add_xy_position(x = "Tissue_compartment")
```

```{r Boxes_overall 2}
boxes_overall + stat_pvalue_manual(label = "p.adj.signif",
                                   stat.test_overallavg, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # this is alright but does not show the paired (cases)

```

```{r Oa data matrix paired,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
a$Average_Stain <- as.factor(a$Average_Stain)
a$PKR <- as.factor(a$PKR)
colnames(a)[4] <- "Tissue_compartment"

## For visualizing the paired cases

Fig_2d <- ggplot(a, aes(x = Tissue_compartment, y = Avg_value)) + 
  geom_boxplot(aes(fill = Tissue_compartment, color = Tissue_compartment), alpha = 0.2) +
  geom_line(aes(group = PKR), colour = "gray", linewidth = 0.3) + 
  geom_point(size = 1.5, aes(fill = Tissue_compartment, color = Tissue_compartment)) + 
  facet_wrap(~ factor(Average_Stain, levels = c("Muc5", "CK17", "HMGA2", "CDX2", "CK5", "CA125"))) +
  stat_pvalue_manual(label = "p.adj.signif",
                     stat.test_overallavg, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme_clean() +
  scale_color_brewer(palette= "Set2", type = "qual") +
  scale_fill_brewer(palette= "Set2", type = "qual") +
  labs(title = "Average expression of PDAC cells in Lobule and Stroma", subtitle = "All cases") +
  xlab(NULL) + ylab("Average + tumor cells") +
  theme(plot.margin = margin(1.1, 1.1, 1.1, 1.1, "cm")) +
  ylim(-1, 120)

Fig_2d
```

```{r fig 2d rois check}
nlevels(a$PKR) #31, number of unique cases

a_return_perstain <- a %>%
  group_by(Average_Stain) %>%
  summarise(N.instances = n())
a_return_perstain #62, the double of 31. one average per case and compartment 

a_return_perstainandcase <- a %>%
  group_by(Average_Stain, PKR) %>%
  summarise(N.instances = n())
a_return_perstainandcase


```


```{r Oa data matrix Extensive and moderate plasticity,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
### only cases with extensive and moderate plasticity: ###

c <- a
str(c)
c$Plasticity <- factor(372)
c$Plasticity <- "None"
c$PKR<-paste("PKR-", as.character(c$PKR), ".", sep= " ")

c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 2 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 16 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 30 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 51 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 24 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 13 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 29 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 46 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 49 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 60 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 67 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 1 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 70 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 11 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 18 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 4 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 50 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 65 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 12 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 55 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 62 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 63 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 14 .', c$PKR), 'Any', .$Plasticity))
c <- c %>% mutate(Plasticity = ifelse(grepl('PKR- 57 .', c$PKR), 'Any', .$Plasticity))

c <- split(c, f = c$Plasticity)
c_plast <- c$Any
c_noplast <- c$None

stat.test_overall_cplast <- c_plast %>%
  group_by(Average_Stain) %>%
  wilcox_test(
    Avg_value ~ Tissue_compartment, paired = T) %>%
  adjust_pvalue(method = "none") %>%
  add_significance("p.adj")
stat.test_overall_cplast

boxes_overall_cplast <- ggboxplot(c_plast, x = "Tissue_compartment", y = "Avg_value",
                             color = "Tissue_compartment",
                             add = "jitter") +
  labs(title = "Average expression of PDAC cells in Lobule and Stroma", subtitle = "Subset: Cases with any degree of plasticity") +
  xlab(NULL) + ylab("Average + tumor cells") +
  theme_clean() +
  ylim(-1, 120) +
  scale_color_brewer(palette= "Accent") +
  theme(plot.margin = margin(1.1, 1.1, 1.1, 1.1, "cm")) +
  facet_wrap(~ Average_Stain)

stat.test_overallavg_cplast <- stat.test_overall_cplast %>% add_xy_position(x = "Tissue_compartment")

boxes_overall_cplast + stat_pvalue_manual(label = "p.adj.signif",
                                     stat.test_overallavg_cplast, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # does not show the paired (cases)
```

```{r Oa data matrix No plasticity,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
## No plasticity: ##
stat.test_overall_noplast <- c_noplast %>%
  group_by(Average_Stain) %>%
  wilcox_test(
    Avg_value ~ Tissue_compartment, paired = T) %>%
  adjust_pvalue(method = "none") %>%
  add_significance("p.adj")
stat.test_overall_noplast

boxes_overall_noplast <- ggboxplot(c_noplast, x = "Tissue_compartment", y = "Avg_value",
                                  color = "Tissue_compartment",
                                  add = "jitter") +
  labs(title = "Average expression of PDAC cells in Lobule and Stroma", subtitle = "Subset: Cases with no plasticity") +
  xlab(NULL) + ylab("Average + tumor cells") +
  theme_clean() +
  ylim(-1, 120) +
  scale_color_brewer(palette= "Accent") +
  theme(plot.margin = margin(1.1, 1.1, 1.1, 1.1, "cm")) +
  facet_wrap(~ Average_Stain)

stat.test_overallavg_noplast <- stat.test_overall_noplast %>% add_xy_position(x = "Tissue_compartment")

boxes_overall_noplast + stat_pvalue_manual(label = "p.adj.signif",
                                          stat.test_overallavg_noplast, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # does not show the paired (cases)
```

```{r Boxes_overall_non-plasticiy_paired}
#no plasticity, visualising paired:
boxes_overall_paired_noplast <- ggplot(c_noplast, aes(x = Tissue_compartment, y = Avg_value)) + 
  geom_boxplot(aes(fill = Tissue_compartment, color = Tissue_compartment), alpha = 0.2) +
  geom_line(aes(group = PKR), colour = "gray") + 
  geom_point(size = 1.5, aes(fill = Tissue_compartment, color = Tissue_compartment)) + 
  facet_wrap(~ Average_Stain) +
  stat_pvalue_manual(label = "p.adj.signif",
                     stat.test_overallavg_noplast, tip.length = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme_clean() +
  scale_color_brewer(palette= "Accent") +
  scale_fill_brewer(palette = "Accent") +
  labs(title = "Average expression of PDAC cells in Lobule and Stroma", subtitle = "Subset: Cases with no plasticity") +
  xlab(NULL) + ylab("Average + tumor cells") +
  theme(plot.margin = margin(1.1, 1.1, 1.1, 1.1, "cm")) +
  ylim(-1, 120)

boxes_overall_paired_noplast

```

# The simpson index

The index ranges between 0-1, where 1 = high diversity and 0 = no diversity 

Matrix with the distribution scores: 
Distribution levels; below called "Intensity scores. This category represents the fraction of positive cells in each ROI:

* 0-5%          0
* 5-20 %        1
* 20-40%        2
* 40-60%        3
* 60-80%        4
* 80-95%        5
* 95-100%       6

```{r Simpson matrix 1,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}

sim_all_combined <- all_combined[, c(1, 2, 3, 5, 12)]
#str(sim_all_combined)
sim_all_combined$Intensity_Score <- numeric(2508)

data_range_0 <- sim_all_combined$Tumor..Positive.. >= 0 & sim_all_combined$Tumor..Positive.. < 5
select_r_0 <- sim_all_combined[data_range_0, ]
select_r_0$Intensity_Score <- 0

data_range_1 <- sim_all_combined$Tumor..Positive.. >= 5 & sim_all_combined$Tumor..Positive.. < 20
select_r_1 <- sim_all_combined[data_range_1, ]
select_r_1$Intensity_Score <- 1

data_range_2 <- sim_all_combined$Tumor..Positive.. >= 20 & sim_all_combined$Tumor..Positive.. < 40
select_r_2 <- sim_all_combined[data_range_2, ]
select_r_2$Intensity_Score <- 2

data_range_3 <- sim_all_combined$Tumor..Positive.. >= 40 & sim_all_combined$Tumor..Positive.. < 60
select_r_3 <- sim_all_combined[data_range_3, ]
select_r_3$Intensity_Score <- 3

data_range_4 <- sim_all_combined$Tumor..Positive.. >= 60 & sim_all_combined$Tumor..Positive.. < 80
select_r_4 <- sim_all_combined[data_range_4, ]
select_r_4$Intensity_Score <- 4

data_range_5 <- sim_all_combined$Tumor..Positive.. >= 80 & sim_all_combined$Tumor..Positive.. < 95
select_r_5 <- sim_all_combined[data_range_5, ]
select_r_5$Intensity_Score <- 5

data_range_6 <- sim_all_combined$Tumor..Positive.. >= 95 & sim_all_combined$Tumor..Positive.. <= 100
select_r_6 <- sim_all_combined[data_range_6, ]
select_r_6$Intensity_Score <- 6

sim_all_c <- rbind(select_r_0, select_r_1, select_r_2, select_r_3, select_r_4, select_r_5, select_r_6)
#str(sim_all_c)

# Generate matrix which allows calculation of the simpson index, on each case ID and stain. 
sim_all_c$Simpson <- numeric((nrow(sim_all_c)))
sim_all_c2 <- split(sim_all_c, f = sim_all_c$PKR) #it is now a list of data frames, one for each case.

sim_all_c4 <- lapply(sim_all_c2, function (x){
  dcast(data = x, formula = Stain ~ Intensity_Score,
        value.var = "Tumor..Positive..")
})

# First, all data frames in the list need to have the stains as rownames:
sim_all_c5 <- lapply(sim_all_c4, function (x){
  rownames(x)<- x$Stain
  x <- x[, -1]
  head(x)
})

## Function to calculate explanation:
# 1) For each stain's intensity categories: calculate n(n-1) (A)
# 2) have a "SUM nr ROIs*((SUM  nr ROIs)-1), for each stain  (B)
# 3) Calculate 1-(A/B) = C = the simpson's index of diversity to use.

sim_all_c6 <- lapply(sim_all_c5, function (x){
  apply(X = x, MARGIN = 1, FUN = function(g){
    y <- g*(g-1)
    A <- sum(y, na.rm = T)
    B <- sum(g)
    BB <- B*(B-1)
    C <- 1-(A/BB)
    C
  })
  })
```

```{r Simpsons PKR-2}
sim_all_c6$`2` #Example of how the data matrix looks like, for just one case, PKR-2
```

```{r Simpson matrix 2 and remove NA,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
Simpsons_2 <- bind_rows(sim_all_c6, .id = "PKR")
simpsons_2 <- melt(Simpsons_2, id.vars = "PKR") # NA:s here are stains not available (not quantified). These should be removed.

colnames(simpsons_2)[2] <- "Stain"
colnames(simpsons_2)[3] <- "Simpsons"
simpsons_2$Group <- character(nrow(simpsons_2))
simpsons_2$Group <- paste("PKR-", simpsons_2$PKR, " ", simpsons_2$Stain, sep="")

simpsons_2 <- simpsons_2[!with(simpsons_2,is.na(Simpsons)),] #Remove NA rows, because we dont have data from these case-stains.
simpline_q <- quantile(simpsons_2$Simpsons, na.rm = T)
simpline <- simpline_q[4]

```

```{r plot Simpsons all cases and stains}
ggplot(simpsons_2, aes(x=PKR, y=Simpsons, 
                     color = Stain, fill = Stain, 
                     position = "dodge")) +
  geom_bar(stat = "identity", width=0.6, position = position_dodge2(width = 0.5, preserve = "single")) +
  scale_fill_brewer(palette = "PuOr") +
  scale_color_brewer(palette = "PuOr") + 
  geom_hline(yintercept = simpline,linetype = "longdash") +
  geom_text(aes(29.5, simpline, label = "3rd quantile", vjust = -1), col = "brown") +
  labs(title = "Simpson index of all cases and stains") +
  theme_clean() 
```

```{r Simpson matrix,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
#Make the dataset simpson_p, containing one more column where all the adjusted p-values for the wilcox tests are stored, for the respective case and stain.
simpsons_p_2 <- simpsons_2
# Spit the dataset on each stain, then match with columns pkr mutate to fill in the pvalues

simp_split_2 <- split(simpsons_p_2, f = simpsons_p_2$Stain)
colnames(stat.test_CK17)[9] <- "p_val"
stat_CK17 <- stat.test_CK17[, c(1, 9)]
simp_CK17_2 <- merge(stat_CK17, simp_split_2$CK17, all.x= T, all.y = T, by = "PKR")
colnames(stat.test_CK5)[9] <- "p_val"
stat_CK5 <- stat.test_CK5[, c(1, 9)]
simp_CK5_2 <- merge(stat_CK5, simp_split_2$CK5, all.x= T, all.y = T, by = "PKR")
colnames(stat.test_Muc5)[9] <- "p_val"
stat_Muc5 <- stat.test_Muc5[, c(1, 9)]
simp_Muc5_2 <- merge(stat_Muc5, simp_split_2$Muc5, all.x= T, all.y = T, by = "PKR")
colnames(stat.test_HMGA2)[9] <- "p_val"
stat_HMGA2 <- stat.test_HMGA2[, c(1, 9)]
simp_HMGA2_2 <- merge(stat_HMGA2, simp_split_2$HMGA2, all.x= T, all.y = T, by = "PKR")
colnames(stat.test_CDX2)[9] <- "p_val"
stat_CDX2 <- stat.test_CDX2[, c(1, 9)]
simp_CDX2_2 <- merge(stat_CDX2, simp_split_2$CDX2, all.x= T, all.y = T, by = "PKR")
colnames(stat.test_CA125)[9] <- "p_val"
stat_CA125 <- stat.test_CA125[, c(1, 9)]
simp_CA125_2 <- merge(stat_CA125, simp_split_2$CA125, all.x= T, all.y = T, by = "PKR")

simpsons_pval_2 <- rbind(simp_CA125_2, simp_CDX2_2, simp_CK17_2, simp_HMGA2_2, simp_Muc5_2, simp_CK5_2)
simpsons_pval_2 <- simpsons_pval_2[, -6]

#simpsons_pval_2 does have NA where statistical testsing could not be done(ie, for cases and stains that were "filled with 0".) Fill these instances with '0.99'.
simpsons_pval_2[is.na(simpsons_pval_2)] <- 0.99

# Making a mirror histogram:
pline <- 0.05

simpsons_lgpval_2 <- simpsons_pval_2
simpsons_lgpval_2$p_val <- log10(simpsons_lgpval_2$p_val)

```

```{r mirror  barchart lg pval, scatterplots and venn}
ggplot(simpsons_lgpval_2, aes(x=PKR, 
                            color = Stain, fill = Stain, 
                            position = "dodge")) +
  geom_bar(aes(y = p_val), stat = "identity", width=0.6, position = position_dodge2(width = 0.5, preserve = "single")) +
  geom_bar(aes(y = Simpsons), stat = "identity", width=0.6, position = position_dodge2(width = 0.5, preserve = "single")) +
  scale_fill_brewer(palette= "BrBG", type = "div") +
  scale_color_brewer(palette= "BrBG", type = "div") + 
  geom_hline(yintercept = log10(pline), linetype = "dashed", color = "black") +
  geom_text(aes(29.5, simpline, label = "3rd quantile", vjust = -1), size = 3, col = "black") +
  geom_text(aes(28.8, log10(pline), label = "p = 0.05", vjust = -1), size = 3, col = "black") +
  geom_hline(yintercept = simpline,linetype = "dashed", col = "black") +
  geom_hline(yintercept = 0,linetype = "solid", color = "black")+
  labs(title = "p-values of wilcoxon test and Simpsons for all cases and stains") +
  ylab(c("lg(adjusted p-values from wilxocon testing)               Simpson index of diversity")) +
  theme_clean()

# Scatterplot

x_line <- simpsons_pval_2$p_val
y_line <- simpsons_pval_2$Simpsons

scatter_simppval <- ggplot(simpsons_pval_2, aes(p_val, Simpsons, color = Stain), size = 4, alpha = 0.8)


Suppl_Fig_10a <- scatter_simppval + geom_point(aes(alpha = 0.7, size = 0.5)) +
  ggtitle("scatterplot of p-values and simpsons index: Suppl fig 10a") +
  geom_hline(yintercept = simpline,linetype = "dashed", col = "black") +
  geom_text(aes(0.75, simpline, label = "3rd quantile", vjust = -1), size = 4, col = "black") +
  geom_vline(xintercept = pline, linetype = "dashed", color = "black") +
  geom_text(aes(0.15, pline, label = "p = 0.05 ", vjust = -1), size = 4, col = "black") +
  scale_color_manual(values = 
                       c("CK17"="#e41a1c", "CK5"="#984ea3", "HMGA2" = "#ff7f00", "CA125" = "#ffff33", "CDX2" = "#4daf4a", "Muc5" = "#377eb8"))+
  theme_clean()

Suppl_Fig_10a

scatter_simppval + geom_point(aes(alpha = 0.7, size = 3)) +
  ggtitle("scatterplot of p-values and simpsons index") +
  geom_hline(yintercept = simpline,linetype = "dashed", col = "black") +
  geom_text(aes(0.75, simpline, label = "3rd quantile", vjust = -1), size = 4, col = "black") +
  geom_vline(xintercept = pline, linetype = "dashed", color = "black") +
  geom_text(aes(0.15, pline, label = "p = 0.05 ", vjust = -1), size = 4, col = "black") +
  scale_color_manual(values = 
                       c("CK17"="#e41a1c", "CK5"="#984ea3", "HMGA2" = "#ff7f00", "CA125" = "#ffff33", "CDX2" = "#4daf4a", "Muc5" = "#377eb8"))+
  geom_smooth(method = "lm", se = FALSE) +
  theme_clean()

simpsons_pval_3 <- simpsons_pval_2
simpsons_pval_3$Event_simp_above <- factor(186)
simpsons_pval_3$Event_pval_below <- factor(186)

simpsons_pval_3 <- simpsons_pval_3 %>% mutate(Event_simp_above = ifelse(Simpsons >= simpline, 
                                                                  'true', .$Event_simp_above))
simpsons_pval_3 <- simpsons_pval_3 %>% mutate(Event_pval_below = ifelse(p_val <= pline, 
                                                                  'true', .$Event_pval_below))

simpsons_pval_3 <- simpsons_pval_3 %>% mutate(Event_simp_above = ifelse(Event_simp_above != "true", 
                                                                  'false', .$Event_simp_above))
simpsons_pval_3 <- simpsons_pval_3 %>% mutate(Event_pval_below = ifelse(Event_pval_below != "true", 
                                                                  'false', .$Event_pval_below))
simpsons_pval_4 <- simpsons_pval_3[, c(1, 3, 5:7)]  
simpsons_pval_4[is.na(simpsons_pval_4)] <- 'false'
simpsons_venn <- simpsons_pval_4[, c(3, 4, 5)]
simpsons_venn <- simpsons_venn  %>% mutate(across(starts_with("Event"), as.logical))

Vennfunc <- lapply(simpsons_venn[, c(2, 3)], function (x)
  which(x == 1))

Venn_simp_pval <- ggVennDiagram(Vennfunc)

Suppl_Fig_10b_old <- Venn_simp_pval +
  scale_fill_gradient(low = "#0072B2", high = "#CC79A7")
  scale_colour_gradient(low = "#0072B2", high = "#CC79A7")
  theme_cleantable()
Suppl_Fig_10b_old #color changed

```


```{r venn diagrams with proportional circle size}
# Venn, version where the size of Venn circles correspond to the numbers of cells in the respective group
# colors now correspond to the group identity instead of nr of ROIs in each group.
listed_z <- NULL

propVenn <- draw.venn(list_x = Vennfunc$Event_simp_above, list_y = Vennfunc$Event_pval_below, list_z = listed_z, title = "Venn diagram of overlapping measures of heterogeneity", xtitle = "Simpsons above threshold", ytitle = "p−value sign", x_c = "#F1BB7B", y_c = "#FD6467",  nrtype = "pct")
Suppl_Fig_10b_new <- propVenn
Suppl_Fig_10b_new
```



```{r simpsons count nr of instances}

simppval_return <- simpsons_venn %>%
  group_by(Event_pval_below, Event_simp_above) %>%
  summarise(N.events = n())

simppval_return 

simpson_return_perstain <- simpsons_pval_2 %>%
  group_by(Stain) %>%
  summarise(N.instances = n())

simpson_return_perstain
sum(simpson_return_perstain$N.instances)
```


```{r Simpson 4 examples dm,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
## To examplify, create a subset of a to re-create this type of graph with one example from each heterogenety/plasticity category:
# PKR-2, 49, 70 and 17.
simpson_2pval_2 <- simpsons_pval_2[simpsons_pval_2$PKR == c("2"), ]
simpson_49pval_2 <- simpsons_pval_2[simpsons_pval_2$PKR == c("49"), ]
simpson_70pval_2 <- simpsons_pval_2[simpsons_pval_2$PKR == c("70"), ]
simpson_17pval_2 <- simpsons_pval_2[simpsons_pval_2$PKR == c("17"), ]
ex_simpson_pval_2 <- rbind(simpson_2pval_2, simpson_49pval_2, simpson_70pval_2, simpson_17pval_2)

exlg_simpson_pval_2 <- ex_simpson_pval_2
exlg_simpson_pval_2$p_val <- log10(exlg_simpson_pval_2$p_val)
```

```{r Exlg pcval Simpsons }
ggplot(exlg_simpson_pval_2, aes(x=PKR, 
                              color = Stain, fill = Stain, 
                              position = "dodge")) +
  geom_bar(aes(y = p_val), stat = "identity", width=0.6, position = position_dodge2(width = 0.5, preserve = "single")) +
  geom_bar(aes(y = Simpsons), stat = "identity", width=0.6, position = position_dodge2(width = 0.5, preserve = "single")) +
  scale_fill_brewer(palette = "PuOr") +
  scale_color_brewer(palette = "PuOr") + 
  geom_hline(yintercept = log10(pline), linetype = "dashed", color = "brown") +
  geom_hline(yintercept = simpline,linetype = "dashed", color = "brown") +
  geom_hline(yintercept = 0,linetype = "solid", color = "black")+
  geom_text(aes(4.4, log10(pline), label = "p < 0.05", vjust = 2), col = "brown") +
  geom_text(aes(4.4, simpline, label = "3rd quantile", vjust = -1), col = "brown") +
  labs(title = "p-values of wilcoxon test and Simpsons") +
  ylab(c("log(adjusted p-values from wilxocon testing)                            Simpson index of diversity")) +
  ylim(-3.8, 1.5) +
  theme_clean()
```

```{r comparing heterogeneity measures dm,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
## Making groups for high heterogenous PKRs based on simpsons. 
simpsons_above <- simpsons_2$Simpsons >= simpline
simpsons_ab <- simpsons_2[simpsons_above, ] #NA is filled in those without the match, remove those...

simpsons_pass <- na.delete(simpsons_ab)
ssimp <- split(simpsons_pass, f = simpsons_pass$PKR)

ssimp2 <- lapply(X= ssimp, function (x) {
  nrow(x)
})
ssimp3 <- bind_rows(ssimp2)
# This is now PKR in column names, and in row 1 how many stains that case had that were considered highly heterogenous (cutoff 3rd quantile) using Simpson index of diversity.
# This can be added to m_Crows, prev. used for metadata in the heatmaps.
ssimp4 <- melt(ssimp3)

rownames(ssimp4) <- ssimp4$variable
rownames(ssimp4) <- paste(rownames(ssimp4), " ", sep="")
str(ssimp4)

het <- merge(ssimp4, m_Crows,by = 'row.names', all = T)
str(het)
## Now same grouping for heterogeneity as plasticity:
# Extensive: 3 or more markers above simpson threshold
# Moderate: 2 markers above simpson threshold
# Slight plasticity: One marker above simpson threshold
het$Heterogeneity <- character(31)
het$Heterogeneity <- "None"
het<- het %>% mutate(Heterogeneity = ifelse(grepl('1', het$value), 'Slight', .$Heterogeneity))
het<- het %>% mutate(Heterogeneity = ifelse(grepl('2', het$value), 'Moderate', .$Heterogeneity))
het<- het %>% mutate(Heterogeneity = ifelse(grepl('3', het$value), 'Extensive', .$Heterogeneity))
het <- het %>% mutate(Heterogeneity = ifelse(grepl('4', het$value), 'Extensive', .$Heterogeneity))

het <- het[, -c(2, 3)]
# het has ordinal variables, which is un-plottable as barcharts.
# So these will needs to be ranked instead.
# None = 0.5 ( this is in order to see those represented and not just as empty space in the charts)
# Slight = 1
# Moderate = 2
# Extensive = 3

het$Heterogeneity_Cat <- numeric(31)
het$Heterogeneity_Cat <- 0
het <- het %>% mutate(Heterogeneity_Cat = ifelse(grepl('Slight', het$Heterogeneity), '1', .$Heterogeneity_Cat))
het<- het %>% mutate(Heterogeneity_Cat = ifelse(grepl('Moderate', het$Heterogeneity), '2', .$Heterogeneity_Cat))
het<- het %>% mutate(Heterogeneity_Cat = ifelse(grepl('Extensive', het$Heterogeneity), '3', .$Heterogeneity_Cat))

het$Plasticity_Cat <- numeric(31)
het$Plasticity_Cat <- 0
het<- het %>% mutate(Plasticity_Cat = ifelse(grepl('Slight', het$Plasticity), '1', .$Plasticity_Cat))
het<- het %>% mutate(Plasticity_Cat = ifelse(grepl('Moderate', het$Plasticity), '2', .$Plasticity_Cat))
het<- het %>% mutate(Plasticity_Cat = ifelse(grepl('Extensive', het$Plasticity), '3', .$Plasticity_Cat))

het$Heterogeneity_Cat <- as.numeric(het$Heterogeneity_Cat)
het$Plasticity_Cat <- as.numeric(het$Plasticity_Cat)
het$PKR <- as.character(het$Row.names)
row.names(het) <- het$PKR
het$PKR<-gsub("PKR-","",as.character(het$PKR))
#het <- het[, -1]
str(het)
het$Group <- character(31)
het$Group <-  "Same"
het <- het %>% mutate(Group=ifelse(Heterogeneity_Cat > Plasticity_Cat, 'Het_Simpsons higher', .$Group))
het <- het %>% mutate(Group=ifelse(Plasticity_Cat > Heterogeneity_Cat, 'Het_Wilcox higher', .$Group))
```

```{r Comparing heterogeneity measures}
ggplot(het, aes(x=PKR, position = "dodge", color = Group, fill = Group)) +
  geom_bar(aes(y = Heterogeneity_Cat), stat = "identity", width=0.6, position = position_dodge2(width = 0.5, preserve = "single")) +
  geom_bar(aes(y = -Plasticity_Cat), stat = "identity", width = 0.6, position = position_dodge2(width = 0.5, preserve = "single")) +
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") + 
  geom_hline(yintercept = 0,linetype = "solid", color = "black")+
  labs(title = "Comparing heterogeneity measures") +
  ylab(c("Plasticity by Wilcox                            Heterogeneity by Simpsons")) +
  scale_y_continuous(breaks=c(-3, -2, -1, 1, 2, 3), label= c("Extensive", "Moderate", "Slight", "Slight", "Moderate", "Extensive")) +
  theme_clean()
```


```{r Aggregated acinar and stromal expression dm,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
##### Aggregated histogram plots #####
# Visualizing if there is a any grouping of low or high heterogeneity, depending on markers expressed in either compartment.
waterr <- all_combined[, c(1, 2, 3, 4, 7, 12)]
water <- split(waterr, f = waterr$PKR)

watertoplot <- lapply(water, function (x){
  a <- x[x$Tissue_compartment %in% c('Lobule'), ]  # only retrieves lobular ROIs.
  b <- a[a$Stain %in% c('CK17', 'CK5', 'HMGA2', 'CA125'), ] #only keep the basal like related markers in "b"
  c <- sum(a$Num.Tumor..Positive) # nr of positive tumor cells in "c" (So not total numer of cell detections.)
  d <- sum(b$Num.Tumor..Positive) # Nr of basal like related positive cells
  cla <- a[a$Stain %in% c('Muc5', 'CDX2'), ]
  e <- sum(cla$Num.Tumor..Positive) #nr of classical related positive cells
  ratio_bas <- d/c
  ratio_cla <- e/c 
  plo <- data_frame(Basal_r = ratio_bas, Classical_r = ratio_cla) # For the individual case, the ratio of basal like and classical markers out of the total nr of positive cells is noted in "plo".
})

bound <- bind_rows(watertoplot, .id = "tib") 
colnames(bound)[1] <- "Case_ID"
head(bound)

wat <- melt(bound, id = "Case_ID",variable.name = "Group", value.name = "Ratio")
head(wat)

wat$Group <- factor(wat$Group, levels = c('Classical_r', 'Basal_r'))
wat <- wat %>%
  arrange(desc(Group), Ratio)
wat <- wat %>%
  mutate(Case_ID = factor(Case_ID, unique(Case_ID)))

ratio_cols <- c("#FFe5ec", "#ceeaf6")
ratio_cols_acinar <- c(Classical_r = "#ceeaf6", Basal_r = "#FFe5ec")
Ratio_Lobule <-ggplot(wat, aes(x=Case_ID, y=Ratio, fill=Group)) +
  geom_col() +
  labs(title = "Ratio of all positive cells in acinar ROIs", subtitle = "All cases") +
  theme_clean()+
  scale_fill_manual(values=ratio_cols_acinar)

Ratio_Lobule

watertoplot_stroma <- lapply(water, function (x){
  s <- x[x$Tissue_compartment %in% c('Stroma'), ]  
  b <- s[s$Stain %in% c('CK17', 'CK5', 'HMGA2', 'CA125'), ] 
  c <- sum(s$Num.Tumor..Positive)
  d <- sum(b$Num.Tumor..Positive) 
  cla <- s[s$Stain %in% c('Muc5', 'CDX2'), ]
  e <- sum(cla$Num.Tumor..Positive) 
  ratio_bas <- d/c
  ratio_cla <- e/c 
  plo <- data_frame(Basal_r = ratio_bas, Classical_r = ratio_cla)
})
head(watertoplot_stroma)

bound_stroma <- bind_rows(watertoplot_stroma, .id = "tib") 
colnames(bound_stroma)[1] <- "Case_ID"

wat_s <- melt(bound_stroma, id = "Case_ID",variable.name = "Group", value.name = "Ratio")
wat_s <- wat_s %>%
  arrange(desc(Group), Ratio)
wat_s <- wat_s %>%
  mutate(Case_ID = factor(Case_ID, unique(Case_ID)))

Ratio_stroma <-ggplot(wat_s, aes(x=Case_ID, y=Ratio, fill=Group)) +
  geom_col() +
  labs(title = "Ratio of all positive cells in stromal ROIs", subtitle = "All cases") +
  theme_clean()+
  scale_fill_manual(values=ratio_cols)

Ratio_stroma

# Now a faceted plot with both acinar and stromal ROIs. 
head(wat)
head(wat_s)
wat_joined <- bind_rows( "Stromal" = wat_s, "Acinar" = wat, .id= "ROI") #The order of "stromal" and "acinar" here determine the order that will show up in the plot "sp"

sp <- ggplot(wat_joined, aes(x = Case_ID, y = Ratio, fill = Group)) +
  geom_col(width = 0.7) +
  theme_clean()+
  scale_fill_manual(values=ratio_cols)

```


```{r Aggregted histograms on stroma and acinar}
sp + 
  facet_grid(ROI ~ .)+ #sort on (basal like) expression in stroma.
  labs(title = "Ratio of classical and basal-like expressing cells in stromal and acinar ROIs", subtitle = "All cases") 
```

```{r Aggregates_with plasticity,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}

# Visualise if there is any clustering or in other of the heterogenetiy classes
# Such as in the metadata m_Crows, a copy is made ("Plasticity_degree") 

Plasticity_degree <- m_Crows
Plasticity_degree$Case_ID <- row.names(Plasticity_degree) 

wat_m <- wat
wat_str_m <- wat_s
wat_m$Case_ID <- as.character(wat_m$Case_ID)
wat_m <- wat_m %>% 
  mutate(Case_ID = str_replace_all(Case_ID," ", ""))
Plasticity_degree <- Plasticity_degree %>% 
  mutate(Case_ID = str_replace_all(Case_ID," ", ""))
wat_meta <- merge(wat_m, Plasticity_degree, by = "Case_ID")

wat_str_m$Case_ID <- as.character(wat_str_m$Case_ID)
wat_str_m <- wat_str_m %>% 
  mutate(Case_ID = str_replace_all(Case_ID," ", ""))
wat_str_meta <- merge(wat_str_m, Plasticity_degree, by = "Case_ID")

wat_joined_meta <- bind_rows("Acinar" = wat_meta,  "Stromal" = wat_str_meta, .id= "ROI")

wat2_s_joined <- wat_joined_meta %>%
  arrange(desc(Plasticity), Ratio)
wat3_s_joined <- wat2_s_joined %>%
  mutate(Case_ID = factor(Case_ID, unique(Case_ID)))
Plast_order <- c("Extensive", "Moderate", "Slight", "None", "Classical", "Basal")
ratios_joined <- wat3_s_joined %>% 
  arrange(factor(Plasticity, levels = Plast_order))

Ratio_joined <- ggplot(ratios_joined, aes(x = Case_ID, y = Ratio, fill = Group)) +
  geom_col(width = 0.7) +
  theme_clean()+
  scale_fill_manual(values=ratio_cols_acinar)
```

```{r Aggreg. Ratio plasticity}
Ratio_joined + 
  facet_grid(ROI ~ .) +
  geom_point(aes(color=Plasticity, alpha = Group)) +
  scale_colour_manual(values = c(Extensive = "#4D004B", Moderate = "#8C6BB1", Slight = "#9EBCDA", None = "#808080", Classical = "#97C8FF", Basal = "#FF6242")) +
  scale_alpha_discrete(range=c(0, 1))+
  labs(title = "Ratio of all positive cells separated on compartment", subtitle = "All cases")  

```

```{r Ratio_no plasticity,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}
### The ratio plots, but with only no plasticity ###
ratio_noplast <- split(ratios_joined, f = ratios_joined$Plasticity)
ratio_noplast <- ratio_noplast$None
ratio_noplast_s <- ratio_noplast %>%
  arrange(desc(Group), Ratio)


# Cases with no plasticity, drawn manually from wat_joined(the sorted datafile):
wat_joined$Case_ID <- as.character(wat_joined$Case_ID)
wat_joined_noplast <- wat_joined[wat_joined$Case_ID %in% c("15", "17", "52a", "66", "32", "58", "47"), ]


sp_noplast <- ggplot(wat_joined_noplast, aes(x = Case_ID, y = Ratio, fill = Group)) +
  geom_col(width = 0.7) +
  theme_clean()+
  scale_fill_manual(values=ratio_cols) +
  labs(title = "Ratio of all positive cells separated on compartment", subtitle = "Subgroup: Only cases with no plasticity") 

```

```{r ratio_aggr no plast}
sp_noplast + 
  facet_grid(ROI ~ .) #sort on (basal like) expression in stroma.
```

# Distribution of % positive cells for all cases

```{r Distrib plot objects,  echo = FALSE, message = FALSE, results = "hide", warning = TRUE}

distrib <- all_combined[, c(1, 2, 3, 5, 11, 12)]
distrib <- split(distrib, f = distrib$Tissue_compartment)
distrib_a <- distrib$Lobule
distrib_s <- distrib$Stroma

## Basal-like in stroma ##
distrib_s_ba <- distrib_s[distrib_s$Stain %in% c('CK17', 'CK5', 'HMGA2', 'CA125'), ]

distrib_stroma_basal <- ggplot(distrib_s_ba, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of basal-like stains'expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  facet_grid(Stain ~ .) +
  labs(x = "Each case + stromal ROI combination",) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())


distrib_s_CK17 <- distrib_s_ba[distrib_s_ba$Stain %in% 'CK17', ]
length(unique(distrib_s_CK17$Group_Case_Class))
n_occur_17 <- data.frame(table(distrib_s_CK17$Group_Case_Class))
n_occur_17[n_occur_17$Freq > 1,] 

# Stain specific
distrib_stroma_CK17 <-ggplot(distrib_s_CK17, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive.., na.rm = TRUE), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CK17 expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CK17+ cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = c("red4"))

distrib_s_CK5 <- distrib_s_ba[distrib_s_ba$Stain %in% 'CK5', ]

distrib_stroma_CK5 <-ggplot(distrib_s_CK5, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CK5 expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CK5+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = c("grey3"))

distrib_s_CA125 <- distrib_s_ba[distrib_s_ba$Stain %in% 'CA125', ]

distrib_stroma_CA125 <-ggplot(distrib_s_CA125, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CA125 expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CA125+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = c("brown2"))

distrib_s_HMGA2 <- distrib_s_ba[distrib_s_ba$Stain %in% 'HMGA2', ]

distrib_stroma_HMGA2 <-ggplot(distrib_s_HMGA2, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of HMGA2 expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% HMGA2+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "sienna4")

## Classical in stromal ROIs ##
distrib_s_cl <- distrib_s[distrib_s$Stain %in% c('Muc5', 'CDX2'), ]

distrib_stroma_classical <-ggplot(distrib_s_cl, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of classical stains'expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  facet_grid(Stain ~ .) +
  labs(x = "Each case + stromal ROI combination",) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values=col.pal)

#Stain specific
distrib_s_Muc5 <- distrib_s_cl[distrib_s_cl$Stain %in% 'Muc5', ]

distrib_stroma_Muc5 <-ggplot(distrib_s_Muc5, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of Muc5 expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% Muc5+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "#005A9C")

distrib_s_CDX2 <- distrib_s_cl[distrib_s_cl$Stain %in% 'CDX2', ]

distrib_stroma_CDX2 <-ggplot(distrib_s_CDX2, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CDX2 expression in the stroma", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CDX2+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "lightsteelblue4")

## Basal like in acinar ROIs ##
distrib_a_ba <- distrib_a[distrib_a$Stain %in% c('CK17', 'CK5', 'HMGA2', 'CA125'), ]

distrib_lobule_basal <-ggplot(distrib_a_ba, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of basal-like stains' expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  facet_grid(Stain ~ .) +
  labs(x = "Each case + stromal ROI combination",) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

# Stain specific
distrib_a_CK17 <- distrib_a_ba[distrib_a_ba$Stain %in% 'CK17', ]

distrib_lobule_CK17 <-ggplot(distrib_a_CK17, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CK17 expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CK17+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "red4")

distrib_a_CK5 <- distrib_a_ba[distrib_a_ba$Stain %in% 'CK5', ]

distrib_lobule_CK5 <-ggplot(distrib_a_CK5, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CK5 expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CK5+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "grey3")

distrib_a_HMGA2 <- distrib_a_ba[distrib_a_ba$Stain %in% 'HMGA2', ]

distrib_lobule_HMGA2 <-ggplot(distrib_a_HMGA2, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of HMGA2 expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% HMGA2+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "sienna4")

distrib_a_CA125 <- distrib_a_ba[distrib_a_ba$Stain %in% 'CA125', ]

distrib_lobule_CA125 <-ggplot(distrib_a_CA125, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CA125 expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CA125+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "brown2")

## Classical in acinar ROIs ##
distrib_a_cl <- distrib_a[distrib_a$Stain %in% c('Muc5', 'CDX2'), ]

distrib_lobule_classical <-ggplot(distrib_a_cl, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of classical stains' expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  facet_grid(Stain ~ .) +
  labs(x = "Each case + stromal ROI combination",) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

# Stain specific
distrib_a_Muc5 <- distrib_a_cl[distrib_a_cl$Stain %in% 'Muc5', ]

distrib_lobule_Muc5 <-ggplot(distrib_a_Muc5, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of Muc5 expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% Muc5+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "dodgerblue4")

distrib_a_CDX2 <- distrib_a_cl[distrib_a_cl$Stain %in% 'CDX2', ]

distrib_lobule_CDX2 <-ggplot(distrib_a_CDX2, aes(x=fct_reorder(Group_Case_Class, Tumor..Positive..), y=Tumor..Positive.., fill = Stain)) +
  geom_col(width = 0.7) +
  labs(title = "Distribution of CDX2 expression in the acinar invasion ROIs", subtitle = "All cases") +
  theme_clean() +
  labs(x = "Each case + stromal ROI combination", y = "% CDX2+ PDAC cells") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  scale_fill_manual(values = "lightsteelblue4")
```

# Basal stains in stroma. First, all pooled, followed by the individual stains.
```{r Draw basal in stroma distribution plots}

distrib_stroma_basal

distrib_stroma_CK17

distrib_stroma_CK5

distrib_stroma_CA125 

distrib_stroma_HMGA2
```

# Classical stains in stroma. First, all pooled, followed by the individual stains.
```{r Draw classical in stroma distribution plots}
distrib_stroma_classical

distrib_stroma_Muc5

distrib_stroma_CDX2
```

# Basal stains in lobular regions. First, all pooled, followed by the individual stains.
```{r Draw basal in lobular distribution plots}
distrib_lobule_basal

distrib_lobule_CK17

distrib_lobule_CK5 

distrib_lobule_HMGA2 

distrib_lobule_CA125 
```

# Classical stains in acinar. First, all pooled, followed by the individual stains.
```{r Draw classical in acinar distribution plots}
distrib_lobule_classical

distrib_lobule_Muc5

distrib_lobule_CDX2
```

Density plots, separated on compartment
using the datasets distrib_s and distrib_a

```{r density plot for all stains in stroma}
density_stroma <- distrib_s %>% 
  ggplot(aes(x = Tumor..Positive.., fill = Stain)) +
  geom_density(color="#e9ecef", alpha=0.8) +
  facet_grid(~Stain)

density_stroma

#As ridgeline
ridges_stroma <- distrib_s %>% 
  ggplot(aes(x = Tumor..Positive.., y = Stain, fill = Stain)) +
  geom_density_ridges(color="#e9ecef", alpha=0.8)

ridges_stroma

#As ridgeline with color gradient
ridges_gradient_stroma <- ggplot(distrib_s, aes(x = `Tumor..Positive..`, y = `Stain`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "E") +
  labs(title = 'Density of stains in stroma') +
  theme_clean() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )+
  xlab("% Positive cells") + ylab(NULL)
ridges_gradient_stroma

# Violin plots
violin_stroma_trimmed <- distrib_s %>% 
  ggplot(aes(x = Stain, y = Tumor..Positive.., fill = Stain)) +
  geom_violin(trim = TRUE) +
  theme_clean()

violin_stroma_trimmed

```

```{r density plots stains in Lobule}
density_lobule <- distrib_a %>% 
  ggplot(aes(x = Tumor..Positive.., fill = Stain)) +
  geom_density(color="#e9ecef", alpha=0.8) +
  facet_grid(~Stain)

density_lobule

#As ridgeline
ridges_lobule <- distrib_a %>% 
  ggplot(aes(x = Tumor..Positive.., y = Stain, fill = Stain)) +
  geom_density_ridges(color="#e9ecef", alpha=0.8)

ridges_lobule

#As ridgeline with color gradient
ridges_gradient_lobule <- ggplot(distrib_a, aes(x = `Tumor..Positive..`, y = `Stain`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "E") +
  labs(title = 'Density of stains in Lobule') +
  theme_clean() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
  xlab("% Positive cells") + ylab(NULL)
ridges_gradient_lobule

# Violin plots
violin_acinar_trimmed <- distrib_a %>% 
  ggplot(aes(x = Stain, y = Tumor..Positive.., fill = Stain)) +
  geom_violin(trim = TRUE) +
  theme_clean()

violin_acinar_trimmed

#Violin plots for both compartments
distrib_all <- all_combined[, c(1, 2, 3, 5, 11, 12)]

violin_both_untrimmed <- distrib_all %>% 
  ggplot(aes(x = Stain, y = Tumor..Positive.., color = Tissue_compartment)) +
  geom_violin(trim = FALSE, adjust = 2) +
  theme_calc()

violin_both_untrimmed
```


```{r Extracting hte number of cases in each heterogenity category}

Plasticity_ID <- m_Crows
Plasticity_ID$PKR <- rownames(Plasticity_ID)
# 'Basal' and 'Classical' both go to the moderate category


Plasticity_ID %>% count(Plasticity)
Plasticity_ID <- Plasticity_ID %>% mutate(Plasticity = ifelse(grepl('Basal', .$Plasticity), 'Moderate', .$Plasticity))
Plasticity_ID <- Plasticity_ID %>% mutate(Plasticity = ifelse(grepl('Classical', .$Plasticity), 'Moderate', .$Plasticity))
Plasticity_ID %>% count(Plasticity)

# Fraction of any type of switching; 24/31 ≈ 77,4%

range(all_combined_exl$Num.Detections)
# between 299-864 cells in ROIs.
```


```{r}
sessionInfo()
```

