---
title: "m-IF_immune panel"
author: "Sara Söderqvist"
date: "2024-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This script explores whether two macrophage types, defined below by their marker expression, are differentially expressed between lobule and stroma.

In brief, larger annotations of lobules and stroma were created in qupath (ROIs, regions of interest). Each ROI was in the vincinity of tumor. Within, cell detection was performed, and a clean-up to enrich for immune cells only was applied. The clean-up consisted of: ________________ removal of PanCK+ cells, removal of cells with very elongated nucleus (these would be more likely to be stromal cells and not immune cells).

In this pipeline, CD68 will be used as pan-macrophage marker, (nb, a pan-monocyte marker). Cells double positive for CD163 and CD68 are called M2 type of macrophage. CD163- CD69+ cells will be called M1 macrophages.

```{r required libraries}
library(rstatix)
library(dplyr)
library(reshape2)
library(ggplot2)
library(tidyr)
library(ggpubr)
library(ggthemes)

```


```{r Read in qupath export file}

setwd("/Users/sara.soderqvist/Library/CloudStorage/OneDrive-KarolinskaInstitutet/Writing_/Adaptive PDAC paper/Code to clean up and add in/read in") #Adjust working directory
immunepanel_cleaned = read.csv("read in_m-IF_immune.csv")

immunepanel_cleaned <- immunepanel_cleaned[, -c(3, 4, 7)]

```


```{r grouping columns and data format cleaned }

#Create individual columns displaying whether cells are positive for a given stain. This information is extracted from the column "Classification", as determined by the qupath classifier.
immunepanel_cleaned$CD68_pos <- factor(nrow(immunepanel_cleaned))
immunepanel_cleaned$CD163_pos <- factor(nrow(immunepanel_cleaned))
immunepanel_cleaned$CD56_pos <- factor(nrow(immunepanel_cleaned))
immunepanel_cleaned$NKp46_pos <- factor(nrow(immunepanel_cleaned))
immunepanel_cleaned$CD3_pos <- factor(nrow(immunepanel_cleaned))

immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD68_pos = ifelse(grepl('CD68_pos:', immunepanel_cleaned$Classification), '1', .$CD68_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD68_pos = ifelse(grepl('CD68_neg:', immunepanel_cleaned$Classification), '0', .$CD68_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD163_pos = ifelse(grepl(': CD163_pos', immunepanel_cleaned$Classification), '1', .$CD163_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD163_pos = ifelse(grepl(': CD163_neg', immunepanel_cleaned$Classification), '0', .$CD163_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD56_pos = ifelse(grepl('CD56_pos: ', immunepanel_cleaned$Classification), '1', .$CD56_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD56_pos = ifelse(grepl('CD56_neg: ', immunepanel_cleaned$Classification), '0', .$CD56_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(NKp46_pos = ifelse(grepl(' NKp46_pos: ', immunepanel_cleaned$Classification), '1', .$NKp46_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(NKp46_pos = ifelse(grepl(' NKp46_neg: ', immunepanel_cleaned$Classification), '0', .$NKp46_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD3_pos = ifelse(grepl('CD3_pos', immunepanel_cleaned$Classification), '1', .$CD3_pos))
immunepanel_cleaned <- immunepanel_cleaned %>% mutate(CD3_pos = ifelse(grepl('CD3_neg', immunepanel_cleaned$Classification), '0', .$CD3_pos)) # 176 422 cells, same as the original read in file.
# 1 row = 1 cell.

```

```{r cleanup including melt}

# the goal is to show what immune cells, specifically macrophages, exist in lobules and in stroma.
# Format the data to enable calculation of fractions of positive cells, depending on the ROI identity:

immunepanel_cleaned_cell <- immunepanel_cleaned[, -c(3, 5, 6)]

immune_cleaned_cell <- melt(immunepanel_cleaned_cell, id = c("Image", "Object.ID", "Parent"))
colnames(immune_cleaned_cell) [4] <- "Stain"
colnames(immune_cleaned_cell) [5] <- "Cell_pos"

immune_cleaned_cell <- separate(data = immune_cleaned_cell, col = "Image", into="Image", sep=c("_"), remove=TRUE)
immune_cleaned_cell$Group <- character(nrow(immune_cleaned_cell))
immune_cleaned_cell$Group <- paste(immune_cleaned_cell$Image, immune_cleaned_cell$Parent, immune_cleaned_cell$Stain, sep = "_")
```

```{r calculate fractions of cells in each ROI}
s_immune_cleaned_cell <- split(immune_cleaned_cell, f = immune_cleaned_cell$Group)

# Fraction calculation. Note that this does not take into account wheather individual cells are double positive etc. for other markers.
tot_det_immune_cleaned_cell <- lapply(X = s_immune_cleaned_cell, FUN = function (x) {
  x$sum_detections_ROI <- nrow(x)
  x
})

fun_immunecells_cleaned_frac <- function (z) {
  nr_pos<- z %>% filter(Cell_pos == "1") # get nr positive cells for each stain
  pos_fr <- nrow(nr_pos)/nrow(z) #calculate the fraction out of all the cells in that specific ROI
  z$pos_fraction <- pos_fr # put the calculated fraction in a column; "pos_fr"
  z
} 

immunecells_cleaned_calc_fr <- lapply (X = tot_det_immune_cleaned_cell, FUN = fun_immunecells_cleaned_frac)

immunecells_calc_frsub_cleaned <- bind_rows(immunecells_cleaned_calc_fr) #882 110 rows 
immunecells_calc_frsub_cleaned <- immunecells_calc_frsub_cleaned[, -c(2, 5)]

immunecells_calc_frsub_cleaned_roi <- unique(immunecells_calc_frsub_cleaned) # 370 rows

# Clearer grouping column; the compartment identity of the ROI (lobular or stromal)
immunecells_calc_frsub_cleaned_roi$Compartment <- factor(nrow(immunecells_calc_frsub_cleaned_roi))
immunecells_calc_frsub_cleaned_roi <- immunecells_calc_frsub_cleaned_roi %>% mutate(Compartment = ifelse(grepl('Lobule', immunecells_calc_frsub_cleaned_roi$Parent), 'Lobule', .$Compartment))
immunecells_calc_frsub_cleaned_roi <- immunecells_calc_frsub_cleaned_roi %>% mutate(Compartment = ifelse(grepl('Stroma', immunecells_calc_frsub_cleaned_roi$Parent), 'Stroma', .$Compartment))
```

```{r plot single positives}

boxes_pan_immune_cleaned_sp<- ggboxplot(immunecells_calc_frsub_cleaned_roi, x = "Stain", y = "pos_fraction", 
                        color = "Compartment", fill = "Compartment", alpha=0.6) +
  scale_color_brewer(palette= "Set2", type = "qual") +
  scale_fill_brewer(palette= "Set2", type = "qual") +
  labs(title = "Fractions of cells with positive immune cells' markers in lobular and stromal ROIs", subtitle = "General single markers' expression") +
  xlab(NULL) + ylab("Average positive cells per ROI") +
  theme_clean()
  
wilcox_pan_immune_cleaned <- immunecells_calc_frsub_cleaned_roi %>%
  group_by(Stain) %>%
  wilcox_test(pos_fraction ~ Compartment) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")

wilcox_pan_immune_cleaned <- wilcox_pan_immune_cleaned %>% add_xy_position(x = "Stain")
wilcox_pan_immune_cleaned


boxes_pan_immune_sp_cleaned_sign <- boxes_pan_immune_cleaned_sp + stat_pvalue_manual(label = "p.adj.signif",
                                                    wilcox_pan_immune_cleaned, tip.length = 0.02)
boxes_pan_immune_sp_cleaned_sign
```


Focused research question: are M1 and M2 macrophages (Mø) and their ratio differentially expressed between lobule and stroma?

1) How many of each Mø subtype are there, the total counts?
```{r subset: the macrophage markers CD68 and CD163}
# Recall subtypes:
# M2 mø = CD68+CD163+
# M1 mø = CD68+CD163-


#1) remove all cells that are positive for CD3 ( they are T cells)
#immunepanel_cleaned derived

non_CD3 <- dplyr::filter(immunepanel_cleaned, paste(CD3_pos) == '0')

# 2) keep all cells that are positive for either CD163 or CD68
mo <- non_CD3 %>%
  filter(CD163_pos == 1 | CD68_pos == 1)

mo$mo_subtype <- factor(nrow(mo))
mo <- mo %>% mutate(mo_subtype = ifelse(grepl('1', mo$CD68_pos) & grepl('1', mo$CD163_pos), 'CD68+_CD163+', .$mo_subtype))
mo <- mo %>% mutate(mo_subtype = ifelse(grepl('1', mo$CD68_pos) & grepl('0', mo$CD163_pos), 'CD68+_CD163-',.$mo_subtype))
mo <- mo %>% mutate(mo_subtype = ifelse(grepl('0', mo$CD68_pos) & grepl('1', mo$CD163_pos), 'CD68-_CD163+',.$mo_subtype))

mo$Compartment <- factor(nrow(mo))
mo <- mo %>% mutate(Compartment = ifelse(grepl('Lobule', mo$Parent), 'Lobule', .$Compartment))
mo <- mo %>% mutate(Compartment = ifelse(grepl('Stroma', mo$Parent), 'Stroma', .$Compartment))

```


```{r plot total counts of macrophage types}
bar_mo <- ggplot(mo, aes(x = mo_subtype, color = Compartment, fill = Compartment, alpha = 0.9)) +
  geom_bar(stat = "count", position = position_dodge2(width = 0.5, preserve = "single")) +
    scale_color_brewer(palette= "Set2", type = "qual") +
  scale_fill_brewer(palette= "Set2", type = "qual") +
  labs(title = "Total count of macrophage types") +
  theme_clean()
bar_mo

```


```{r melt and create histograms of mo subset}

mo_cell <- mo[, -c(3, 7:11)]

melt_mo_cell <- melt(mo_cell, id = c("Image", "Object.ID", "Parent", "mo_subtype", "Compartment"))
colnames(melt_mo_cell) [6] <- "Stain" 
colnames(melt_mo_cell) [7] <- "Intensity"

melt_mo_cell <- separate(data = melt_mo_cell, col = "Image", into="Image", sep=c("_"), remove=TRUE)
melt_mo_cell$Group <- character(nrow(melt_mo_cell))
melt_mo_cell$Group <- paste(melt_mo_cell$Image, melt_mo_cell$Parent, melt_mo_cell$Stain, sep = "_")

# histogram of mo:s intensity of the two stains of interest

hist_mo <- gghistogram(melt_mo_cell, x = "Intensity", color = "mo_subtype", alpha = 0.4, bins = 200, size = 0.1) +
  facet_grid(mo_subtype~Stain) +
  theme_clean()
hist_mo

```

```{r calculate fractions of macrophage subtypes per ROI mo}
s_mo <- split(melt_mo_cell, f = melt_mo_cell$Group)

tot_det_s_mo <- lapply(X = s_mo, FUN = function (x) {
  x$sum_detections_ROI <- nrow(x)
  x
})

# new functions to calculate mo subtype fractions in each ROI:
fun_immunecells_frac_mo <- function (z) {
  nr_M2<- z %>% filter(mo_subtype == "CD68+_CD163+") # get nr of M2 mo:s in the ROI
  M2pos_fr <- nrow(nr_M2)/nrow(z) #calculate the fraction of M2 mo:s out of all the cells in this specific ROI
  z$M2pos_fraction <- M2pos_fr # put M2 fraction in a column
  
  nr_M1<- z %>% filter(mo_subtype == "CD68+_CD163-") # get nr of M1 mo:s in the ROI
  M1pos_fr <- nrow(nr_M1)/nrow(z) #calculate the fraction of M1 mo:s out of all the cells in this specific ROI
  z$M1pos_fraction <- M1pos_fr # put M1 fraction in a column
  
  nr_CD163<- z %>% filter(mo_subtype == "CD68-_CD163+") # get nr of M1 mo:s in the ROI
  CD163pos_fr <- nrow(nr_CD163)/nrow(z) #calculate the fraction of M1 mo:s out of all the cells in this specific ROI
  z$CD163pos_fr <- CD163pos_fr # put cell id fraction in a column
  
  z
} 

immunecells_calc_fr_mo <- lapply (X = tot_det_s_mo, FUN = fun_immunecells_frac_mo)

immunecells_calc_frsub_mo <- bind_rows(immunecells_calc_fr_mo) #77 576 rows . 1 row for each cell.


immunecells_calc_frsub_mo <- immunecells_calc_frsub_mo[, -c(2, 4, 6:8)] #remove cell-specific row information: "group", enabling unique-ing to the ROIs in next command
immunecells_calc_frsub_roi_mo <- unique(immunecells_calc_frsub_mo) # 74 rows. 1 row for each ROI.

#melt it down

melt_mofractions <- melt(immunecells_calc_frsub_roi_mo, id = c("Image", "Parent", "Compartment", "sum_detections_ROI"), variable.name = "Immune_subtype", value.name = "Pos_fraction") # 222 rows = 74*3. 3 cell types per ROI.
head(melt_mofractions)


```

```{r boxplot mo subtype per compartment}

boxes_mo<- ggboxplot(melt_mofractions, x = "Immune_subtype", y = "Pos_fraction", 
                        color = "Compartment", fill = "Compartment", alpha=0.6) +
  scale_color_brewer(palette= "Set2", type = "qual") +
  scale_fill_brewer(palette= "Set2", type = "qual") +
  labs(title = "Fractions of macrophages and CD163 single+ in lobular and stromal ROIs", subtitle = "Only ROIs that have tumor in the immediate vicinity, only cells with positivity for either CD163 or CD68") +
  xlab(NULL) + ylab("Average positive cells per ROI") +
  theme_clean()

wilcox_mo <- melt_mofractions %>%
  group_by(Immune_subtype) %>%
  wilcox_test(Pos_fraction ~ Compartment) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")

t_test_mo <- melt_mofractions %>%
  group_by(Immune_subtype) %>%
  t_test(Pos_fraction ~ Compartment) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj")

t_test_mo

wilcox_mo <- wilcox_mo %>% add_xy_position(x = "Immune_subtype")
wilcox_mo


boxes_mo_sign <- boxes_mo + stat_pvalue_manual(label = "p.adj.signif",
                                                    wilcox_mo, tip.length = 0.01)
boxes_mo_sign


```




